### 2022-07-11 includes/acronyms.tex
updates on includes

\newacronym{hyq}{HyQ}{Hydraulically actuated Quadruped}

\newacronym{lf}{LF}{Left-Front}
\newacronym{rf}{RF}{Right-Front}
\newacronym{lh}{LH}{Left-Hind}
\newacronym{rh}{RH}{Right-Hind}

\newacronym{haa}{HAA}{Hip Adduction-Abduction}
\newacronym{hfe}{HFE}{Hip Flexion-Extension}
\newacronym{kfe}{KFE}{Knee Flexion-Extension}

\newacronym{imu}{IMU}{Inertial Measurement Unit}
\newacronym{dofs}{DoFs}{Degrees of Freedom}
\newacronym{rt}{RT}{Real Time}

\newacronym{com}{CoM}{Center of Mass}
\newacronym{cop}{CoP}{Center of Pressure}
\newacronym{zmp}{ZMP}{Zero Moment Point}
\newacronym{icp}{ICP}{Instantaneous Capture Point}
\newacronym{cp}{CP}{Capture Point}
\newacronym{cmp}{CMP}{Centroidal Moment Pivot}
\newacronym{grfs}{GRFs}{Ground Reaction Forces}

\newacronym{ls}{LS}{Least Square}

\newacronym{slip}{SLIP}{Spring Loaded Inverted Pendulum}
\newacronym{eom}{EoM}{Equation of Motions}
\newacronym{qp}{QP}{Quadratic Program}
\newacronym{sqp}{SQP}{Sequential Quadratic Programming}
\newacronym{mic}{MIC}{Mixed-Integer Convex}
\newacronym{cmaes}{CMA-ES}{Covariance Matrix Adaptation Evolution Strategy}
\newacronym{ara}{ARA*}{Anytime Repairing A*}
\newacronym{pca}{PCA}{Principal Component Analysis}
\newacronym{cpg}{CPG}{Central Pattern Generator}
\newacronym{wbc}{WBC}{Whole-Body Control}

\newacronym{mpc}{MPC}{Model Predictive Control}
\newacronym{ik}{IK}{Inverse Kinematic}
\newacronym{ocp}{OCP}{Optimal Control Problem}
\newacronym{nlp}{NLP}{Nonlinear Programming}
\newacronym{ltv}{LTV}{Linear Time Varying}

% SOFT TERRAIN ADAPTATION
\newacronym{awbc}{c$^3$WBC}{Compliant Contact Consistent Whole-Body Control}
\newacronym{swbc}{sWBC}{Standard Whole-Body Control}
\newacronym{c3wbc}{c$^3$WBC}{Compliant Contact Consistent Whole-Body Control}
\newacronym{ste}{TCE}{Terrain Compliance Estimator}
\newacronym{c3}{\texttt{c}$^3$}{compliant contact consistent}

\newacronym{stance}{STANCE}{\textbf{S}oft \textbf{T}errain \textbf{A}daptation a\textbf{n}d \textbf{C}ompliance \textbf{E}stimation}

\newacronym{wbopt}{WBOpt}{Whole Body Optimization}

\newacronym{hc}{HC}{Hunt and Crossley's}
\newacronym{kv}{KV}{Kelvin-Voigt's}

\newacronym{wllsr}{WLLSR}{Weighted Linear Least Squared Regression}

\newcommand{\grfs}{\gls{grfs}~}

\newacronym{mae}{MAE}{Mean Absolute Tracking Error}

\newacronym{ode}{ODE}{Open Dynamics Engine}

\newacronym{cmg}{CMG}{Control Moment Gyroscope}
\newacronym{ocs}{OCS}{Orientation Control System}

### 2022-07-11 mdpi22roscia.tex
Intro

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal,letterpaper]{IEEEtran}
% include the list of acronyms, math commands and new commands used in this paper
\input{includes/used_packages.tex}
\input{includes/acronyms.tex}
\input{includes/math_commands.tex}
\input{includes/newcommands.tex}
\hypersetup{draft}
\makeglossaries

\title{Reaction Wheels: Enhancing Aerial Maneuvers for
	Legged Robots (Tentative)}
\author{Andrea Cumerlotti$^{1, \, 2}$, Francesco Roscia$^{2, \, 3}$, Michele Focchi$^{2, \, 4}$, Andrea Del Prete$^{1}$ and Claudio Semini$^2$
	\thanks{$^1$ Industrial Engineering Department (DII), University of Trento, Trento, Italy.

	$^2$ Dynamic Legged Systems (DLS) lab, Istituto Italiano di Tecnologia (IIT), Genoa, Italy.

	$^3$ Computer Science and Technology, Bioengineering, Robotics and Systems Engineering (DIBRIS), University of Genoa, Genoa, Italy.

	$^4$ Department of Information Engineering and Computer Science (DISI), University of Trento, Trento, Italy.

}}

\begin{document}
\maketitle
\thispagestyle{empty}
\pagestyle{empty}

\begin{abstract}%150-250 word abstract
Aerial motions represent a challenge for quadrupedal robots. They become inevitable to overpass obstacles that cannot be circumvent with standard locomotion gaits. In these cases, the robot must perform a leap to position itself on the obstruction or even flyover it. In this work we propose an orientation control system consisting on a couple of rotating and actuated masses (named flywheels or reaction wheels) for the lightweight open-source robot Solo12. Because of the conservation of angular momentum, their rotation can be adjusted to steer the robot base orientation even when there are no contacts with the ground. The axes of rotation of the flywheels are designed to be incident, allowing us to have a compact orientation control system that is capable of controlling both roll and pitch angles of the platform considering the different moment of inertia in the two directions. We proof the concept with simulation and experiments on the real robot.
\end{abstract}

\begin{IEEEkeywords}

\end{IEEEkeywords}

\section{Introduction}\label{sec:introduction}
Legged locomotion is designed for traversing rough terrain.
Different types of gait, such as trot or crawl, have been developed to move quadrupedal robots. 
Thanks to the progress of the last two decades, robots become lighter and able to generate higher torques and forces at the joints, enabling the possibility of doing highly dynamic maneuvers.
Sometimes it is not possible to get around an obstacle with the gaits mentioned above, and jumps may be required.

When the robot is in air, the \acrfull{com} moves on the ballistic trajectory, that is completely defined by the lift-off position and velocity. On the other hand, the base orientation can be changed exploiting the conservation of the system angular momentum. This means that it is possible to control the base angular velocity by changing the inertia of the robot, e.g., changing the joints configuration.

Nevertheless, the majority of quadrupeds are designed to respect the massless leg assumption, resulting in limbs having small influence on the total angular momentum.

\subsection{Related Work}
Quadrupedal animals, like cats, can rearrange the tail and trunk to correct the orientation during a fall \cite{kane1969dynamical}.
% Tail
Many works use an additional link as a tail, like in \cite{chu2019null} and \cite{wenger2016frontal}.
This link rotates around an axis that does not pass through the robot \acrshort{com}: the distances between the axis of rotation and the \acrshort{com} of both trunk and tail allow to obtain large effect on the total angular momentum even with a small tail mass.
However, the placement of the additional link makes the resulting robot asymmetric. Moreover, due to its limited range of motion, a tail can be used only for a single jump, not for a repeated sequence \cite{johnson2012tail}.\annotation{Andrea me la spieghi?}

% Legs reconfig
It is possible to obtain a similar result by creating repetitive circular motions with the feet, like in \cite{hoffman2021exploiting} and \cite{kurtz2021mini}.

In the latter, authors proposes special heavy boots for Mini Cheetah and use a neural network to calculate online joint trajectory. However, this solution unnecessarily increases the inertia of the legs, complicating the locomotion problem.\annotation{trova un modo per dire che questa complicazione è data dalla massless assumption che non è più valida}

% Gyros
Another option is to use a \acrfull{cmg}.
It consists of a wheel, spinning at a constant angular velocity inside two or three actuated gimbals.
Tilting the wheel's axes of rotation generates the gyroscopic torque.
This system is widespread in spacecrafts reorientation \cite{yoon2002spacecraft}, but less frequently exploited in robot locomotion, either wheeled \cite{brown1996single} or legged \cite{mikhalkov2021gyrubot}.
The \acrshort{cmg} presents interesting capabilities, but the presence of a pan-tilt \annotation{Andrea mi spieghi che intendi} unit to the drive the gyroscope makes it impractical to mount it on a small, lightweight robot.

% Reaction wheel
Reaction wheels represent an additional option for controlling the robot orientation.
Changing the angular velocity of a rotating mass attached to the trunk generates a torque that can reorient and stabilize the robot. Also this device is borrowed by spacecraft orientation \cite{oland2009reaction}, and it was sporadically investigated in legged locomotion, both for bipeds \cite{Brown2016}  \cite{xiong2020sequential} and quadrupeds \cite{kolvenbach2019towards}, \cite{vasilopoulos2016quadruped}.
Compared with tails, reaction wheels do not have position limit, and since it rotates around its center of mass, its angular momentum results holonomic \cite{machairas2015quadruped}.\annotation{Andrea mi spieghi che intendi}
To get a fast response, it is necessary to have an abrupt change in the reaction wheel angular velocity (angular acceleration).
Using a brake avoids the employment of a motor able to deliver higher torques \cite{gajamohan2012cubli}, keeping the system compact: the motor speeds up slowly the reaction wheel, and when a reorientation is required, the break stops its spin.
Since the effect of the break is unidirectional, it is possible to generate a rotation of the base only in the opposite direction of the reaction wheel angular velocity.

\subsection{Proposed Approach and Contribution}
In this work, we present an \acrlong{ocs} based on two flywheels, mounted on the trunk of the quadruped Solo12 \cite{grimminger2020open}. In order to control roll and pitch orientations with a compact device, the axes of rotation are set to be incident, creating an angle that balances the effect in both the directions. 
Flywheels can correct orientation errors due to disturbances during the flight and inaccuracies in to the angular momentum achieved at the lift-off (e.g., given by tracking issues and non-idealities). 
They enable the robot to land with a desired angular velocity (possibly zero) and orientation.
In addition, they can enhance the landing phase by significantly reducing oscillations.  
The presence of these additional joints, whose only function is to control the orientation, gives the possibility to relieve the effort of the legs.
In more complex scenarios, like in a somersault, legs and flywheels can work in parallel to achieve a rotation angle larger than the one achievable only with legs (e.g., due to torque limitation).

\subsection{Outline}
%

\section{Background}

\section{Conclusions}
\label{sec:conclusion}

\small
\section*{Acknowledgements}

\printbibliography
\end{document}

### 2022-07-11 mdpi22roscia.tex
background and design (inertia selection in review)

@@
 \subsection{Proposed Approach and Contribution}
-In this work, we present an \acrlong{ocs} based on two flywheels, mounted on the trunk of the quadruped Solo12 \cite{grimminger2020open}.
+In this work, we present an \acrfull{ocs} based on two flywheels, mounted on the trunk of the quadruped Solo12 \cite{grimminger2020open}.
 In order to control roll and pitch orientations with a compact device, the axes of rotation are set to be incident, creating an angle that balances the effect in both the directions.
@@
 The presence of these additional joints, whose only function is to control the orientation, gives the possibility to relieve the effort of the legs.
-In more complex scenarios, like in a somersault, legs and flywheels can work in parallel to achieve a rotation angle larger than the one achievable only with legs (e.g., due to torque limitation).
+In more complex scenarios, like for a somersault, legs and flywheels can work in parallel to achieve a rotation angle larger than the one achievable only with legs (e.g., due to torque limitation).
+Furthermore, the \acrshort{ocs} can also be exploited to stabilize dynamic gaits, compensating for trunks oscillation.
 
@@
 \section{Background}
+The starting point for any \acrshort{ocs} is the Euler's equation.
+For any mechanical system, the time derivative of the angular momentum computed with the respect of with a reference point $O$ fixed in an inertial frame equals
+the resultant torque with respect to the same reference point:
+\begin{equation}
+	\dot{\bm{L}} = \sum_i \bm{M}_i
+\end{equation}
+When the external moment applied to the system is zero, Euler’s equation simplifies in
+\begin{equation}
+\dot{\bm{L}} = 0 \quad \Rightarrow \quad \bm{L}(t) = const.
+\end{equation}
+which is known as conservation of angular momentum.
+Referring to legged robots, this condition occurs whether the system is not in contact with the ground or other objects, e.g., during a fall or the flight phase of a jump.
+In the case, it is possible to change the angular velocity of the base, changing the joints velocities, as a result of the \textit{non-holonomy} of the angular momentum \cite{Wieber16}: if the angular momentum of a certain body changes, the one of the others must modify to maintain the total sum constant.
 
+\section{\acrshort{ocs} Design}
+We want to control simultaneously the robot roll and pitch orientation.
+To accomplish this task, the reaction wheels' contribution to the total angular momentum should lie on both $x$- and $y$-axes of the base frame, see Fig. \ref{fig:angVel}. 
+To remove unnecessary complications, we design two symmetrical modules.
+The axes of rotation of the left and right wheel, identified in the base reference frame with the unit vectors $\hat{\bm{a}}_l$ and $\hat{\bm{a}}_r$\annotation{$\hat{\bm{a}}$ sta per axis, se non ci piace si può trasformare in $\hat{\bm{c}}$}, are set to be incident, spanning a plane parallel to the $xy-$ plane of the base reference frame.
+We denote with $\alpha \leq \pi/2$ the non-negative angle between the wheels'axes of rotation and the robot lateral direction\annotation{I suggest to call $\alpha$ incident angle}.
+As long as the matrix
+\begin{equation}
+	\bm{C} = 
+	\left[ \begin{array}{cc}
+	\hat{\bm{a}}_l & \hat{\bm{a}}_r
+	\end{array} \right] = 
+	\left[ \begin{array}{cc}
+	\sin(\alpha) & \sin(\alpha) \\
+	-\cos(\alpha) & \cos(\alpha) \\
+	0 & 0
+	\end{array} \right]
+\end{equation}
+is full column rank, it is possible to control both roll and pitch angles.
+If $\alpha = 0$, the roll results to be uncontrollable through the \acrshort{ocs}; otherwise, if $\alpha = \pi/2$, the pitch is not uncontrollable.
+%
+%\subsection{Inertia Selection}
+%n order to find the inertia necessary to comply with the requirements of the orientation control system, some analyses are performed with the “Elroy’s Beanie” shown in figure \ref{fig:elroy-beanie}.
+%This model consists of two bodies connected with a revolute joint on their Center of Mass (CoM).
+%One of the bodies represents the robot (inertia $I_r$), simplified as a single rigid body, and the other one the flywheel (inertia $I_{fw}$).
+%I perform the analysis for only one flywheel and only then extend the results to two bodies.
+%\begin{figure}[H] 
+%	\centering
+%	\includegraphics[height=5.9cm]{Images/Flywheel_System_Design/elroys-beanie.pdf}
+%	\caption{Schematic representation of the Elroy's Beanie model used for the preliminary analysis of the pitch motion.}
+%	\label{fig:elroy-beanie}
+%\end{figure}
+%\noindent
+%The angular momentum $L$ of this system can be written as
+%\begin{equation}\label{eq:ang_mom_elroy}
+%L = \left(I_r + I_{fw}\right)\dot{\theta} + I_{fw}\dot{\gamma}
+%\end{equation}
+%in which $I_r$ is the $I_{yy}$ component of the centroidal inertia matrix of the robot and $I_{fw}$ is composed by both the reaction wheels inertia of the along their axis of rotation, $\dot{\theta}$ and $\dot{\gamma}$ correspond respectively to the pitch-rate and angular speed of the wheels.
+%I employed the only the $I_{yy}$ component because, it gives the information about the pitch rotation.
+%The other elements of the tensor of inertia that affect the pitch, such as $I_{yx}$ and $I_{yz}$, are not considered since they are negligible with the respect to  $I_{yy}$.
+%Using the $I_{xx}$ component of the centroidal inertia matrix instead of the $I_{yy}$ allows us to perform the the same analysis as below for the roll rotation.
+%In this case, the variable $\theta$ should be substitute by $\phi$, the roll angle.
+%Under the condition of conservation of the angular momentum \ref{sec:cons_ang_mom} it is possible to estimate of the lower bound for $I_{fw}$ to get a desired angular velocity of the main body $\dot{\theta}_{des}$.
+%Assuming that at the lift-off the reaction wheels are not spinning ($\dot{\gamma}_{lo} = 0$), the angular momentum is $L_{lo} = \left(I_r + I_{fw}\right)\dot{\theta}_{lo}$, in which $\dot{\theta}_{lo}$ is the base pitch rate at lift-off.
+This value does not change for the whole duration of the flight.
+%The minimum inertia necessary to reach $\dot{\theta}_{des}$ is obtained when the flywheels rotate at their maximum speed. For higher inertia it is sufficient a smaller angular velocity:
+%\begin{equation}
+%I_{fw} = I_r \frac{\dot{\theta}_{lo} - \dot{\theta}_{des}}{\dot{\theta}_{des} + \dot{\gamma}_{max} - \dot{\theta}_{lo}} \ .
+%\end{equation}
+%The variable $\dot{\gamma}_{max}$ is the maximum motor speed that correspond to $5000 \ \mathrm{RPM}$ ( $523.6 \ \mathrm{rad/s}$).
+%In figure \ref{fig:pre_ana_kin} the results for different initial and desired angular velocities are reported.
+%It is evident that high inertia is required when the inertial and the desired angular velocity are opposite.
+%Usually, this condition is avoided thanks to a good thrusting phase.
+%\begin{figure}[H]
+%	\centering
+%	\includegraphics[width=0.7\textwidth]{Images/Flywheel_System_Design/Inertia_selection_by_kinematic_limits.eps}
+%	\caption{Minimum inertia necessary to obtain the angular velocity $\dot{\theta}_{des}$ starting with different lift-off $\dot{\theta}_{lo}$ conditions.}
+%	\label{fig:pre_ana_kin}
+%\end{figure}
+%
+%The kinematic model employed before has not considered the transient to reach the desired angular speed.
+%For this reason, the dynamical model of the "Elroy's Beanie" mechanism is derived.
+%The dynamics of this system is obtained with the Lagrangian method.
+%To calculate the Lagrangian $L$ it is necessary to compute the kinetic energy $K$ and the potential energy $U$.
+%In this quantities must be included also the translation part, therefore are used the masses of the bodies, $m_r$ and $m_{fw}$, and the position of the CoM, $x$ and $z$. 
+%\begin{equation*}
+%K = \frac{1}{2}\left(m_r + m_{fw}\right)\left(\dot{x}^2 + \dot{z}^2\right) + \frac{1}{2}I_r \dot{\theta}^2 + \frac{1}{2} I_{fw} \left(\dot{\theta} + \dot{\gamma}\right)^2
+%\end{equation*}
+%\begin{equation*}
+%U = \left(m_r + m_{fw}\right) g z 
+%\end{equation*}
+%\begin{equation}
+%\begin{split}\label{eq:lagrangian}
+%L &= K - U\\
+%&= \frac{1}{2}\left(m_r + m_{fw}\right)\left(\dot{x}^2 + \dot{z}^2\right) + \frac{1}{2}\left(I_r + I_{fw}\right)\dot{\theta}^2 + \frac{1}{2} I_{fw}\dot{\gamma}^2 + I_{fw} \dot{\theta}\dot{\gamma} - \left(m_r + m_{fw}\right) g z
+%\end{split}
+%\end{equation}
+%\begin{equation}\label{eq:LagToEqMo}
+%\frac{\partial L}{\partial \mathbf{q}} - \frac{d}{dt}\left(\frac{\partial L}{\partial \dot{\mathbf{q}}}\right) =  \boldsymbol{\tau}
+%\end{equation}
+%From the Lagrangian $L$ is possible to derive the equation of motion using \ref{eq:LagToEqMo}, in which $\mathbf{q} = \left[\begin{array}{cccc} x & z & \theta & \gamma \end{array}\right]^T$ are the generalized coordinates and $ \boldsymbol{\tau} $ are the generalized forces applied to the bodies.
+%In this particular case, it is analysed the flight phase in which are not applied external forces or torques, apart from the flywheels' ones.
+%Then we can set $\boldsymbol{\tau} = \left[\begin{array}{cccc} 0 & 0 & 0 & \tau_m \end{array}\right]^T$, in which $\tau_m$ is the direct-drive torque applied by the motor that control the flywheel.
+%The resulting equation of motion are
+%\begin{equation}\label{eq:EqOfMo}
+%\left[\begin{array}{cccc}
+%m_r + m_{fw} & 0 & 0 & 0\\
+%0 & m_r + m_{fw} & 0 & 0\\
+%0 & 0 & I_r+I_{fw} & I_{fw}\\
+%0 & 0 & I_{fw} & I_{fw} \\
+%\end{array}\right]
+%\left[\begin{array}{c}
+%\ddot{x} \\ \ddot{y} \\ \ddot{\theta} \\ \ddot{\phi} \\
+%\end{array}\right] + 
+%\left[\begin{array}{c}
+%0 \\ \left(m_r + m_{fw}\right) g \\ 0 \\ 0 \\
+%\end{array}\right] = 
+%\left[\begin{array}{c}
+%0 \\ 0 \\ 0 \\ \tau \\
+%\end{array}\right] \ .
+%\end{equation}
+%To obtain the values of $\mathbf{q}$, the differential equations are integrated using Runge-Kutta 4th order method.
+%From \ref{eq:EqOfMo}, we can see that the linear and the angular dynamic are decoupled.
+%The linear part, which corresponds to the well-known motion of a projectile, is used to calculate the time of flight.
+%I suppose that the robot lands on a flat surface at height zero.
+%The landing is estimated when the CoM reach the height of $0.23 \ \mathrm{m}$ (distance between CoM and ground in homing configuration) with a negative velocity.
+%This information is needed to understand the maximum time in which the reorientation maneuver must be concluded.
+%The linear dynamics can be neglected, imposing a fixed time constraint based on the specification of the duration of the flight phase to $1 \ \mathrm{s}$.
+%
+%The inertia of the reaction wheels can be chosen minimizing the work done by the motor to perform a reorientation maneuver.
+In particular, the task analyzed was a rotation of the main body of $30^\circ$ ($\simeq 0.5 \ \mathrm{rad}$) in the time  $t_f = 1 \ \mathrm{s}$.
+%A fifth-order polynomial generates the desired trajectory of the pitch of the robot that performs the rotation starting and finishing with zero angular velocity and acceleration.
+%In order to track the trajectory, we assume that we use a proportional and derivative (PD) controller.
+%The gains $k_p$ and $k_d$ are chosen to obtain a small settling time avoiding overshoots.
+%\begin{equation*}
+%\tau_m(t) = k_p \left(\theta_{des}(t) - \theta(t)\right) + k_d \left(\dot{\theta}_{des}(t) - \dot{\theta}(t)\right)
+%\end{equation*}
+%To select the inertia we intend to perform an optimization where we minimize the integral of the motor power absolute value, during the reorientation task.
+%The inertia $I_r$ affects the dynamics of the system and then modifies the evolution of $\tau_m(t)$ and $\dot{\gamma}(t)$ during the task.
+%\begin{equation*}
+%\min_{I_r} \int_{0}^{t_f} |\tau_m(t)\dot{\gamma}(t)| \,dt
+%\end{equation*}
+%In figure \ref{fig:inertia_vs_energy} are shown the values of the work done by the motors,for correcting both roll and pitch of Solo12.
+The dissipated energy is monotonically decreasing in the flywheel's inertia.
+%From the results, it is evident that there is not a local minimum and the optimal value for the inertia $I_{fw}$ is the greatest as possible.
+%\begin{figure}[H]
+%	\centering
+%	\includegraphics[height=6cm]{Images/Flywheel_System_Design/Inertia_vs_energy.eps}
+%	\caption{Work done by the motor $W_m$ using different values of inertia $I_{fw}$ for the reaction wheel  during a reorientation manoeuvrer of $30^\circ$ in $1s$.
+%		The $x$-axis is in logarithmic scale to show a wide spectrum of inertias.
+%		The results of the roll (in red) and pitch (in blue) are difference due to the different value of $I_r$, indeed the components of the centroidal inertia of the robot $I_{xx}$ and $I_{yy}$ are very different.}
+%	\label{fig:inertia_vs_energy}
+%\end{figure}
+%
+%\noindent
+%With the same analysis, I investigated whether the inclusion of a gearbox between the motor and the reaction wheel would bring some benefits.
+%The dynamical model is modified including the gearbox, with a gear reduction ratio $n$.
+%In this case, the motor angular velocity and the one of the reaction wheel are no more equal.
+%I continue to use $\gamma$ to define the motor angle, while the one of the flywheel is $\gamma / n$.
+%The torque applied at the reaction wheel is the torque of the motor magnified by a a factor $n$.
+%\begin{equation}\label{eq:EqOfMo_wGB}
+%\left[\begin{array}{cccc}
+%m_r + m_{fw} & 0 & 0 & 0\\
+%0 & m_r + m_{fw} & 0 & 0\\
+%0 & 0 & I_r+I_{fw} & I_{fw}/n\\
+%0 & 0 & I_{fw} & I_{fw}/n \\
+%\end{array}\right]
+%\left[\begin{array}{c}
+%\ddot{x} \\ \ddot{y} \\ \ddot{\theta} \\ \ddot{\gamma} \\
+%\end{array}\right] + 
+%\left[\begin{array}{c}
+%0 \\ \left(m_r + m_{fw}\right) g \\ 0 \\ 0 \\
+%\end{array}\right] = 
+%\left[\begin{array}{c}
+%0 \\ 0 \\ 0 \\ n\tau_m \\
+%\end{array}\right]
+%\end{equation}
+%The problem of minimization is modified including the gear reduction ratio as a decision variable.
+%\begin{equation}
+%\min_{I_r,\, n} \int_{0}^{t_f} |\tau_m(t)\dot{\gamma}(t)| \,dt
+%\end{equation}
+%The results, in figure \ref{fig:gearbox_vs_energy}, point out that the presence of a gearbox with a reduction ratio as great as possible results in a minimization of the energy used to perform the reorientation maneuvers.
+%The simulation of the task points out that adding the gearbox reduce the tracking error, but decrease of a factor $1/n$ the maximum speed of the motor.
+%I decide to maintain the value of $n = 1$ (without gearbox), to avoid the saturation of the motor velocity during a common task and because using a greater value do not comport a relevant decrease in energy.
+%\begin{figure}[H]
+%	\centering
+%	\includegraphics[height=6cm]{Images/Flywheel_System_Design/gearbox_vs_energy.eps}
+%	\caption{Work done by the motor $W_m$ using different values of reduction ratio $n$ for the reaction wheel  during a reorientation manoeuvrer of $30^\circ$ in $1s$.
+%		The $x$-axis is in logarithmic scale to show a wide spectrum of inertias.
+%		The results of the roll (in red) and pitch (in blue) are difference due to the different value of $I_r$, indeed the components of the centroidal inertia of the robot $I_{xx}$ and $I_{yy}$ are very different.}
+%	\label{fig:gearbox_vs_energy}
+%\end{figure}
+%\subsection{Incident Angle Selection}
+%
+%Let $\bm{I}^r = \mathrm{diag}{I^r_xx, \, I^r_yy, \, I^r_zz}$ be the inertia tensor of the robot in nominal configuration, expressed at its \acrshort{com}.
+The angle $\alpha$ that allows to maximize the joint contribution to the angular momentum on both the directions is
+\begin{figure}
+	\centering
+	\includegraphics[width=0.7\linewidth]{figures/angular_velocities.eps}
+	\caption{Representation of the flywheels rotation axes, seen from the top of the robot}
+	\label{fig:angVel}
+\end{figure}
 \section{Conclusions}

### 2022-07-15 mdpi22roscia.tex
update

--- 2022-07-11 mdpi22roscia.tex
@@
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[journal,letterpaper]{IEEEtran}
+\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper
+
+%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper
+
+\IEEEoverridecommandlockouts                              % This command is only needed if 
+% you want to use the \thanks command
+
+\overrideIEEEmargins                                      % Needed to meet printer requirements.
+
+%\documentclass[journal,letterpaper]{IEEEtran}
 % include the list of acronyms, math commands and new commands used in this paper
@@
 
-\title{Reaction Wheels: Enhancing Aerial Maneuvers for
-	Legged Robots (Tentative)}
-\author{Andrea Cumerlotti$^{1, \, 2}$, Francesco Roscia$^{2, \, 3}$, Michele Focchi$^{2, \, 4}$, Andrea Del Prete$^{1}$ and Claudio Semini$^2$
-	\thanks{$^1$ Industrial Engineering Department (DII), University of Trento, Trento, Italy.
-
-	$^2$ Dynamic Legged Systems (DLS) lab, Istituto Italiano di Tecnologia (IIT), Genoa, Italy.
-
-	$^3$ Computer Science and Technology, Bioengineering, Robotics and Systems Engineering (DIBRIS), University of Genoa, Genoa, Italy.
-
-	$^4$ Department of Information Engineering and Computer Science (DISI), University of Trento, Trento, Italy.
-
+\title{Orientation Control System: Enhancing Aerial Maneuvers \\ for Quadruped Robots}
+\author{
+	Francesco Roscia$^{1}$, Andrea Cumerlotti$^{1, \, 2}$, Andrea Del Prete$^{2}$, Claudio Semini$^{1}$, and Michele Focchi$^{1, \, 3}$
+	\thanks{$^1$ Dynamic Legged Systems (DLS) lab, Istituto Italiano di Tecnologia (IIT), Genoa, Italy.
+		$^2$ Industrial Engineering Department (DII), University of Trento, Trento, Italy.
+		$^3$ Department of Information Engineering and Computer Science (DISI), University of Trento, Trento, Italy.
 }}
@@
 \begin{abstract}%150-250 word abstract
-Aerial motions represent a challenge for quadrupedal robots.
-They become inevitable to overpass obstacles that cannot be circumvent with standard locomotion gaits.
-In these cases, the robot must perform a leap to position itself on the obstruction or even flyover it.
-In this work we propose an orientation control system consisting on a couple of rotating and actuated masses (named flywheels or reaction wheels) for the lightweight open-source robot Solo12.
-Because of the conservation of angular momentum, their rotation can be adjusted to steer the robot base orientation even when there are no contacts with the ground.
-The axes of rotation of the flywheels are designed to be incident, allowing us to have a compact orientation control system that is capable of controlling both roll and pitch angles of the platform considering the different moment of inertia in the two directions.
-We proof the concept with simulation and experiments on the real robot.
+For legged robots, aerial motions are the only option to overpass obstacles that cannot be circumvent with standard locomotion gaits.
+In these cases, the robot must perform a leap to either jump onto the obstacle or even fly over it.
+However, these movements represent a challenge because during the flight phase the center of mass cannot be controlled, and the robot orientation has limited controllability.
+This paper focuses on the latter issue and proposes an \gls{ocs} consisting of two rotating and actuated masses (flywheels or reaction wheels) to gain control authority on the robot orientation.
+Because of the conservation of angular momentum, their rotational velocity can be adjusted to steer the robot orientation even when there are no contacts with the ground.
+The axes of rotation of the flywheels are designed to be incident, leading to a compact orientation control system that is capable of controlling both roll and pitch angles, considering the different moment of inertia in the two directions.
+We tested the concept with simulations on the robot Solo12.
 \end{abstract}
 
-\begin{IEEEkeywords}
+%\begin{IEEEkeywords}
 
-\end{IEEEkeywords}
+%\end{IEEEkeywords}
 
 \section{Introduction}\label{sec:introduction}
-Legged locomotion is designed for traversing rough terrain.
-Different types of gait, such as trot or crawl, have been developed to move quadrupedal robots. 
-Thanks to the progress of the last two decades, robots become lighter and able to generate higher torques and forces at the joints, enabling the possibility of doing highly dynamic maneuvers.
-Sometimes it is not possible to get around an obstacle with the gaits mentioned above, and jumps may be required.
+Legged robots are designed for traversing rough terrain.
+Different types of gaits, such as trot or crawl, have been developed for quadrupedal robots. 
+Thanks to the progress of the last two decades, robots have become lighter and stronger, which has enabled them to perform agile locomotion.
+However, sometimes it is not possible to get around or over an obstacle with the gaits mentioned above, and jumps are required.
 
-When the robot is in air, the \acrfull{com} moves on the ballistic trajectory, that is completely defined by the lift-off position and velocity.
+When the robot is in the air, the \gls{com} moves on the ballistic trajectory, that is completely defined by the lift-off position and velocity.
 On the other hand, the base orientation can be changed exploiting the conservation of the system angular momentum.
@@
 
-Nevertheless, the majority of quadrupeds are designed to respect the massless leg assumption, resulting in limbs having small influence on the total angular momentum.
+Nevertheless, the majority of quadrupeds are designed with light legs, resulting in limbs that have small influence on the total angular momentum.
 
-\subsection{Related Work}
-Quadrupedal animals, like cats, can rearrange the tail and trunk to correct the orientation during a fall \cite{kane1969dynamical}.
+\begin{figure}[t!]
+	\centering
+	\includegraphics[width=.7\linewidth]{figures/solo_flywheels_pic.png}
+	\caption{\small The proposed Orientation Control System for Solo12 has two flywheels with incident rotation axes. Each wheel is located in a 3D-printed shell and mounted on the trunk body.}
+	\label{fig:solo12flywheels}
+\end{figure}
+
+Quadrupedal animals, like cats, can rearrange the tail and trunk to correct the orientation during a fall~\cite{kane1969dynamical}.
 % Tail
-Many works use an additional link as a tail, like in \cite{chu2019null} and \cite{wenger2016frontal}.
-This link rotates around an axis that does not pass through the robot \acrshort{com}: the distances between the axis of rotation and the \acrshort{com} of both trunk and tail allow to obtain large effect on the total angular momentum even with a small tail mass.
+Many works in the field of robotics used an additional link as a tail, like in \cite{chu2019null} and \cite{wenger2016frontal}.
+This link rotates around an axis that does not pass through the robot \gls{com}: the distances between the axis of rotation and the \gls{com} of both trunk and tail result in a large effect on the total angular momentum even with a small tail mass.
 However, the placement of the additional link makes the resulting robot asymmetric.
-Moreover, due to its limited range of motion, a tail can be used only for a single jump, not for a repeated sequence \cite{johnson2012tail}.\annotation{Andrea me la spieghi?}
-
+Moreover, due to its limited range of motion, a tail can be used only for a single jump, not for a repeated sequence~\cite{johnson2012tail}.
 % Legs reconfig
 It is possible to obtain a similar result by creating repetitive circular motions with the feet, like in \cite{hoffman2021exploiting} and \cite{kurtz2021mini}.
-
-In the latter, authors proposes special heavy boots for Mini Cheetah and use a neural network to calculate online joint trajectory.
-However, this solution unnecessarily increases the inertia of the legs, complicating the locomotion problem.\annotation{trova un modo per dire che questa complicazione è data dalla massless assumption che non è più valida}
-
+In the latter work, the authors proposed special heavy boots for Mini Cheetah and used a neural network to calculate online joint trajectories.
+However, this solution unnecessarily increases the inertia of the legs, complicating the locomotion problem, which can no longer rely on the massless leg assumption.
+%\annotation{trova un modo per dire che questa complicazione è data dalla massless assumption che non è più valida}
 % Gyros
-Another option is to use a \acrfull{cmg}.
+Another option is to use a \gls{cmg}.
 It consists of a wheel, spinning at a constant angular velocity inside two or three actuated gimbals.
-Tilting the wheel's axes of rotation generates the gyroscopic torque.
+Tilting the wheel's axis of rotation generates the gyroscopic torque.
 This system is widespread in spacecrafts reorientation \cite{yoon2002spacecraft}, but less frequently exploited in robot locomotion, either wheeled \cite{brown1996single} or legged \cite{mikhalkov2021gyrubot}.
-The \acrshort{cmg} presents interesting capabilities, but the presence of a pan-tilt \annotation{Andrea mi spieghi che intendi} unit to the drive the gyroscope makes it impractical to mount it on a small, lightweight robot.
+The \gls{cmg} presents interesting capabilities, but the presence of a pan-tilt unit to drive the gyroscope makes it impractical to mount it on a small, lightweight robot.
 
@@
 Reaction wheels represent an additional option for controlling the robot orientation.
-Changing the angular velocity of a rotating mass attached to the trunk generates a torque that can reorient and stabilize the robot.
-Also this device is borrowed by spacecraft orientation \cite{oland2009reaction}, and it was sporadically investigated in legged locomotion, both for bipeds \cite{Brown2016}  \cite{xiong2020sequential} and quadrupeds \cite{kolvenbach2019towards}, \cite{vasilopoulos2016quadruped}.
-Compared with tails, reaction wheels do not have position limit, and since it rotates around its center of mass, its angular momentum results holonomic \cite{machairas2015quadruped}.\annotation{Andrea mi spieghi che intendi}
+Changing the angular velocity of a rotating mass attached to the trunk generates a torque that can reorient the robot.
+This device has been first applied for spacecraft orientation \cite{oland2009reaction}, and later it has been only sporadically investigated in legged locomotion for controlling pitch orientation, both for bipeds \cite{Brown2016}, \cite{xiong2020sequential} and quadrupeds \cite{kolvenbach2019towards}, \cite{vasilopoulos2016quadruped}.
+Compared with tails, reaction wheels do not have position limit, and since their rotation axes pass though the \gls{com}, its angular momentum results holonomic \cite{machairas2015quadruped}.
 To get a fast response, it is necessary to have an abrupt change in the reaction wheel angular velocity (angular acceleration).
-Using a brake avoids the employment of a motor able to deliver higher torques \cite{gajamohan2012cubli}, keeping the system compact: the motor speeds up slowly the reaction wheel, and when a reorientation is required, the break stops its spin.
-Since the effect of the break is unidirectional, it is possible to generate a rotation of the base only in the opposite direction of the reaction wheel angular velocity.
+Using a brake avoids the employment of a motor able to deliver higher torques \cite{gajamohan2012cubli}, 
+keeping the system compact: the motor slowly accelerates the wheel to a certain speed to store angular momentum and, when a reorientation is required, the break stops its spin.
+Since the effect of the break is unidirectional, a limitation of this approach is given by the fact that it is possible to generate a rotation of the base only in the opposite direction of the reaction wheel angular velocity making them unsuitable for applications where \textit{continuous} controllability is required.
+On the other hand, direct-drive controlled flywheels enable to create accelerations in both directions,  and  to implement continuous control laws.
+The benefits of this kind of  \gls{ocs} are multiple: 1) flywheels can correct orientation errors during the flight (e.g.
+due to disturbances or tracking inaccuracies in the angular momentum achieved at the lift-off)   % and non-idealities).
+in a continuous manner. 2) They enable to track a time-varying reference (e.g. to have the robot land with a desired angular velocity (possibly zero) and orientation).
+3) Even in presence of contacts, they can enhance the landing phase or stabilize dynamic gaits (e.g. trot), by significantly reducing trunk oscillations.  
+4) The presence of these additional joints, whose only function is to control the orientation, gives the possibility to relieve the effort of the legs.
+In more complex scenarios, like for a somersault, limbs and \gls{ocs} can operate in parallel to achieve a rotation angle larger than the one achievable only with legs (e.g., due to torque limitation).
 
 \subsection{Proposed Approach and Contribution}
-In this work, we present an \acrlong{ocs} based on two flywheels, mounted on the trunk of the quadruped Solo12 \cite{grimminger2020open}.
-In order to control roll and pitch orientations with a compact device, the axes of rotation are set to be incident, creating an angle that balances the effect in both the directions.
-Flywheels can correct orientation errors due to disturbances during the flight and inaccuracies in to the angular momentum achieved at the lift-off (e.g., given by tracking issues and non-idealities).
-They enable the robot to land with a desired angular velocity (possibly zero) and orientation.
-In addition, they can enhance the landing phase by significantly reducing oscillations.  
-The presence of these additional joints, whose only function is to control the orientation, gives the possibility to relieve the effort of the legs.
-In more complex scenarios, like in a somersault, legs and flywheels can work in parallel to achieve a rotation angle larger than the one achievable only with legs (e.g., due to torque limitation).
+In this work, we present an \gls{ocs} based on two flywheels,  mounted on the trunk of the quadruped Solo12 \cite{grimminger2020open}, see Fig. \ref{fig:solo12flywheels}. 
+The contributions of the paper are:
+\begin{itemize}
+	\item a novel design that enables to control the orientation of a legged robot during a jump in an effective way, while keeping the design simple. 
+	In particular,  the axes of rotation of the flywheels are set to be incident, enabling \textit{continuous} controllability in both the directions (roll and pitch) while keeping the  device compact.
+	\item simulations  with the quadruped Solo12 that demonstrate the effectivess of the proposed approach.
+\end{itemize}
 
 \subsection{Outline}
+The reminder of this paper is organized as follow.
+In Section \ref{sec:background} the of conservation of the total angular momentum is recalled.
+In Section \ref{sec:ocs}, this principle is used in a simplified model for deriving the lower bound for the inertia of the \gls{ocs} and it is shown how to exploit it for simultaneously controlling the robot roll and pitch orientations.
+Section \ref{sec:simulations} presents simulation in different scenarios that proof the capability of our \gls{ocs} for rejecting disturbances and tracking angular references when there is no contact with the groung, and dumping base oscillation after the touch down.
+Conclusions and possible evolution of the work are reported in Section \ref{sec:conclusions}.
 %
@@
 \section{Background}
+\label{sec:background}
+The starting point for any \gls{ocs} is Euler's equation.
+For any mechanical system, the time derivative of the angular momentum computed with the respect to a reference point $O$, fixed in an inertial frame,  equals the resultant torque with respect to the same reference point:
+\begin{equation}
+\dot{\bm{L}} = \sum_i \bm{M}_i
+\end{equation}
+When the resultant of the external moments applied to the system is zero, Euler’s equation simplifies to:
+\begin{equation}
+\dot{\bm{L}} = 0 \quad \Rightarrow \quad \bm{L}(t) = const,
+\label{eq:cons_ang_mom}
+\end{equation}
+which is known as conservation of angular momentum.
+Referring to legged robots, this condition occurs when the system is not in contact with the ground or other objects, e.g., during a fall or the flight phase of a jump.
+In this case, it is possible to change the angular velocity of the base link changing the joint positions and velocities, as a result of the \textit{non-holonomy} of the total angular momentum \cite{Wieber16}: if the angular momentum of a certain body changes, the one of the others must change to maintain the total sum constant.
+
+\section{Orientation Control System}
+\label{sec:ocs}
+In the following, we present a procedure for selecting the inertia of the flywheels, and we show how to exploit them for simultaneously controlling the robot roll and pitch orientation.
+\subsection{Inertia Selection}
+A preliminary analysis for the selection of the flywheels' inertia can be performed with the Elroy's Beanie model, depicted in Fig. \ref{fig:elroy-beanie}.
+\begin{figure}
+	\centering
+	\includegraphics[width=.7\linewidth]{figures/elroys-beanie.eps}
+	\caption{\small Schematic representation of the Elroy's Beanie model used for the preliminary analysis of the pitch motion.}
+	\label{fig:elroy-beanie}
+\end{figure} 
+This consists of two rigid bodies connected through their \gls{com} with a revolute joint.
+One of the bodies represents the robot in its nominal configuration.
+The other models the \gls{ocs}, here represented by a single wheel for the sake of simplicity.
+The aim is to examine the rotational dynamics of the system as a whole.
+To simplify the analysis, in the following we will refer only to the effects on the pitch angle of the robot, keeping in mind that the same arguments apply also to the roll.
+Let us identify the moment of inertia of the robot in the nominal configuration as $I_r$ and the moment of inertia of the two flywheels as $I_f$, both referred to the system \gls{com}.
+The angular momentum $L$ of this system can be written as
+\begin{equation}\label{eq:ang_mom_elroy}
+L(t) = \left(I_r + I_{f}\right)\dot{\theta}(t) + I_{f}\dot{\gamma}(t)
+\end{equation}
+where $\dot{\theta}$ and $\dot{\gamma}$ are respectively the robot pitch rate and the angular speed of the wheel.
+We want to drive the robot to a desired pitch rate $\dot{\theta}_{des}$ acting on the wheel speed.
+Without loss of generalities, we can assume the reaction wheel is stationary with the robot at the instant at which the reorientation maneuver starts ($\dot{\gamma}_0 = 0$), having a system angular momentum of $L_0 = \left(I_r + I_{f}\right)\dot{\theta}_0$.
+Under the condition of conservation of the angular momentum, this quantity is constant over time and it is possible to estimate the lower bound for $I_{f}$ given a desired pitch rate and the maximum velocity of the flywheels $\dot{\gamma}_{max}$:
+\begin{equation}
+I_{f} \geq I_r \dfrac{ \Delta \dot{\theta} }{\Delta \dot{\theta}+\dot{\gamma}_{max}}.
+\end{equation}
+where $\Delta \dot{\theta} =  \left\lvert \dot{\theta}_0-\dot{\theta}_{des} \right\rvert$ is the base velocity variation.
+Fig. \ref{fig:inertia_limits} reports the lower bound of $I_{f}$ given the desired base velocity variation and the maximum velocity of the actuator.
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{figures/lowerbound.eps}
+	\caption{\small Minimum inertia of the \gls{ocs} (normalized by the robot inertia) necessary to achieve a base velocity variation $\Delta \dot{\theta}$, given the actuation bound $\dot \gamma^{max}$}
+	\label{fig:inertia_limits}
+\end{figure}
+
+Once the desired inertia $I_f$ is chosen with the 2-D Elroy's Beanie model, it must be used to realize the 3-D \gls{ocs} (Fig. \ref{fig:flywheel}). 
+\begin{figure}
+	\centering
+	\includegraphics[width=0.7\linewidth]{figures/flywheel_opaque.png}
+	\caption{\small The final design of the reaction wheel, shown together with its inertia principal axes.
+Here, $h$ is the wheel thickness, $r$ is the inner radius of the wheel and $R$ is the outer one.}
+	\label{fig:flywheel}
+\end{figure}
+We decided to realize the reaction wheels in the shape of hollow cylinders.
+This shape increases the inertia by locating the mass far away from the rotation axis.
+The inertia tensor expressed in its principal axes, $\bm{I}_f = \mathrm{diag}\left\{ I_{f, xx}, \, I_{f, yy}, \, I_{f, zz}  \right\}$, depends on the cylinders inner and outer radii, $r$ and $R$, its height $h$ and the material density $\rho$:
+\begin{equation}
+\begin{split}
+I_{f, xx} = I_{f, yy} &= \frac{1}{12}\pi \rho h \left(3\left(R^4 - r^4\right) + h^2\left(R^2 - r^2\right)\right) \\
+I_{f, zz} &= \frac{1}{2}\pi \rho h \left(R^4 - r^4\right)
+\end{split}
+\end{equation}
+Notice that in the Elroy's Beanie model, there is a single body that models the complete \gls{ocs}.
+To meet the specification of controlling both the robot roll and pitch, it is possible to select one of two design options:
+\begin{enumerate}
+	\item use two couples of flywheels, each couple having the rotation axes aligned and orthogonal to the ones of the other couple \MF{metti foto con i due design uno di fianco all altro}; \ADP{Not clear why you would need 4 flywheels and not just 2.
+Also, I think it's not interesting to explain this design if then we don't use it.}
+	\item use two flywheels, with incident axes of rotation.
+\end{enumerate}
+The first case would lead to a bulky \gls{ocs}, requiring brakes to achieve relevant angular momentum.
+Thus, the second option is considered.
+The inertia introduced in the Elroy's Beanie model has to be split between the two reaction wheels: $I_f = 2 \, I_{f, zz}$.
+\ADP{Not clear why the inertia should be split because you haven't discussed yet the fact that the flywheels won't be aligned with the axes.
+Move this to a later point.}
+The parameters $R$ and $h$ can be set to have a compact \gls{ocs} and $\rho$ depends on the chosen material, that is stainless steel in our case.
+The inner radius $r$ can be adjusted to obtain the desired inertia.
+Spokes with negligible mass connect the wheel to the motor shaft.
+All the parameters are reported in Table \ref{tab:dim_fw}, together with the flywheels final inertia and mass.
+\begin{table}
+	\centering
+	\caption{\small Size and dynamic parameters of a single reaction wheel}
+	\begin{tabular}{ccc}
+		Parameter & Value & Unit \\
+		\hline
+		\hline
+		$r$ & $2.20\cdot 10^{-2}$ & $\mathrm{m}$ \\
+		$R$ & $3.00\cdot 10^{-2}$ & $\mathrm{m}$ \\
+		$h$   & $1.02\cdot 10^{-2}$ & $\mathrm{m}$ \\
+		$\rho$   & $7.860 \cdot 10^{3}$ & $\mathrm{kg/m^3}$ \\
+		$m$   & $1.02\cdot 10^{-1}$ & $\mathrm{kg}$ \\
+		$I_{xx}, I_{yy}$   & $3.64 \cdot 10^{-5}$ & $\mathrm{kg \cdot m^2}$ \\
+	$I_{zz}$   & $7.11 \cdot 10^{-5}$ & $ \mathrm{kg \cdot m^2}$ \\
+	\end{tabular}
+	\label{tab:dim_fw}
+\end{table}
+
+\subsection{Flywheels Axes of Rotation}
+
+The orientation of the axis of rotation of each flywheel influences their contribution to the total angular momentum.
+To meet the specification of controlling both the robot roll and pitch, the axes of rotation of the left and right wheel, identified in the base reference frame with the unit vectors $\hat{\bm{a}}_l$ and $\hat{\bm{a}}_r$, are set to be incident, laying on a plane parallel to the $xy-$plane of the base reference frame (see Fig.
+\ref{fig:axes}).
+To remove unnecessary complications, we design two symmetrical modules.
+We denote with $\alpha \leq \pi/2$ the non-negative incident angle between the wheels' axes of rotation and the robot lateral direction.
+The matrix
+\begin{equation}
+\bm{C} = 
+\left[ \begin{array}{cc}
+\hat{\bm{a}}_l & \hat{\bm{a}}_r
+\end{array} \right] = 
+\left[ \begin{array}{cc}
+\sin(\alpha) & -\sin(\alpha) \\
+\cos(\alpha) & \cos(\alpha) \\
+0 & 0
+\end{array} \right]
+\end{equation}
+maps the reaction wheel moments into scalar torques $\bm{u} = \left[\begin{array}{cc} \tau_{fl} & \tau_{fr} \end{array} \right]^T$ about the flywheel axes $\hat{\bm{a}}_l$ and $\hat{\bm{a}}_r$ (expressed in the base frame).
+As long as $\bm{C}$ is full column rank, it is possible to control both roll and pitch angles.
+If $\alpha = 0$, the roll results to be uncontrollable through the \gls{ocs}; otherwise, if $\alpha = \pi/2$, the pitch is uncontrollable.
+\begin{figure}
+	\centering
+	\includegraphics[width=0.7\linewidth]{figures/axes.eps}
+	\caption{\small Representation of the flywheels rotation axes, seen from the top of the robot}
+	\label{fig:axes}
+\end{figure}
+The angle is selected considering the ratio of the eigenvalues along the $x$ and $y$ directions of the ellipsoid of the robot inertia:
+\begin{equation*}
+\alpha^{*}=\tan^{-1}\left(\dfrac{I_{r, xx}}{I_{r, yy}}\right).
+\end{equation*}
+With these considerations, the angular momentum produced by the wheels, expressed in the robot base frame, is:
+\begin{equation}
+\begin{split}
+\tensor[_b]{\bm{L}}{_f} &= \tensor[_b]{\bm{L}}{_{fl}} + \tensor[_b]{\bm{L}}{_{fr}} \\
+&=I_{f, zz} \, \bm{\omega}_{fl} + I_{f, zz} \, \bm{\omega}_{fl}\\
+&=I_{f, zz}
+\left[\begin{array}{c}
+\left(\omega_{fl} - \omega_{fr}\right)\sin(\alpha) \\ 
+\left(\omega_{fl} +  \omega_{fr}\right)\cos(\alpha) \\
+0
+\end{array}\right]
+\end{split}
+\label{eq:bLf}
+\end{equation}
+in which $\bm{\omega}_{fl}= \omega_{fl} \, \hat{\bm{a}}_{l}$ and $\bm{\omega}_{fr}= \omega_{fr} \, \hat{\bm{a}}_{r}$ are the angular velocity vectors of the two wheels, and $\omega_{fl}$ and $\omega_{fr}$ are the speeds provided to each flywheel by its actuation system.
+The latter equation shows that the sum of the two angular speeds impacts on roll rotations, while their difference can be used to adjust the pitch, see Fig.
+\ref{fig:sum-diff}.
+Using the definition of $\bm{C}$, \eqref{eq:bLf} rewrites as
+\begin{equation}
+\begin{split}
+\tensor[_b]{\bm{L}}{_f} & = I_{f,zz}
+\left[
+\begin{array}{cc}
+\sin(\alpha) & \sin(\alpha) \\
+\cos(\alpha) & -\cos(\alpha) \\
+0 & 0
+\end{array}
+\right]
+\left[
+\begin{array}{cc}
+\omega_{fl} \\
+\omega_{fr}
+\end{array}
+\right] \\
+&= I_{f,zz} \bm{C} \bm{\omega}_f.
+\end{split}
+\end{equation}
+\begin{figure}
+	\centering
+	\includegraphics[width=.8\linewidth]{figures/sum_diff_mod.eps}
+	\caption{\small Having incident rotation axes, the \gls{ocs} allows to control both the robot roll and pitch.
+Notice that the roll is influenced by the sum of the speeds of the flywheels, $\omega_x = \left(\omega_{fl} +  \omega_{fr}\right)\sin(\alpha)$.
+On the contrary, the pitch depends on their difference $\omega_y = \left(\omega_{fl} -  \omega_{fr}\right)\cos(\alpha)$.}
+	\label{fig:sum-diff}
+\end{figure}
+
+\subsection{Flywheels Control Law}
+\label{subsec:control_law}
+To derive a control law based on the robot base orientation, we make use of \eqref{eq:cons_ang_mom}, expressing all the contributions to the time derivative of total angular momentum with the respect to the base reference frame:
+\begin{equation*}
+\bm{I}_f \dot{\bm{\omega}}_f + \bm{I}_b \dot{\bm{\omega}}_b + \bm{\omega}_b \times \left( \bm{I}_b \bm{\omega}_b\right) = 0
+\end{equation*}
+From this expression, we can define the moment on the base caused by the acceleration of the flywheels, that can be used as a feedback torque $\bm{\tau}_{fb}$:
+\begin{equation*}
+\begin{split}
+\bm{\tau}_{fb} &= \bm{I}_f \dot{\bm{\omega}}_f \\
+&= - \bm{I}_b \dot{\bm{\omega}}_b - \bm{\omega}_b \times \left( \bm{I}_b \bm{\omega}_b\right)
+\end{split}
+\end{equation*}
+We can use the PD action $\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}}$ to generate a reference acceleration for the base $\dot{\bm{\omega}}_b$, where $\bm{K}_p$ and $\bm{K}_d$ are diagonal and positive-definite gain matrices for the error in attitude and angular velocity.
+The error $\bm{e} \in SO(3)$ needs the algebra of the special rotational group to be calculated. To avoid singular configurations, we represent the orientation with quaternions.
+The derivative error can be calculated using $\dot{\bm{e}} = \bm{\omega}^{des}_b - \bm{\omega}_b$, in which $\bm{\omega}^{des}_b$ and $\bm{\omega}_b$ are, respectively, the desired and actual angular velocity of the base:
+\begin{equation*}
+\bm{\tau}_{fb} = - \bm{I}_b \left(\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}}\right) - \bm{\omega}_b \times \left( \bm{I}_b \bm{\omega}_b\right)
+\end{equation*}
+Projecting $\bm{\tau}_{fb}$ onto the flywheel axes with $\bm{C}^T$, we obtain the control action $\bm{u}$
+\begin{equation}
+\label{eq:ctrl_law}
+\bm{u} = \bm{C}^T \bm{\tau}_{fb}
+\end{equation}
+
+\section{Results of Simulations}
+\label{sec:simulations}
+To validate our \gls{ocs}, we present three simulations on different scenarios. We want to test the capability of the proposed approach to: 
+reject disturbance when the robot is in the flight phase of a jump, damp trunk oscillations after the touch down, 
+and work in parallel with the joints of the legs to achieve a highly dynamic motion. 
+All the simulations are performed in Gazebo.
+The robot full dynamics is modeled with Pinocchio \cite{carpentier2019pinocchio}.
+The references for the joints of the legs are computed off-line using Crocoddyl \cite{mastalli2020crocoddyl} and tracked with a PD joint controller.
+The \gls{ocs} is commanded to
+track base orientation references using the control law introduced in \ref{subsec:control_law}.\footnote{The video with all the simulations is available at web link \url{https://www.dropbox.com/scl/fo/wevvl0zvfp8gofie6x7rw/h?dl=0&rlkey=pap76cr813a0skrvrxl7qgrb9}.}
+
+%this is the main test! 
+The necessity of having an \gls{ocs} is faced up in the case of a disturbance that occurs when the robot has no contact with the ground. 
+During the flight phase of a forward jump, $0.1~\mathrm{s}$ after the lift off, an external disturbance moment is applied to the robot base, deflecting the robot orientation: this disturbance is set to $\dot{\bm{L}}_{b, dist} = \left[\begin{array}{ccc}
+0.2 & 0.8 & 0.0
+\end{array}
+\right]^T \ \mathrm{N \cdot m}$ and applies for $0.05 \ \mathrm{s}$. 
+If the flywheels are not used, Solo12 falls after the touch down. Instead, 
+using the flywheels %pairing the legs joint PD controller with \ref{eq:ctrl_law},
+it is possible to drive the robot to a safe configuration after the landing (Fig. \ref{fig:fw_jump}), without the implementation of a specific landing strategy, as the one in \cite{jeon2021real}.
+\begin{figure}
+	\centering
+	\includegraphics[width=1\linewidth]{figures/fw_jump.eps}
+	\caption{\small A purely angular disturbance on the trunk can be compensated only if the \gls{ocs} is turned on.
+If it does not, the robot is unable to restore a safe configuration after the touch down}
+	\label{fig:fw_jump}
+\end{figure}
+
+In the second test, we demonstrate the ability of the \gls{ocs} to reorient the base.
+We let the robot fall from a height of $1 \ \mathrm{m}$ with an initial pitch orientation of $30^\circ$ and zero base angular velocity.
+If we do not actuate the reaction wheels, the robot touches the ground with the same initial orientation and the trunk oscillates.
+If the \gls{ocs} drives the robot base to be horizontal when it is still in the air, the oscillations after the touch down are drastically reduced, both in the pitch angle and forward direction (Fig.
+\ref{fig:fall_plot}).
+\begin{figure}
+	\centering
+	\includegraphics[width=1\linewidth]{figures/fall_plot.eps}
+	\caption{\small The \gls{ocs}, driving to zero the robot orientation during a fall, allows for damping the base oscillations after the touch down (vertical dashed line), even without implementing a landing strategy}
+	\label{fig:fall_plot}
+\end{figure}
+
+%In a second test, we demonstrate the versatility use the presented \gls{ocs} to dump the base oscillations generated at the landing. 
+%% , even though we designed it to control the orientation during the flight phase.
+%To demonstrate this feature, we use the same joint trajectory to perform two simulations: 
+%in the first, we control the reaction wheels to track an orientation parallel to the ground, while in the other they are not actuated.
+%The joint trajectory realizes a jump with a height of $0.4 \ \mathrm{m}$, a forward distance of $0.8 \ \mathrm{m}$, and a flight phase of $0.5 \ \mathrm{s}$.
+\MF{why you detail here the jump and not before?
+is the same jump?}
+%The results point out the possibility to use this device even when the angular momentum is not conserved (i.e. when contacts are present).
+
+Finally, we demonstrate how reaction wheels can alleviate the effort of the leg joints to achieve a highly dynamic motion: a back-flip. 
+For this, we target a space application carrying out a simulation with Moon gravity ($1.62 \ \mathrm{m/s^2}$).
+In this way, it is possible to obtain high jumps with a long flight phase, without having to select more powerful actuators.  
+The leg joint trajectory computed off-line describes a purely vertical jump of $1 \ \mathrm{m}$, having a flight phase that lasts $2 \ \mathrm{s}$.
+Right after the lift-off, the reaction wheels start the reorientation task of performing a back-flip, that is a spin of $360^\circ$ on the pitch.
+For this maneuver, the ideal value of the incident angle $\alpha$ is $0^\circ$ since no roll rotation is required.
+% and it is preferred to use all the contributions of the \gls{ocs} on the pitch.
+Our hardware design allows the manual change of this value before performing the task.
+This simulation demonstrates that reaction wheels alleviate the effort to be applied on the legs, which are only used for the linear motion.
 
 \section{Conclusions}
-\label{sec:conclusion}
+\label{sec:conclusions}
+
+In this work, we have presented the design of an \gls{ocs}, composed of two flywheels, that enables to control the trunk of a legged robot platform, 
+increasing the accuracy of aerial maneuvers both during the under-actuated phases (i.e., flight phase) as well as 
+enhancing stability when in contact by damping oscillations.
+The novelty of the design, that involves the flywheels to be attached with incident rotational axes on the trunk, allows to control the orientation
+in both the roll and pitch directions while keeping the device compact. Their joint effect is maximized considering the inertial property of the mechanical structure. 
+Several simulations are reported with the quadruped Solo12 that demonstrate the effectiveness of the proposed approach to: 
+reject disturbances during the flight phase, stabilize the platform after the touch down even in absence of a specific landing strategy and 
+achieve a fast reorientation maneuver (a back-flip) in a reduced gravity environment.
+
+In future work we plan to proof our concept performing some experiment. 
+The control strategy for defining the base desired angular velocity can be improved using \gls{nmpc}, that takes in consideration the future samples of the orientation reference.
+We expect this feature allows for stabilizing the yaw to a desired value, even if it is not locally controllable, enhancing the non-holonomy property of the angular momentum, i.e.
+doing a preliminary roll and pitch maneuver.
+A control strategy that may leads to the same result is dynamic feedback linearization, widely used for flight with quadricopters.
 

### 2022-07-15 humanoids22roscia.tex
last modifications (Michele + Claudio comments) + rename file

--- 2022-07-15 mdpi22roscia.tex
@@
 For legged robots, aerial motions are the only option to overpass obstacles that cannot be circumvent with standard locomotion gaits.
-In these cases, the robot must perform a leap to either jump onto the obstacle or even fly over it.
-However, these movements represent a challenge because during the flight phase the center of mass cannot be controlled, and the robot orientation has limited controllability.
+In these cases, the robot must perform a leap to either jump onto the obstacle or fly over it.
+However, these movements represent a challenge because during the flight phase the \gls{com} cannot be controlled, and the robot orientation has limited controllability.
 This paper focuses on the latter issue and proposes an \gls{ocs} consisting of two rotating and actuated masses (flywheels or reaction wheels) to gain control authority on the robot orientation.
@@
 	\centering
-	\includegraphics[width=.7\linewidth]{figures/solo_flywheels_pic.png}
-	\caption{\small The proposed Orientation Control System for Solo12 has two flywheels with incident rotation axes. Each wheel is located in a 3D-printed shell and mounted on the trunk body.}
+	\includegraphics[width=.7\linewidth]{figures/solo_fw_pic_compressed.png}
+	\caption{\small The proposed Orientation Control System for 
+		for the $2.5 \ \mathrm{kg}$ quadruped robot Solo12 consists of two $0.1 \ \mathrm{kg}$ flywheels with incident rotation axes.
+Each wheel is located in a 3D-printed shell and mounted on the trunk body.}
 	\label{fig:solo12flywheels}
@@
 
-Quadrupedal animals, like cats, can rearrange the tail and trunk to correct the orientation during a fall~\cite{kane1969dynamical}.
+Quadrupedal animals, like cats, can rearrange the tail and trunk to correct the orientation during a fall \cite{kane1969dynamical}.
 % Tail
@@
 
-% Reaction wheel
-Reaction wheels represent an additional option for controlling the robot orientation.
+% flywheel
+flywheels represent an additional option for controlling the robot orientation.
 Changing the angular velocity of a rotating mass attached to the trunk generates a torque that can reorient the robot.
 This device has been first applied for spacecraft orientation \cite{oland2009reaction}, and later it has been only sporadically investigated in legged locomotion for controlling pitch orientation, both for bipeds \cite{Brown2016}, \cite{xiong2020sequential} and quadrupeds \cite{kolvenbach2019towards}, \cite{vasilopoulos2016quadruped}.
-Compared with tails, reaction wheels do not have position limit, and since their rotation axes pass though the \gls{com}, its angular momentum results holonomic \cite{machairas2015quadruped}.
-To get a fast response, it is necessary to have an abrupt change in the reaction wheel angular velocity (angular acceleration).
+Compared with tails, flywheels do not have position limits, and since their rotation axes pass through the \gls{com}, its angular momentum results holonomic \cite{machairas2015quadruped}.
+To get a fast response, it is necessary to have an abrupt change in the flywheels angular velocity (angular acceleration).
 Using a brake avoids the employment of a motor able to deliver higher torques \cite{gajamohan2012cubli}, 
 keeping the system compact: the motor slowly accelerates the wheel to a certain speed to store angular momentum and, when a reorientation is required, the break stops its spin.
-Since the effect of the break is unidirectional, a limitation of this approach is given by the fact that it is possible to generate a rotation of the base only in the opposite direction of the reaction wheel angular velocity making them unsuitable for applications where \textit{continuous} controllability is required.
+Since the effect of the break is unidirectional, a limitation of this approach is given by the fact that it is possible to generate a rotation of the base only in the opposite direction of the flywheel angular velocity making them unsuitable for applications where \textit{continuous} controllability is required.
 On the other hand, direct-drive controlled flywheels enable to create accelerations in both directions,  and  to implement continuous control laws.
@@
 \subsection{Proposed Approach and Contribution}
-In this work, we present an \gls{ocs} based on two flywheels,  mounted on the trunk of the quadruped Solo12 \cite{grimminger2020open}, see Fig. \ref{fig:solo12flywheels}. 
+In this work, we present a compact \gls{ocs} based on two flywheels, mounted on the trunk of the lightweight quadruped Solo12 \cite{grimminger2020open}, see Fig. \ref{fig:solo12flywheels}. 
 The contributions of the paper are:
 \begin{itemize}
-	\item a novel design that enables to control the orientation of a legged robot during a jump in an effective way, while keeping the design simple. 
+	\item the design of a novel \gls{ocs} that enables to control the orientation of a legged robot during a jump in an effective way, while keeping the design simple. 
 	In particular,  the axes of rotation of the flywheels are set to be incident, enabling \textit{continuous} controllability in both the directions (roll and pitch) while keeping the  device compact.
-	\item simulations  with the quadruped Solo12 that demonstrate the effectivess of the proposed approach.
+	\item simulations  with the quadruped Solo12 that demonstrate the effectiveness of the proposed approach.
 \end{itemize}
@@
 The reminder of this paper is organized as follow.
-In Section \ref{sec:background} the of conservation of the total angular momentum is recalled.
-In Section \ref{sec:ocs}, this principle is used in a simplified model for deriving the lower bound for the inertia of the \gls{ocs} and it is shown how to exploit it for simultaneously controlling the robot roll and pitch orientations.
-Section \ref{sec:simulations} presents simulation in different scenarios that proof the capability of our \gls{ocs} for rejecting disturbances and tracking angular references when there is no contact with the groung, and dumping base oscillation after the touch down.
-Conclusions and possible evolution of the work are reported in Section \ref{sec:conclusions}.
+In Section \ref{sec:background}, the law of conservation of the total angular momentum is recalled.
+In Section \ref{sec:ocs}, the design principles are presented together with the strategy for simultaneously control the robot roll and pitch orientations.
+Section \ref{sec:simulations} presents simulation in different scenarios that demonstrate the the capability of our \gls{ocs} to reject disturbances and to track angular references when there is no contact with the ground, and damping of base oscillation after the touch-down.
+Conclusions and a possible evolution of the work ar reported in Section \ref{sec:conclusions}.
 %
@@
 \label{sec:background}
-The starting point for any \gls{ocs} is Euler's equation.
-For any mechanical system, the time derivative of the angular momentum computed with the respect to a reference point $O$, fixed in an inertial frame,  equals the resultant torque with respect to the same reference point:
+The starting point for any \gls{ocs} is the Euler's equation.
+For any mechanical system, the time derivative of the angular momentum $\bm{L}$ computed with the respect to a reference point $O$, fixed in an inertial frame, equals the sum of the moments $\bm{M}_i$ applied to the system with respect to the same reference point:
 \begin{equation}
@@
 \end{equation}
-When the resultant of the external moments applied to the system is zero, Euler’s equation simplifies to:
+When the resultant of the external moments applied to the system is zero, the Euler's equation simplifies to:
 \begin{equation}
@@
 \label{sec:ocs}
-In the following, we present a procedure for selecting the inertia of the flywheels, and we show how to exploit them for simultaneously controlling the robot roll and pitch orientation.
+In this section, we present a procedure to select the inertia of the flywheels, and we show how to exploit them for simultaneously controlling the robot roll and pitch orientation.
 \subsection{Inertia Selection}
@@
 where $\dot{\theta}$ and $\dot{\gamma}$ are respectively the robot pitch rate and the angular speed of the wheel.
-We want to drive the robot to a desired pitch rate $\dot{\theta}_{des}$ acting on the wheel speed.
-Without loss of generalities, we can assume the reaction wheel is stationary with the robot at the instant at which the reorientation maneuver starts ($\dot{\gamma}_0 = 0$), having a system angular momentum of $L_0 = \left(I_r + I_{f}\right)\dot{\theta}_0$.
+We want to drive the robot to a desired pitch rate $\dot{\theta}_{des}$ by acting on the wheel speed.
+Without loss of generality, we can assume the flywheel is stationary with the robot at the instant at which the reorientation maneuver starts ($\dot{\gamma}_0 = 0$), having a system angular momentum of $L_0 = \left(I_r + I_{f}\right)\dot{\theta}_0$.
 Under the condition of conservation of the angular momentum, this quantity is constant over time and it is possible to estimate the lower bound for $I_{f}$ given a desired pitch rate and the maximum velocity of the flywheels $\dot{\gamma}_{max}$:
@@
 	\includegraphics[width=\linewidth]{figures/lowerbound.eps}
-	\caption{\small Minimum inertia of the \gls{ocs} (normalized by the robot inertia) necessary to achieve a base velocity variation $\Delta \dot{\theta}$, given the actuation bound $\dot \gamma^{max}$}
+	\caption{\small Minimum inertia of the \gls{ocs} (normalized by the robot inertia) necessary to achieve a base velocity variation $\Delta \dot{\theta}$, given the actuation bound $\dot \gamma^{max}$.}
 	\label{fig:inertia_limits}
 \end{figure}
-
-Once the desired inertia $I_f$ is chosen with the 2-D Elroy's Beanie model, it must be used to realize the 3-D \gls{ocs} (Fig. \ref{fig:flywheel}). 
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{figures/flywheel_opaque.png}
-	\caption{\small The final design of the reaction wheel, shown together with its inertia principal axes.
-Here, $h$ is the wheel thickness, $r$ is the inner radius of the wheel and $R$ is the outer one.}
-	\label{fig:flywheel}
-\end{figure}
-We decided to realize the reaction wheels in the shape of hollow cylinders.
-This shape increases the inertia by locating the mass far away from the rotation axis.
-The inertia tensor expressed in its principal axes, $\bm{I}_f = \mathrm{diag}\left\{ I_{f, xx}, \, I_{f, yy}, \, I_{f, zz}  \right\}$, depends on the cylinders inner and outer radii, $r$ and $R$, its height $h$ and the material density $\rho$:
-\begin{equation}
-\begin{split}
-I_{f, xx} = I_{f, yy} &= \frac{1}{12}\pi \rho h \left(3\left(R^4 - r^4\right) + h^2\left(R^2 - r^2\right)\right) \\
-I_{f, zz} &= \frac{1}{2}\pi \rho h \left(R^4 - r^4\right)
-\end{split}
-\end{equation}
-Notice that in the Elroy's Beanie model, there is a single body that models the complete \gls{ocs}.
-To meet the specification of controlling both the robot roll and pitch, it is possible to select one of two design options:
-\begin{enumerate}
-	\item use two couples of flywheels, each couple having the rotation axes aligned and orthogonal to the ones of the other couple \MF{metti foto con i due design uno di fianco all altro}; \ADP{Not clear why you would need 4 flywheels and not just 2.
-Also, I think it's not interesting to explain this design if then we don't use it.}
-	\item use two flywheels, with incident axes of rotation.
-\end{enumerate}
-The first case would lead to a bulky \gls{ocs}, requiring brakes to achieve relevant angular momentum.
-Thus, the second option is considered.
-The inertia introduced in the Elroy's Beanie model has to be split between the two reaction wheels: $I_f = 2 \, I_{f, zz}$.
-\ADP{Not clear why the inertia should be split because you haven't discussed yet the fact that the flywheels won't be aligned with the axes.
-Move this to a later point.}
-The parameters $R$ and $h$ can be set to have a compact \gls{ocs} and $\rho$ depends on the chosen material, that is stainless steel in our case.
-The inner radius $r$ can be adjusted to obtain the desired inertia.
-Spokes with negligible mass connect the wheel to the motor shaft.
-All the parameters are reported in Table \ref{tab:dim_fw}, together with the flywheels final inertia and mass.
-\begin{table}
-	\centering
-	\caption{\small Size and dynamic parameters of a single reaction wheel}
-	\begin{tabular}{ccc}
-		Parameter & Value & Unit \\
-		\hline
-		\hline
-		$r$ & $2.20\cdot 10^{-2}$ & $\mathrm{m}$ \\
-		$R$ & $3.00\cdot 10^{-2}$ & $\mathrm{m}$ \\
-		$h$   & $1.02\cdot 10^{-2}$ & $\mathrm{m}$ \\
-		$\rho$   & $7.860 \cdot 10^{3}$ & $\mathrm{kg/m^3}$ \\
-		$m$   & $1.02\cdot 10^{-1}$ & $\mathrm{kg}$ \\
-		$I_{xx}, I_{yy}$   & $3.64 \cdot 10^{-5}$ & $\mathrm{kg \cdot m^2}$ \\
-	$I_{zz}$   & $7.11 \cdot 10^{-5}$ & $ \mathrm{kg \cdot m^2}$ \\
-	\end{tabular}
-	\label{tab:dim_fw}
-\end{table}
 
@@
 \ref{fig:axes}).
-To remove unnecessary complications, we design two symmetrical modules.
+To remove unnecessary complications, we designed two identical modules, to be mounted on the trunk of Solo12, symmetrically with the respect of the robot sagittal plane.
 We denote with $\alpha \leq \pi/2$ the non-negative incident angle between the wheels' axes of rotation and the robot lateral direction.
@@
 \end{equation}
-maps the reaction wheel moments into scalar torques $\bm{u} = \left[\begin{array}{cc} \tau_{fl} & \tau_{fr} \end{array} \right]^T$ about the flywheel axes $\hat{\bm{a}}_l$ and $\hat{\bm{a}}_r$ (expressed in the base frame).
+maps the flywheel moments into scalar torques $\bm{u} = \left[\begin{array}{cc} \tau_{fl} & \tau_{fr} \end{array} \right]^T$ about the flywheel axes $\hat{\bm{a}}_l$ and $\hat{\bm{a}}_r$ (expressed in the base frame).
 As long as $\bm{C}$ is full column rank, it is possible to control both roll and pitch angles.
@@
 	\includegraphics[width=0.7\linewidth]{figures/axes.eps}
-	\caption{\small Representation of the flywheels rotation axes, seen from the top of the robot}
+	\caption{\small Representation of the flywheels rotation axes, seen from the top of the robot.}
 	\label{fig:axes}
 \end{figure}
-The angle is selected considering the ratio of the eigenvalues along the $x$ and $y$ directions of the ellipsoid of the robot inertia:
+The angle is selected considering the ratio of the eigenvalues along the $x$ and $y$ directions of the ellipsoid of the robot inertia(Fig. \ref{fig:ellipsoid}):
 \begin{equation*}
@@
 \end{equation*}
-With these considerations, the angular momentum produced by the wheels, expressed in the robot base frame, is:
+\begin{figure}
+	\centering
+	\includegraphics[width=1\linewidth]{figures/inertiaEllipsoid.eps}
+	\caption{\small The inertia tensor can be seen as an ellipsoid. Its principal axes in the direction of the eigenvectors of the tensor and their length depends on the eigenvalues.}
+	\label{fig:ellipsoid}
+\end{figure}
+In the case of Solo12, this angle is about $40^{\circ}$. With these considerations, the angular momentum produced by the wheels, expressed in the robot base frame, is:
 \begin{equation}
@@
 \tensor[_b]{\bm{L}}{_f} &= \tensor[_b]{\bm{L}}{_{fl}} + \tensor[_b]{\bm{L}}{_{fr}} \\
-&=I_{f, zz} \, \bm{\omega}_{fl} + I_{f, zz} \, \bm{\omega}_{fl}\\
+&=I_{f, zz} \, \bm{\omega}_{fl} + I_{f, zz} \, \bm{\omega}_{fr}\\
 &=I_{f, zz}
@@
 \end{equation}
-in which $\bm{\omega}_{fl}= \omega_{fl} \, \hat{\bm{a}}_{l}$ and $\bm{\omega}_{fr}= \omega_{fr} \, \hat{\bm{a}}_{r}$ are the angular velocity vectors of the two wheels, and $\omega_{fl}$ and $\omega_{fr}$ are the speeds provided to each flywheel by its actuation system.
-The latter equation shows that the sum of the two angular speeds impacts on roll rotations, while their difference can be used to adjust the pitch, see Fig.
+in which $\bm{\omega}_{fl}= \omega_{fl} \, \hat{\bm{a}}_{l}$ and $\bm{\omega}_{fr}= \omega_{fr} \, \hat{\bm{a}}_{r}$ are the angular velocity vectors of the two wheels, and $\omega_{fl}$ and $\omega_{fr}$ are the angular speeds provided to each flywheel by its actuation system.
+The latter equation shows that the difference of the two angular speeds impacts on roll rotations, while their difference can be used to adjust the pitch, see Fig.
 \ref{fig:sum-diff}.
@@
 \begin{array}{cc}
-\sin(\alpha) & \sin(\alpha) \\
-\cos(\alpha) & -\cos(\alpha) \\
+\sin(\alpha) & -\sin(\alpha) \\
+\cos(\alpha) & \cos(\alpha) \\
 0 & 0
@@
 	\caption{\small Having incident rotation axes, the \gls{ocs} allows to control both the robot roll and pitch.
-Notice that the roll is influenced by the sum of the speeds of the flywheels, $\omega_x = \left(\omega_{fl} +  \omega_{fr}\right)\sin(\alpha)$.
-On the contrary, the pitch depends on their difference $\omega_y = \left(\omega_{fl} -  \omega_{fr}\right)\cos(\alpha)$.}
+Notice that the roll is influenced by the difference of the angular speeds of the flywheels, $\omega_x = \left(\omega_{fl} - \omega_{fr}\right)\sin(\alpha)$.
+On the contrary, the pitch depends on the sum of the angular speeds $\omega_y = \left(\omega_{fl} -  \omega_{fr}\right)\cos(\alpha)$.}
 	\label{fig:sum-diff}
 \end{figure}
+
+\subsection{Inertia Selection}
+Once the desired inertia $I_f$ is chosen, according to the 2-D Elroy's Beanie model, it is used to realize the 3-D \gls{ocs} (Fig. \ref{fig:flywheel}). 
+\begin{figure}
+	\centering
+	\includegraphics[width=0.7\linewidth]{figures/flywheel_opaque.png}
+	\caption{\small The final design of the flywheel, shown together with its inertia principal axes. Here, $h$ is the wheel thickness, $r$ is the inner radius of the wheel and $R$ is the outer one.}
+	\label{fig:flywheel}
+\end{figure}
+We decided to design the flywheels in the shape of hollow cylinders.
+This shape increases the inertia by locating the mass far away from the rotation axis.
+The inertia tensor expressed in its principal axes, $\bm{I}_f = \mathrm{diag}\left\{ I_{f, xx}, \, I_{f, yy}, \, I_{f, zz}  \right\}$, depends on the cylinders inner and outer radii, $r$ and $R$, its height $h$ and the material density $\rho$:
+\begin{equation}
+\begin{split}
+I_{f, xx} = I_{f, yy} &= \frac{1}{12}\pi \rho h \left(3\left(R^4 - r^4\right) + h^2\left(R^2 - r^2\right)\right) \\
+I_{f, zz} &= \frac{1}{2}\pi \rho h \left(R^4 - r^4\right)
+\end{split}
+\end{equation}
+Notice that in the Elroy's Beanie model, there is a single body that models the complete \gls{ocs}.
+The inertia introduced in the Elroy's Beanie model has to be split between the two flywheels: $I_{f, zz} = I_f/\left(2\cos(\alpha)\right)$.
+The parameters $R$ and $h$ can be set to have a compact \gls{ocs} and $\rho$ depends on the chosen material, that is stainless steel in our case.
+The inner radius $r$ can be adjusted to obtain the desired inertia.
+Spokes with negligible mass connect the wheel to the motor shaft.
+All the parameters are reported in Table \ref{tab:dim_fw}, together with the selected inertia and mass of a single flywheel.
+\begin{table}
+	\centering
+	\caption{\small Size and dynamic parameters of a single flywheel}
+	\begin{tabular}{ccc}
+		Parameter & Value & Unit \\
+		\hline
+		\hline
+		$r$ & $2.20\cdot 10^{-2}$ & $\mathrm{m}$ \\
+		$R$ & $3.00\cdot 10^{-2}$ & $\mathrm{m}$ \\
+		$h$   & $1.02\cdot 10^{-2}$ & $\mathrm{m}$ \\
+		$\rho$   & $7.860 \cdot 10^{3}$ & $\mathrm{kg/m^3}$ \\
+		$m$   & $1.02\cdot 10^{-1}$ & $\mathrm{kg}$ \\
+		$I_{xx}, I_{yy}$   & $3.64 \cdot 10^{-5}$ & $\mathrm{kg \cdot m^2}$ \\
+		$I_{zz}$   & $7.11 \cdot 10^{-5}$ & $ \mathrm{kg \cdot m^2}$ \\
+	\end{tabular}
+	\label{tab:dim_fw}
+\end{table}
 
@@
 \begin{equation*}
-\bm{I}_f \dot{\bm{\omega}}_f + \bm{I}_b \dot{\bm{\omega}}_b + \bm{\omega}_b \times \left( \bm{I}_b \bm{\omega}_b\right) = 0
+\tensor[_b]{\bm{I}}{_f} \, \tensor[_b]{\dot{\bm{\omega}}}{_f} + \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\dot{\bm{\omega}}}{_r} + \tensor[_b]{\bm{\omega}}{_r} \times \left( \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\bm{\omega}}{_r}\right) = 0
 \end{equation*}
@@
 \bm{\tau}_{fb} &= \bm{I}_f \dot{\bm{\omega}}_f \\
-&= - \bm{I}_b \dot{\bm{\omega}}_b - \bm{\omega}_b \times \left( \bm{I}_b \bm{\omega}_b\right)
+&= - \tensor[_b]{\bm{I}}{_r} \dot{\bm{\omega}}_b - \tensor[_b]{\bm{\omega}}{_r} \times \left( \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\bm{\omega}}{_r}\right)
 \end{split}
 \end{equation*}
-We can use the PD action $\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}}$ to generate a reference acceleration for the base $\dot{\bm{\omega}}_b$, where $\bm{K}_p$ and $\bm{K}_d$ are diagonal and positive-definite gain matrices for the error in attitude and angular velocity.
-The error $\bm{e} \in SO(3)$ needs the algebra of the special rotational group to be calculated. To avoid singular configurations, we represent the orientation with quaternions.
-The derivative error can be calculated using $\dot{\bm{e}} = \bm{\omega}^{des}_b - \bm{\omega}_b$, in which $\bm{\omega}^{des}_b$ and $\bm{\omega}_b$ are, respectively, the desired and actual angular velocity of the base:
+We can use the PD action $\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}_r}$, where $\bm{K}_p$ and $\bm{K}_d$ are diagonal and positive-definite gain matrices for the error in attitude and angular velocity.
+The orientation error $\bm{e} \in SO(3)$ needs the algebra of the special rotational group to be computed. To avoid singular configurations, we represent the orientation with quaternions.
+The derivative error can be computed using $\dot{\bm{e}} = \tensor[_b]{\bm{\omega}}{_r^{des}} - \tensor[_b]{\bm{\omega}}{_r}$, in which $\tensor[_b]{\bm{\omega}}{_r^{des}}$ and $\tensor[_b]{\bm{\omega}}{_r}$ are, respectively, the desired and actual angular velocity of the base:
 \begin{equation*}
-\bm{\tau}_{fb} = - \bm{I}_b \left(\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}}\right) - \bm{\omega}_b \times \left( \bm{I}_b \bm{\omega}_b\right)
+\bm{\tau}_{fb} = - \tensor[_b]{\bm{I}}{_r} \left(\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}}\right) - \tensor[_b]{\bm{\omega}}{_r} \times \left( \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\bm{\omega}}{_r}\right)
 \end{equation*}
-Projecting $\bm{\tau}_{fb}$ onto the flywheel axes with $\bm{C}^T$, we obtain the control action $\bm{u}$
+Projecting the moment $\bm{\tau}_{fb}$ onto the flywheel axes with $\bm{C}^T$, we obtain the control action $\bm{u}$
 \begin{equation}
@@
 To validate our \gls{ocs}, we present three simulations on different scenarios. We want to test the capability of the proposed approach to: 
-reject disturbance when the robot is in the flight phase of a jump, damp trunk oscillations after the touch down, 
+reject a disturbance when the robot is in the flight phase of a jump, damp trunk oscillations after the touch-down, 
 and work in parallel with the joints of the legs to achieve a highly dynamic motion. 
@@
 The \gls{ocs} is commanded to
-track base orientation references using the control law introduced in \ref{subsec:control_law}.\footnote{The video with all the simulations is available at web link \url{https://www.dropbox.com/scl/fo/wevvl0zvfp8gofie6x7rw/h?dl=0&rlkey=pap76cr813a0skrvrxl7qgrb9}.}
+track base orientation references using the control law introduced in Subsection \ref{subsec:control_law}.\footnote{The video with all the simulations is available at web link \url{https://www.dropbox.com/scl/fo/wevvl0zvfp8gofie6x7rw/h?dl=0&rlkey=pap76cr813a0skrvrxl7qgrb9}.}
 
 %this is the main test! 
-The necessity of having an \gls{ocs} is faced up in the case of a disturbance that occurs when the robot has no contact with the ground. 
-During the flight phase of a forward jump, $0.1~\mathrm{s}$ after the lift off, an external disturbance moment is applied to the robot base, deflecting the robot orientation: this disturbance is set to $\dot{\bm{L}}_{b, dist} = \left[\begin{array}{ccc}
+The necessity of having an \gls{ocs} is revealed with the first test: disturbance that occurs when the robot has no contact with the ground. 
+During the flight phase of a forward jump, $0.1~\mathrm{s}$ after the lift-off, an external disturbance moment is applied to the robot base, deflecting the robot orientation: this disturbance is set to $\bm{\tau}_{dist} = \left[\begin{array}{ccc}
 0.2 & 0.8 & 0.0
 \end{array}
-\right]^T \ \mathrm{N \cdot m}$ and applies for $0.05 \ \mathrm{s}$. 
-If the flywheels are not used, Solo12 falls after the touch down. Instead, 
+\right]^T \ \mathrm{N \cdot m}$ and applied for $0.05 \ \mathrm{s}$. 
+If the flywheels are not used, Solo12 falls after the touch-down. Instead, 
 using the flywheels %pairing the legs joint PD controller with \ref{eq:ctrl_law},
-it is possible to drive the robot to a safe configuration after the landing (Fig. \ref{fig:fw_jump}), without the implementation of a specific landing strategy, as the one in \cite{jeon2021real}.
+it is possible to drive the robot to a safe configuration after the landing (Fig. \ref{fig:fw_jump}), without need to implement of a specific landing strategy, as the one in \cite{jeon2021real}.
 \begin{figure}
@@
 	\includegraphics[width=1\linewidth]{figures/fw_jump.eps}
-	\caption{\small A purely angular disturbance on the trunk can be compensated only if the \gls{ocs} is turned on.
-If it does not, the robot is unable to restore a safe configuration after the touch down}
+	\caption{\small First Test: simulation results showing the $\mathrm{CoM}_x$ and pitch trajectory vs.
+time plots.
+A disturbance moment $\bm{\tau}_{dist}$ on the trunk can be compensated only if the \gls{ocs} is enabled.
+If it is disabled, the robot is unable to restore a safe configuration after the touch-down and eventually falls.}
 	\label{fig:fw_jump}
@@
 We let the robot fall from a height of $1 \ \mathrm{m}$ with an initial pitch orientation of $30^\circ$ and zero base angular velocity.
-If we do not actuate the reaction wheels, the robot touches the ground with the same initial orientation and the trunk oscillates.
-If the \gls{ocs} drives the robot base to be horizontal when it is still in the air, the oscillations after the touch down are drastically reduced, both in the pitch angle and forward direction (Fig.
+If we do not actuate the flywheels, the robot touches the ground with the same initial orientation and the trunk oscillates.
+If the \gls{ocs}is enabled, it drives the robot base to be horizontal when it is still in the air and the oscillations after the touch-down are drastically reduced, both in the pitch angle and forward direction (Fig.
 \ref{fig:fall_plot}).
@@
 	\includegraphics[width=1\linewidth]{figures/fall_plot.eps}
-	\caption{\small The \gls{ocs}, driving to zero the robot orientation during a fall, allows for damping the base oscillations after the touch down (vertical dashed line), even without implementing a landing strategy}
+	\caption{\small Second Test: simulation results showing the $\mathrm{CoM}_x$ and pitch trajectory vs.
+time plots.
+The \gls{ocs} drives the robot orientation during a fall to be horizontal.
+This allows to damp for damping the base oscillations after the touch-down (vertical dashed line), even without implementing a landing strategy}
 	\label{fig:fall_plot}
@@
 %To demonstrate this feature, we use the same joint trajectory to perform two simulations: 
-%in the first, we control the reaction wheels to track an orientation parallel to the ground, while in the other they are not actuated.
+%in the first, we control the flywheels to track an orientation parallel to the ground, while in the other they are not actuated.
 %The joint trajectory realizes a jump with a height of $0.4 \ \mathrm{m}$, a forward distance of $0.8 \ \mathrm{m}$, and a flight phase of $0.5 \ \mathrm{s}$.
@@
 
-Finally, we demonstrate how reaction wheels can alleviate the effort of the leg joints to achieve a highly dynamic motion: a back-flip. 
+Finally, we demonstrate how flywheels can alleviate the effort of the leg joints to achieve a highly dynamic motion: a back-flip. 
 For this, we target a space application carrying out a simulation with Moon gravity ($1.62 \ \mathrm{m/s^2}$).
@@
 The leg joint trajectory computed off-line describes a purely vertical jump of $1 \ \mathrm{m}$, having a flight phase that lasts $2 \ \mathrm{s}$.
-Right after the lift-off, the reaction wheels start the reorientation task of performing a back-flip, that is a spin of $360^\circ$ on the pitch.
-For this maneuver, the ideal value of the incident angle $\alpha$ is $0^\circ$ since no roll rotation is required.
+Right after the lift-off, the flywheels start the reorientation task of performing a back-flip, that is a spin of $360^\circ$ on the pitch.
+For this maneuver, the value of the incident angle $\alpha$ is set to $0^\circ$ since no roll rotation is required.
 % and it is preferred to use all the contributions of the \gls{ocs} on the pitch.
 Our hardware design allows the manual change of this value before performing the task.
-This simulation demonstrates that reaction wheels alleviate the effort to be applied on the legs, which are only used for the linear motion.
+This simulation demonstrates that the \gls{ocs} alleviate the effort to be applied on the legs: indeed, that it would not have been possible to do a back-flip without the flywheels because the legs are used only to achieve the linear motion.
 
@@
 
-In this work, we have presented the design of an \gls{ocs}, composed of two flywheels, that enables to control the trunk of a legged robot platform, 
-increasing the accuracy of aerial maneuvers both during the under-actuated phases (i.e., flight phase) as well as 
+In this work, we have presented the design of a novel \gls{ocs}, composed of two flywheels, that enables to control the trunk of a legged robot platform, 
+increasing the accuracy of aerial maneuvers during the under-actuated phases (i.e., flight phase) as well as 
 enhancing stability when in contact by damping oscillations.
 The novelty of the design, that involves the flywheels to be attached with incident rotational axes on the trunk, allows to control the orientation
-in both the roll and pitch directions while keeping the device compact. Their joint effect is maximized considering the inertial property of the mechanical structure. 
+in both the roll and pitch directions while keeping the device compact. Their effect is optimized considering the inertial property of the mechanical structure. 
 Several simulations are reported with the quadruped Solo12 that demonstrate the effectiveness of the proposed approach to: 
-reject disturbances during the flight phase, stabilize the platform after the touch down even in absence of a specific landing strategy and 
+reject disturbances during the flight phase, stabilize the platform after the touch-down even in absence of a specific landing strategy and 
 achieve a fast reorientation maneuver (a back-flip) in a reduced gravity environment.
 
-In future work we plan to proof our concept performing some experiment. 
-The control strategy for defining the base desired angular velocity can be improved using \gls{nmpc}, that takes in consideration the future samples of the orientation reference.
-We expect this feature allows for stabilizing the yaw to a desired value, even if it is not locally controllable, enhancing the non-holonomy property of the angular momentum, i.e.
+In future works, we plan to demonstrate our concept performing experiment with the real platform.
+The control strategy for defining the base desired angular velocity can be improved using \gls{nmpc}, that takes into consideration the future samples of the orientation reference.
+We expect this feature will allows the stabilization of the yaw to a desired value, that if it is not locally controllable, enhancing the non-holonomy property of the angular momentum, i.e.
 doing a preliminary roll and pitch maneuver.
-A control strategy that may leads to the same result is dynamic feedback linearization, widely used for flight with quadricopters.
+A control strategy that may lead to the same result is dynamic feedback linearization, widely used for flight with quadricopters.
 
@@
 \section*{Acknowledgements}	
+All the authors want to thank Dr. Roy Featherstone for interesting tips that started this work. 
+\newline
+The publication was created with the co-financing of the European Union FSE-REACT-EU, PON Research and Innovation 2014-2020 DM1062 / 2021.
 

### 2022-10-27 humanoids22roscia.tex
bibliography backend and starting comments of reviewers

@@
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper
-
-%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper
+\documentclass[letterpaper, 10 pt, conference]{ieeeconf}      % Use this line for a4 paper
 
@@
 
-%\documentclass[journal,letterpaper]{IEEEtran}
+% these commands are needed because we use ieeeconf with natbib
+\let\labelindent\relax
+
+\makeatletter
+\let\NAT@parse\undefined
+\makeatother
+
 % include the list of acronyms, math commands and new commands used in this paper
@@
 \subsection{Outline}
-The reminder of this paper is organized as follow.
+The remainder of this paper is organized as follow.
 In Section \ref{sec:background}, the law of conservation of the total angular momentum is recalled.
@@
 
+\FR{In Solo12, even if the each limb accounts for about $13.4\%$ of the total mass, changes of the total robot inertia come with the motion of only the upperleg (thigh) and the lowerleg (calf), since the displacement of distribution of the hip mass with the system \acrshort{com} is approximately constant.
+In view of this, each limb accounts for only $7.5\%$ of the total mass in varying the the total robot inertia.}
+
 \section{Orientation Control System}
@@
 In this section, we present a procedure to select the inertia of the flywheels, and we show how to exploit them for simultaneously controlling the robot roll and pitch orientation.
-\subsection{Inertia Selection}
+\subsection{Bounds on the Inertia}
 A preliminary analysis for the selection of the flywheels' inertia can be performed with the Elroy's Beanie model, depicted in Fig. \ref{fig:elroy-beanie}.
@@
 If we do not actuate the flywheels, the robot touches the ground with the same initial orientation and the trunk oscillates.
-If the \gls{ocs}is enabled, it drives the robot base to be horizontal when it is still in the air and the oscillations after the touch-down are drastically reduced, both in the pitch angle and forward direction (Fig.
+If the \gls{ocs} is enabled, it drives the robot base to be horizontal when it is still in the air and the oscillations after the touch-down are drastically reduced, both in the pitch angle and forward direction (Fig.
 \ref{fig:fall_plot}).
@@
 
-\printbibliography
+\bibliography{references/bibliography}
+%\printbibliography
 \end{document}

### 2022-11-16 humanoids22roscia.tex
typos

@@
 
-Quadrupedal animals, like cats, can rearrange the tail and trunk to correct the orientation during a fall \cite{kane1969dynamical}.
+Quadrupedal animals, like cats, can rearrange tail and trunk to correct the orientation during a fall \cite{kane1969dynamical}.
 % Tail
 Many works in the field of robotics used an additional link as a tail, like in \cite{chu2019null} and \cite{wenger2016frontal}.
-This link rotates around an axis that does not pass through the robot \gls{com}: the distances between the axis of rotation and the \gls{com} of both trunk and tail result in a large effect on the total angular momentum even with a small tail mass.
+This link rotates around an axis that does not pass through the robot \gls{com}: the distances between the axis of rotation and the \gls{com} of both trunk and tail result in a large effect on the total inertia even with a small tail mass.
 However, the placement of the additional link makes the resulting robot asymmetric.
@@
 \subsection{Inertia Selection}
-Once the desired inertia $I_f$ is chosen, according to the 2-D Elroy's Beanie model, it is used to realize the 3-D \gls{ocs} (Fig. \ref{fig:flywheel}). 
+Once the desired inertia $I_f$ is chosen, according to the 2D Elroy's Beanie model, it is used to realize the 3D \gls{ocs} (Fig. \ref{fig:flywheel}). 
 \begin{figure}

### 2022-11-16 mdpi2022roscia.tex
switch to mdpi format

--- 2022-11-16 humanoids22roscia.tex
@@
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[letterpaper, 10 pt, conference]{ieeeconf}      % Use this line for a4 paper
+%  LaTeX support: latex@mdpi.com 
+%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.
 
-\IEEEoverridecommandlockouts                              % This command is only needed if 
-% you want to use the \thanks command
+%=================================================================
+\documentclass[sensors,article,submit,pdftex,moreauthors]{Definitions/mdpi} 
+% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept".
+The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}.
+This is especially recommended for submission to arXiv, where line numbers should be removed before posting.
+For preprints.org, the editorial staff will make this change immediately prior to posting.
 
-\overrideIEEEmargins                                      % Needed to meet printer requirements.
+%--------------------
+% Class Options:
+%--------------------
+%----------
+% journal
+%----------
+% Choose between the following MDPI journals:
+% acoustics, actuators, addictions, admsci, adolescents, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, alloys, analytica, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, arts, asc, asi, astronomy, atmosphere, atoms, audiolres, automation, axioms, bacteria, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomass, biomechanics, biomed, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, blsf, brainsci, breath, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, chips, cimb, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coasts, coatings, colloids, colorants, commodities, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, covid, crops, cryptography, crystals, csmf, ctn, curroncol, currophthalmol, cyber, dairy, data, dentistry, dermato, dermatopathology, designs, diabetology, diagnostics, dietetics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, ent, entomology, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fintech, fire, fishes, fluids, foods, forecasting, forensicsci, forests, foundations, fractalfract, fuels, futureinternet, futureparasites, futurepharmacol, futurephys, futuretransp, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, hazardousmatters, healthcare, hearts, hemato, heritage, highthroughput, histories, horticulturae, humanities, humans, hydrobiology, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijtm, ijtpp, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jal, jcdd, jcm, jcp, jcs, jdb, jeta, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jzbg, kidney, kidneydial, knowledge, land, languages, laws, life, liquids, literature, livers, logics, logistics, lubricants, lymphatics, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, merits, metabolites, metals, meteorology, methane, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, microplastics, minerals, mining, modelling, molbank, molecules, mps, msf, mti, muscles, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutraceuticals, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, organoids, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacoepidemiology, pharmacy, philosophies, photochem, photonics, phycology, physchem, physics, physiologia, plants, plasma, pollutants, polymers, polysaccharides, poultry, powders, preprints, proceedings, processes, prosthesis, proteomes, psf, psych, psychiatryint, psychoactives, publications, quantumrep, quaternary, qubs, radiation, reactions, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, rheumato, risks, robotics, ruminants, safety, sci, scipharm, seeds, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, software, soilsystems, solar, solids, sports, standards, stats, stresses, surfaces, surgeries, suschem, sustainability, symmetry, synbio, systems, taxonomy, technologies, telecom, test, textiles, thalassrep, thermo, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis
 
-% these commands are needed because we use ieeeconf with natbib
-\let\labelindent\relax
+%---------
+% article
+%---------
+% The default type of manuscript is "article", but can be replaced by: 
+% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
+% supfile = supplementary materials
 
+%----------
+% submit
+%----------
+% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
+This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information.
+Also, line numbering will be removed.
+Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
+
+%------------------
+% moreauthors
+%------------------
+% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
+
+%---------
+% pdftex
+%---------
+% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
+
+%=================================================================
+% MDPI internal commands
+\firstpage{1} 
 \makeatletter 
-\let\NAT@parse\undefined
+\setcounter{page}{\@firstpage} 
 \makeatother
+\pubvolume{1}
+\issuenum{1}
+\articlenumber{0}
+\pubyear{2022}
+\copyrightyear{2022}
+%\externaleditor{Academic Editor: Firstname Lastname}
+\datereceived{} 
+%\daterevised{} % Only for the journal Acoustics
+\dateaccepted{} 
+\datepublished{} 
+%\datecorrected{} % Corrected papers include a "Corrected: XXX" date in the original paper.
+%\dateretracted{} % Corrected papers include a "Retracted: XXX" date in the original paper.
+\hreflink{https://doi.org/} % If needed use \linebreak
+%\doinum{}
+%------------------------------------------------------------------
+% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
+%\pdfoutput=1
 
-% include the list of acronyms, math commands and new commands used in this paper
-\input{includes/used_packages.tex}
+%=================================================================
+% Add packages and commands here.
+The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
+\usepackage{glossaries}
+\usepackage{bm}
+\usepackage{tensor}
+%\input{includes/used_packages.tex}
 \input{includes/acronyms.tex}
-\input{includes/math_commands.tex}
-\input{includes/newcommands.tex}
-\hypersetup{draft}
-\makeglossaries
+%\input{includes/math_commands.tex}
+%\input{includes/newcommands.tex}
 
-\title{Orientation Control System: Enhancing Aerial Maneuvers \\ for Quadruped Robots}
-\author{
-	Francesco Roscia$^{1}$, Andrea Cumerlotti$^{1, \, 2}$, Andrea Del Prete$^{2}$, Claudio Semini$^{1}$, and Michele Focchi$^{1, \, 3}$
-	\thanks{$^1$ Dynamic Legged Systems (DLS) lab, Istituto Italiano di Tecnologia (IIT), Genoa, Italy.
-		$^2$ Industrial Engineering Department (DII), University of Trento, Trento, Italy.
-		$^3$ Department of Information Engineering and Computer Science (DISI), University of Trento, Trento, Italy.
-}}
+\newcommand{\MF}[1]{\textcolor{red}{\textbf{mfocchi}: #1}}
+\newcommand{\CS}[1]{\textcolor{violet}{\textbf{csemini}: #1}}
+\newcommand{\FR}[1]{\textcolor{teal}{\textbf{froscia}: #1}}
+\newcommand{\ADP}[1]{\textcolor{blue}{\textbf{adelprete}: #1}}
+%=================================================================
+%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
+%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
 
-\begin{document}
-\maketitle
-\thispagestyle{empty}
-\pagestyle{empty}
+%=================================================================
+% Full title of the paper (Capitalized)
+\Title{Orientation Control System: Enhancing Aerial Maneuvers for Quadruped Robots}
 
-\begin{abstract}%150-250 word abstract
-For legged robots, aerial motions are the only option to overpass obstacles that cannot be circumvent with standard locomotion gaits.
+% MDPI internal command: Title for citation in the left column
+\TitleCitation{Orientation Control System: Enhancing Aerial Maneuvers for Quadruped Robots}
+
+% Author Orchid ID: enter ID or remove command
+%\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
+%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
+
+% Authors, for the paper (add full first names)
+\Author{Francesco Roscia $^{1, ^*}$, Andrea Cumerlotti $^{?}$, Andrea Del Prete $^{2}$, Claudio Semini $^{1}$ and Michele Focchi$^{1, \, 3}$}
+
+%\longauthorlist{yes}
+
+% MDPI internal command: Authors, for metadata in PDF
+\AuthorNames{Francesco Roscia, Andrea Cumerlotti, Andrea Del Prete, Claudio Semini and Michele Focchi}
+
+% MDPI internal command: Authors, for citation in the left column
+\AuthorCitation{Roscia, F.; Cumerlotti, A.; Del Prete, A., Semini, C.; Focchi, M.}
+% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
+
+% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
+\address{%
+$^{1}$ \quad Dynamic Legged Systems (DLS) lab, Istituto Italiano di Tecnologia (IIT), Genoa, Italy\\
+$^{2}$ \quad Industrial Engineering Department (DII), University of Trento, Trento, Italy\\
+$^{3}$ \quad Department of Information Engineering and Computer Science (DISI), University of Trento, Trento, Italy}
+
+% Contact information of the corresponding author
+\corres{Correspondence: francesco.roscia@iit.it}
+
+% Current address and/or shared authorship
+%\firstnote{Current address: Affiliation 3.} 
+%\secondnote{These authors contributed equally to this work.}
+% The commands \thirdnote{} till \eighthnote{} are available for further notes
+
+%\simplesumm{} % Simple summary
+
+%\conference{} % An extended version of a conference paper
+
+% Abstract (Do not insert blank lines, i.e. \\) 
+\abstract{For legged robots, aerial motions are the only option to overpass obstacles that cannot be circumvent with standard locomotion gaits.
 In these cases, the robot must perform a leap to either jump onto the obstacle or fly over it.
@@
 The axes of rotation of the flywheels are designed to be incident, leading to a compact orientation control system that is capable of controlling both roll and pitch angles, considering the different moment of inertia in the two directions.
-We tested the concept with simulations on the robot Solo12.
-\end{abstract}
+We tested the concept with simulations on the robot Solo12.}
 
-%\begin{IEEEkeywords}
+\makeglossaries
+% Keywords
+\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}
 
-%\end{IEEEkeywords}
+% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
+%\PACS{J0101}
+%\MSC{}
+%\JEL{}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Diversity
+%\LSID{\url{http://}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Applied Sciences
+%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Data
+%\dataset{DOI number or link to the deposited data set if the data set is published separately.
+If the data set shall be published as a supplement to this paper, this field will be filled by the journal editors.
+In this case, please submit the data set as a supplement.}
+%\datasetlicense{License under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Toxins
+%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Encyclopedia
+%\encyclopediadef{For entry manuscripts only: please provide a brief overview of the entry title instead of an abstract.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Advances in Respiratory Medicine
+%\addhighlights{yes}
+%\renewcommand{\addhighlights}{%
+
+%\noindent This is an obligatory section in “Advances in Respiratory Medicine”, whose goal is to increase the discoverability and readability of the article via search engines and other scholars.
+Highlights should not be a copy of the abstract, but a simple text allowing the reader to quickly and simplified find out what the article is about and what can be cited from it.
+Each of these parts should be devoted up to 2~bullet points.\vspace{3pt}\\
+%\textbf{What are the main findings?}
+% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
+% \item First bullet.
+% \item Second bullet.
+% \end{itemize}\vspace{3pt}
+%\textbf{What is the implication of the main finding?}
+% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
+% \item First bullet.
+% \item Second bullet.
+% \end{itemize}
+%}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}\label{sec:introduction}
@@
 	\centering
-	\includegraphics[width=.7\linewidth]{figures/solo_fw_pic_compressed.png}
+	\includegraphics[width=.5\linewidth]{figures/solo_fw_pic_compressed.png}
 	\caption{\small The proposed Orientation Control System for 
@@
 Conclusions and a possible evolution of the work ar reported in Section \ref{sec:conclusions}.
-%
 
@@
 All the parameters are reported in Table \ref{tab:dim_fw}, together with the selected inertia and mass of a single flywheel.
-\begin{table}
-	\centering
-	\caption{\small Size and dynamic parameters of a single flywheel}
-	\begin{tabular}{ccc}
-		Parameter & Value & Unit \\
-		\hline
-		\hline
+\begin{table}[H] 
+	\caption{Size and dynamic parameters of a single flywheel}
+	\label{tab:dim_fw}
+	\newcolumntype{C}{>{\centering\arraybackslash}X}
+	\begin{tabularx}{\textwidth}{CCC}
+		\toprule
+		\textbf{Parameter} & \textbf{Value} & \textbf{Unit}\\
+		\midrule
 		$r$ & $2.20\cdot 10^{-2}$ & $\mathrm{m}$ \\
@@
 		$I_{xx}, I_{yy}$   & $3.64 \cdot 10^{-5}$ & $\mathrm{kg \cdot m^2}$ \\
-		$I_{zz}$   & $7.11 \cdot 10^{-5}$ & $ \mathrm{kg \cdot m^2}$ \\
-	\end{tabular}
-	\label{tab:dim_fw}
+		$I_{zz}$   & $7.11 \cdot 10^{-5}$ & $ \mathrm{kg \cdot m^2}$\\
+		\bottomrule
+	\end{tabularx}
 \end{table}
-
+\unskip
 \subsection{Flywheels Control Law}
@@
 
-\small
-\section*{Acknowledgements}	
-All the authors want to thank Dr. Roy Featherstone for interesting tips that started this work. 
-\newline
-The publication was created with the co-financing of the European Union FSE-REACT-EU, PON Research and Innovation 2014-2020 DM1062 / 2021.
+-------------------------------------------------------------------------------
 
-\bibliography{references/bibliography}
-%\printbibliography
+\section{Introduction}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Materials and Methods}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results}
+
+\subsection{Figures, Tables and Schemes}
+
+All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, Table~\ref{tab2}, etc.
+
+\begin{figure}[H]
+\includegraphics[width=10.5 cm]{Definitions/logo-mdpi}
+\caption{This is a figure.
+Schemes follow the same formatting.
+If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel.
+(\textbf{b}) Description of what is contained in the second panel.
+Figures should be placed in the main text near to the first time they are cited.
+A caption on a single line should be centered.\label{fig1}}
+\end{figure}   
+\unskip
+
+\begin{table}[H] 
+\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
+\newcolumntype{C}{>{\centering\arraybackslash}X}
+\begin{tabularx}{\textwidth}{CCC}
+\toprule
+\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
+\midrule
+Entry 1		& Data			& Data\\
+Entry 2		& Data			& Data\\
+\bottomrule
+\end{tabularx}
+\end{table}
+\unskip
+
+\begin{table}[H]
+\caption{This is a wide table.\label{tab2}}
+	\begin{adjustwidth}{-\extralength}{0cm}
+		\newcolumntype{C}{>{\centering\arraybackslash}X}
+		\begin{tabularx}{\fulllength}{CCCC}
+			\toprule
+			\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}     & \textbf{Title 4}\\
+			\midrule
+			Entry 1		& Data			& Data			& Data\\
+			Entry 2		& Data			& Data			& Data \textsuperscript{1}\\
+			\bottomrule
+		\end{tabularx}
+	\end{adjustwidth}
+	\noindent{\footnotesize{\textsuperscript{1} This is a table footnote.}}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Patents}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{6pt}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+%\supplementary{The following supporting information can be downloaded at:  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.}
+
+% Only for the journal Methods and Protocols:
+% If you wish to submit a video article, please do so with any other supplementary material.
+% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
+A supporting video article is available at doi: link.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\authorcontributions{Conceptualization, F.R.
+and M.F.; software, F.R., A.C.
+and M.F.; validation, F.R., A.C.
+and M.F.; formal analysis, F.R., A.C.
+and M.F.; investigation, F.R.
+and A.C.; resources, C.S.; data curation, F.R., A.C.
+and M.F.; writing---original draft preparation, F.R.; writing---review and editing, F.R., A.C., A.D.P., C.S.
+and M.F.; visualization, F.R.
+and A.C.; supervision, A.D.P.
+and M.F; project administration, F.R.
+and M.F.; funding acquisition, \FR{?}.
+All authors have read and agreed to the published version of the manuscript.}
+%\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
+The following statements should be used ``Conceptualization, X.X.
+and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
+and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
+All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
+Authorship must be limited to those who have contributed substantially to the work~reported.}
+
+\funding{The publication was created with the co-financing of the European Union FSE-REACT-EU, PON Research and Innovation 2014-2020 DM1062 / 2021.}
+
+\institutionalreview{Not applicable.}
+
+\informedconsent{Not applicable.}
+
+\dataavailability{Not applicable.}
+
+\acknowledgments{All the authors want to thank Dr. Roy Featherstone for interesting tips that started this work. }
+
+\conflictsofinterest{The authors declare no conflict of interest.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
+%\sampleavailability{Samples of the compounds ... are available from the authors.}
+
+%% Only for journal Encyclopedia
+%\entrylink{The Link to this entry published on the encyclopedia platform.}
+
+\abbreviations{Abbreviations}{
+The following abbreviations are used in this manuscript:\\
+
+\noindent 
+\begin{tabular}{@{}ll}
+CoM & Center Of Mass\\
+OCS & Orientation Control System\\
+CMG & Control Moment Gyroscope\\
+NMPC & Nonlinear Model Predictive Control
+\end{tabular}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
+%\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A").
+If the appendix sections contain a heading then change the argument to "yes".
+%\appendixstart
+%\appendix
+%\section[\appendixname~\thesection]{}
+%\subsection[\appendixname~\thesubsection]{}
+%The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data.
+Mathematical proofs of results not central to the paper can be added as an appendix.
+
+%\begin{table}[H] 
+%\caption{This is a table caption.\label{tab5}}
+%\newcolumntype{C}{>{\centering\arraybackslash}X}
+%\begin{tabularx}{\textwidth}{CCC}
+%\toprule
+%\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
+%\midrule
+%Entry 1		& Data			& Data\\
+%Entry 2		& Data			& Data\\
+%\bottomrule
+%\end{tabularx}
+%\end{table}
+
+%\section[\appendixname~\thesection]{}
+%All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{adjustwidth}{-\extralength}{0cm}
+%\printendnotes[custom] % Un-comment to print a list of endnotes
+
+\reftitle{References}
+
+% Please provide either the correct journal abbreviation (e.g.
+according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
+% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
+
+%=====================================
+% References, variant A: external bibliography
+%=====================================
+\bibliography{references/bibliography.bib}
+
+%=====================================
+% References, variant B: internal bibliography
+%=====================================
+%\begin{thebibliography}{999}
+% Reference 1
+%\bibitem[Author1(year)]{ref-journal}
+%Author~1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
+% Reference 2
+%\bibitem[Author2(year)]{ref-book1}
+%Author~2, L. The title of the cited contribution. In {\em The Book Title}; Editor 1, F., Editor 2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
+% Reference 3
+%\bibitem[Author3(year)]{ref-book2}
+%Author 1, A.; Author 2, B. \textit{Book Title}, 3rd ed.; Publisher: Publisher Location, Country, 2008; pp. 154--196.
+% Reference 4
+%\bibitem[Author4(year)]{ref-unpublish}
+%Author 1, A.B.; Author 2, C. Title of Unpublished Work. \textit{Abbreviated Journal Name} year, \textit{phrase indicating stage of publication (submitted; accepted; in press)}.
+% Reference 5
+%\bibitem[Author5(year)]{ref-communication}
+%Author 1, A.B. (University, City, State, Country); Author 2, C. (Institute, City, State, Country). Personal communication, 2012.
+% Reference 6
+%\bibitem[Author6(year)]{ref-proceeding}
+%Author 1, A.B.; Author 2, C.D.; Author 3, E.F.
+Title of presentation.
+In Proceedings of the Name of the Conference, Location of Conference, Country, Date of Conference (Day Month Year); Abstract Number (optional), Pagination (optional).
+% Reference 7
+%\bibitem[Author7(year)]{ref-thesis}
+%Author 1, A.B. Title of Thesis. Level of Thesis, Degree-Granting University, Location of University, Date of Completion.
+% Reference 8
+%\bibitem[Author8(year)]{ref-url}
+%Title of Site. Available online: URL (accessed on Day Month Year).
+%\end{thebibliography}
+
+% If authors have biography, please use the format below
+%\section*{Short Biography of Authors}
+%\bio
+%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
+%{\textbf{Firstname Lastname} Biography of first author}
+%
+%\bio
+%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
+%{\textbf{Firstname Lastname} Biography of second author}
+
+% For the MDPI journals use author-date citation, please follow the formatting guidelines on http://www.mdpi.com/authors/references
+% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
+% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
+328; \citeyear{ref-journal-3b}, p.475).
+This produces: Wong (1999, p.
+328; 2000, p.
+475)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% for journal Sci
+%\reviewreports{\\
+%Reviewer 1 comments and authors’ response\\
+%Reviewer 2 comments and authors’ response\\
+%Reviewer 3 comments and authors’ response
+%}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\end{adjustwidth}
 \end{document}

### 2022-11-16 matlab/plotoptions.m

@@ -1,0 +1,36 @@
+% Set grids on all axis.
+set(0,'defaultAxesXGrid','on');
+set(0,'defaultAxesYGrid','on');
+set(0,'defaultAxesZGrid','on');
+
+size = 30;
+
+set(0,'defaultAxesXGrid','on');set(0,'defaultAxesYGrid','on');set(0,'defaultAxesZGrid','on');
+% set some other default values
+set(0, 'RecursionLimit', 50);set(0, 'DefaultFigurePaperType', 'A0');
+set(0, 'Defaultlinelinewidth',5);
+set(0,'defaultaxeslinewidth',1)
+set(0, 'defaultpatchlinewidth',1);set(0, 'DefaultFigureWindowStyle', 'normal');
+set(0, 'DefaultAxesBox', 'on');set(0, 'DefaultTextFontSize', size);
+set(0, 'DefaultAxesFontSize', size);set(0, 'DefaultUicontrolFontSize', size);
+set(0, 'Defaulttextinterpreter','latex')
+label_size=50
+%     area(-feet_pen_RFX(idxA),ones(size(time(idxA))),-0.5 ,'facecolor',0.0*ones(1,3),'edgecolor','none','FaceAlpha',0.2,'HandleVisibility','off'); hold on
+%%
+% Blue color:    ,'Color',[0, 0.4470, 0.7410]);
+% Red color:     ,'Color',[0.8500, 0.3250, 0.0980]);
+%%
+% SIM
+%export_fig('../../figures/softTerrainSimComparisonGRF.pdf','-pdf','-transparent');%MAE!!!
+%export_fig('../../figures/softTerrainSimComparisonpen.pdf','-pdf','-transparent');
+%export_fig('../../figures/softTerrainSimComparison.pdf','-pdf','-transparent');
+%export_fig('../../figures/aggresiveTrunkTraj.pdf','-pdf','-transparent');
+%export_fig('../../figures/speedTest.pdf','-pdf','-transparent');
+%export_fig('../../figures/powerConsumption.pdf','-pdf','-transparent');
+% MIX IS SVG
+%export_fig('../../figures/mix2.pdf','-pdf','-transparent');
+% EXP
+%export_fig('../../figures/softTerraineXPComparison.pdf','-dpdf','-transparent'); % MAE!!!!
+%export_fig('../../figures/LongTransExp.pdf','-pdf','-transparent'); % MAE!!!!
[…]

### 2022-12-02 mdpi2022roscia.tex
update

@@
 % Authors, for the paper (add full first names)
-\Author{Francesco Roscia $^{1, ^*}$, Andrea Cumerlotti $^{?}$, Andrea Del Prete $^{2}$, Claudio Semini $^{1}$ and Michele Focchi$^{1, \, 3}$}
+\Author{Francesco Roscia $^{1}$, Andrea Cumerlotti $^{?}$, Andrea Del Prete $^{2}$, Claudio Semini $^{1}$ and Michele Focchi$^{1, \, 3, ^*}$}
 
@@
 % Contact information of the corresponding author
-\corres{Correspondence: francesco.roscia@iit.it}
+\corres{Correspondence: michele.focchi@unitn.it}
 
@@
 % Keywords
-\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}
+\keyword{Legged robot, Orientation Control, Articulated Multi-body System}
 
@@
 	\centering
-	\includegraphics[width=.5\linewidth]{figures/solo_fw_pic_compressed.png}
+	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/solo_fw_pic_compressed.png}
 	\caption{\small The proposed Orientation Control System for 
@@
 % flywheel
-flywheels represent an additional option for controlling the robot orientation.
+Flywheels represent an additional option for controlling the robot orientation.
 Changing the angular velocity of a rotating mass attached to the trunk generates a torque that can reorient the robot.
@@
 On the other hand, direct-drive controlled flywheels enable to create accelerations in both directions,  and  to implement continuous control laws.
-The benefits of this kind of  \gls{ocs} are multiple: 1) flywheels can correct orientation errors during the flight (e.g.
-due to disturbances or tracking inaccuracies in the angular momentum achieved at the lift-off)   % and non-idealities).
-in a continuous manner. 2) They enable to track a time-varying reference (e.g. to have the robot land with a desired angular velocity (possibly zero) and orientation).
-3) Even in presence of contacts, they can enhance the landing phase or stabilize dynamic gaits (e.g. trot), by significantly reducing trunk oscillations.  
-4) The presence of these additional joints, whose only function is to control the orientation, gives the possibility to relieve the effort of the legs.
+\\
+The benefits of this kind of \gls{ocs} are multiple. 
+\begin{enumerate}
+\item Flywheels can correct orientation errors during the flight (e.g. due to disturbances or tracking inaccuracies in the angular momentum achieved at the lift-off)   % and non-idealities).  
+in a continuous manner. 
+\item They enable to track a time-varying reference (e.g. to have the robot land with a desired angular velocity (possibly zero) and orientation).
+\item  Even in presence of contacts, they can enhance the landing phase or stabilize dynamic gaits (e.g. trot), by significantly reducing trunk oscillations.  
+\item  The presence of these additional joints, whose only function is to control the orientation, gives the possibility to relieve the effort of the legs.
+\end{enumerate}
 In more complex scenarios, like for a somersault, limbs and \gls{ocs} can operate in parallel to achieve a rotation angle larger than the one achievable only with legs (e.g., due to torque limitation).
@@
 In this case, it is possible to change the angular velocity of the base link changing the joint positions and velocities, as a result of the \textit{non-holonomy} of the total angular momentum \cite{Wieber16}: if the angular momentum of a certain body changes, the one of the others must change to maintain the total sum constant.
-
-\FR{In Solo12, even if the each limb accounts for about $13.4\%$ of the total mass, changes of the total robot inertia come with the motion of only the upperleg (thigh) and the lowerleg (calf), since the displacement of distribution of the hip mass with the system \acrshort{com} is approximately constant.
-In view of this, each limb accounts for only $7.5\%$ of the total mass in varying the the total robot inertia.}
+\subsection{Preliminary Analysis}
+Most of quadruped robots are designed in such a way that the largest amount of the mass is located at the main body. As a consequence that the moments given the other bodies is moderate. \\
+As an example, let consider the case of Solo12.
+Even if each limb accounts for about $13.4\%$ of the total mass, changes of the total robot inertia come with the motion of only the upperleg (thigh) and the lowerleg (calf), since the displacement of distribution of the hip mass with the system \acrshort{com} is approximately constant.
+In view of this, each limb accounts for only $7.5\%$ of the total mass in varying the the total robot inertia.
+Flywheels can be used for alleviating the lack of control authority.
 
@@
 In this section, we present a procedure to select the inertia of the flywheels, and we show how to exploit them for simultaneously controlling the robot roll and pitch orientation.
+
 \subsection{Bounds on the Inertia}
-A preliminary analysis for the selection of the flywheels' inertia can be performed with the Elroy's Beanie model, depicted in Fig. \ref{fig:elroy-beanie}.
+Investigation for the selection of the flywheels' inertia can be performed with the Elroy's Beanie model, depicted in Fig. \ref{fig:elroy-beanie}.
 \begin{figure}
@@
 	\label{fig:elroy-beanie}
+\end{figure} 
+\begin{figure}
+	\centering
+	\includegraphics[width=.8\linewidth]{figures/lowerbound2.eps}
+	\caption{\small Minimum inertia of the \gls{ocs} (normalized by the robot inertia) necessary to achieve a base velocity variation $\Delta \dot{\theta}$, given the actuation bound $\dot \gamma^{max}$.}
+	\label{fig:inertia_limits}
 \end{figure}
@@
 Fig. \ref{fig:inertia_limits} reports the lower bound of $I_{f}$ given the desired base velocity variation and the maximum velocity of the actuator.
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{figures/lowerbound.eps}
-	\caption{\small Minimum inertia of the \gls{ocs} (normalized by the robot inertia) necessary to achieve a base velocity variation $\Delta \dot{\theta}$, given the actuation bound $\dot \gamma^{max}$.}
-	\label{fig:inertia_limits}
-\end{figure}
 
 \subsection{Flywheels Axes of Rotation}
-
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/axes.eps}
+	\caption{\small Representation of the flywheels rotation axes, seen from the top of the robot.}
+	\label{fig:axes}
+\end{figure}
 The orientation of the axis of rotation of each flywheel influences their contribution to the total angular momentum.
@@
 If $\alpha = 0$, the roll results to be uncontrollable through the \gls{ocs}; otherwise, if $\alpha = \pi/2$, the pitch is uncontrollable.
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{figures/axes.eps}
-	\caption{\small Representation of the flywheels rotation axes, seen from the top of the robot.}
-	\label{fig:axes}
-\end{figure}
 The angle is selected considering the ratio of the eigenvalues along the $x$ and $y$ directions of the ellipsoid of the robot inertia(Fig. \ref{fig:ellipsoid}):
@@
 \end{equation*}
-\begin{figure}
+\begin{figure}[t]
 	\centering
@@
 	\centering
-	\includegraphics[width=.8\linewidth]{figures/sum_diff_mod.eps}
+	\includegraphics[width=.7\linewidth]{figures/sum_diff_mod.eps}
 	\caption{\small Having incident rotation axes, the \gls{ocs} allows to control both the robot roll and pitch.
-Notice that the roll is influenced by the difference of the angular speeds of the flywheels, $\omega_x = \left(\omega_{fl} - \omega_{fr}\right)\sin(\alpha)$.
-On the contrary, the pitch depends on the sum of the angular speeds $\omega_y = \left(\omega_{fl} -  \omega_{fr}\right)\cos(\alpha)$.}
+Notice that the roll angle is influenced by the difference of the angular speeds of the flywheels, $\omega_x = \left(\omega_{fl} - \omega_{fr}\right)\sin(\alpha)$.
+On the contrary, the pitch angle depends on the sum of the angular speeds $\omega_y = \left(\omega_{fl} +  \omega_{fr}\right)\cos(\alpha)$.}
 	\label{fig:sum-diff}
@@
 Once the desired inertia $I_f$ is chosen, according to the 2D Elroy's Beanie model, it is used to realize the 3D \gls{ocs} (Fig. \ref{fig:flywheel}). 
-\begin{figure}
-	\centering
-	\includegraphics[width=0.7\linewidth]{figures/flywheel_opaque.png}
-	\caption{\small The final design of the flywheel, shown together with its inertia principal axes. Here, $h$ is the wheel thickness, $r$ is the inner radius of the wheel and $R$ is the outer one.}
-	\label{fig:flywheel}
-\end{figure}
 We decided to design the flywheels in the shape of hollow cylinders.
@@
 \begin{table}[H] 
-	\caption{Size and dynamic parameters of a single flywheel}
+	\caption{Sizes and dynamic parameters of a single flywheel}
 	\label{tab:dim_fw}
@@
 		$h$   & $1.02\cdot 10^{-2}$ & $\mathrm{m}$ \\
-		$\rho$   & $7.860 \cdot 10^{3}$ & $\mathrm{kg/m^3}$ \\
+		$\rho$   & $7.86 \cdot 10^{+3}$ & $\mathrm{kg/m^3}$ \\
 		$m$   & $1.02\cdot 10^{-1}$ & $\mathrm{kg}$ \\
@@
 		$I_{zz}$   & $7.11 \cdot 10^{-5}$ & $ \mathrm{kg \cdot m^2}$\\
-		\bottomrule
-	\end{tabularx}
+		\bottomrule	\end{tabularx}
 \end{table}
+\begin{figure}[t!]
+	\centering
+	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/flywheel_opaque.png}
+	\caption{\small The final design of the flywheel, shown together with its inertia principal axes. Here, $h$ is the wheel thickness, $r$ is the inner radius of the wheel and $R$ is the outer one.}
+	\label{fig:flywheel}
+\end{figure}
 \unskip
+
 \subsection{Flywheels Control Law}
@@
 \begin{equation*}
-\bm{\tau}_{fb} = - \tensor[_b]{\bm{I}}{_r} \left(\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}}\right) - \tensor[_b]{\bm{\omega}}{_r} \times \left( \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\bm{\omega}}{_r}\right)
+\bm{\tau}_{fb} = - \tensor[_b]{\bm{I}}{_r} \left(\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}}\right) - \tensor[_b]{\bm{\omega}}{_r} \times \left( \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\bm{\omega}}{_r}\right).
 \end{equation*}
@@
 \label{eq:ctrl_law}
-\bm{u} = \bm{C}^T \bm{\tau}_{fb}
+\bm{u} = \bm{C}^T \bm{\tau}_{fb}.
 \end{equation}
@@
 
-%this is the main test! 
-The necessity of having an \gls{ocs} is revealed with the first test: disturbance that occurs when the robot has no contact with the ground. 
-During the flight phase of a forward jump, $0.1~\mathrm{s}$ after the lift-off, an external disturbance moment is applied to the robot base, deflecting the robot orientation: this disturbance is set to $\bm{\tau}_{dist} = \left[\begin{array}{ccc}
-0.2 & 0.8 & 0.0
-\end{array}
-\right]^T \ \mathrm{N \cdot m}$ and applied for $0.05 \ \mathrm{s}$. 
-If the flywheels are not used, Solo12 falls after the touch-down. Instead, 
-using the flywheels %pairing the legs joint PD controller with \ref{eq:ctrl_law},
-it is possible to drive the robot to a safe configuration after the landing (Fig. \ref{fig:fw_jump}), without need to implement of a specific landing strategy, as the one in \cite{jeon2021real}.
-\begin{figure}
+\begin{figure}[t!]
 	\centering
-	\includegraphics[width=1\linewidth]{figures/fw_jump.eps}
+	\includegraphics[width=.7\linewidth]{figures/fw_jump.eps}
 	\caption{\small First Test: simulation results showing the $\mathrm{CoM}_x$ and pitch trajectory vs.
@@
 \end{figure}
-
-In the second test, we demonstrate the ability of the \gls{ocs} to reorient the base.
-We let the robot fall from a height of $1 \ \mathrm{m}$ with an initial pitch orientation of $30^\circ$ and zero base angular velocity.
-If we do not actuate the flywheels, the robot touches the ground with the same initial orientation and the trunk oscillates.
-If the \gls{ocs} is enabled, it drives the robot base to be horizontal when it is still in the air and the oscillations after the touch-down are drastically reduced, both in the pitch angle and forward direction (Fig.
-\ref{fig:fall_plot}).
-\begin{figure}
+\begin{figure}[H]
 	\centering
-	\includegraphics[width=1\linewidth]{figures/fall_plot.eps}
+	\includegraphics[width=.7\linewidth]{figures/fall_plot.eps}
 	\caption{\small Second Test: simulation results showing the $\mathrm{CoM}_x$ and pitch trajectory vs.
@@
 %The results point out the possibility to use this device even when the angular momentum is not conserved (i.e. when contacts are present).
+%this is the main test! 
+The necessity of having an \gls{ocs} is revealed with the first test: disturbance that occurs when the robot has no contact with the ground. 
+During the flight phase of a forward jump, $0.1~\mathrm{s}$ after the lift-off, an external disturbance moment is applied to the robot base, deflecting the robot orientation: this disturbance is set to $\bm{\tau}_{dist} = \left[\begin{array}{ccc}
+0.2 & 0.8 & 0.0
+\end{array}
+\right]^T \ \mathrm{N \cdot m}$ and applied for $0.05 \ \mathrm{s}$. 
+If the flywheels are not used, Solo12 falls after the touch-down. Instead, 
+using the flywheels %pairing the legs joint PD controller with \ref{eq:ctrl_law},
+it is possible to drive the robot to a safe configuration after the landing (Fig. \ref{fig:fw_jump}), without need to implement of a specific landing strategy, as the one in \cite{jeon2021real}.
+
+In the second test, we demonstrate the ability of the \gls{ocs} to reorient the base.
+We let the robot fall from a height of $1 \ \mathrm{m}$ with an initial pitch orientation of $30^\circ$ and zero base angular velocity.
+If we do not actuate the flywheels, the robot touches the ground with the same initial orientation and the trunk oscillates.
+If the \gls{ocs} is enabled, it drives the robot base to be horizontal when it is still in the air and the oscillations after the touch-down are drastically reduced, both in the pitch angle and forward direction (Fig.
+\ref{fig:fall_plot}).
 
@@
 
--------------------------------------------------------------------------------
-
-\section{Introduction}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-
-\subsection{Figures, Tables and Schemes}
-
-All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, Table~\ref{tab2}, etc.
-
-\begin{figure}[H]
-\includegraphics[width=10.5 cm]{Definitions/logo-mdpi}
-\caption{This is a figure.
-Schemes follow the same formatting.
-If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel.
-(\textbf{b}) Description of what is contained in the second panel.
-Figures should be placed in the main text near to the first time they are cited.
-A caption on a single line should be centered.\label{fig1}}
-\end{figure}   
-\unskip
-
-\begin{table}[H] 
-\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
-\newcolumntype{C}{>{\centering\arraybackslash}X}
-\begin{tabularx}{\textwidth}{CCC}
-\toprule
-\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
-\midrule
-Entry 1		& Data			& Data\\
-Entry 2		& Data			& Data\\
-\bottomrule
-\end{tabularx}
-\end{table}
-\unskip
-
-\begin{table}[H]
-\caption{This is a wide table.\label{tab2}}
-	\begin{adjustwidth}{-\extralength}{0cm}
-		\newcolumntype{C}{>{\centering\arraybackslash}X}
-		\begin{tabularx}{\fulllength}{CCCC}
-			\toprule
-			\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}     & \textbf{Title 4}\\
-			\midrule
-			Entry 1		& Data			& Data			& Data\\
-			Entry 2		& Data			& Data			& Data \textsuperscript{1}\\
-			\bottomrule
-		\end{tabularx}
-	\end{adjustwidth}
-	\noindent{\footnotesize{\textsuperscript{1} This is a table footnote.}}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Patents}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 and M.F; project administration, F.R.
-and M.F.; funding acquisition, \FR{?}.
+and M.F.; funding acquisition, M.F., C.S.
 All authors have read and agreed to the published version of the manuscript.}

### 2022-12-02 mdpi2022roscia.tex
andrea cumerlotti affiliation

@@
 % Authors, for the paper (add full first names)
-\Author{Francesco Roscia $^{1}$, Andrea Cumerlotti $^{?}$, Andrea Del Prete $^{2}$, Claudio Semini $^{1}$ and Michele Focchi$^{1, \, 3, ^*}$}
+\Author{Francesco Roscia $^{1}$, Andrea Cumerlotti $^{1, \, 2}$, Andrea Del Prete $^{3}$, Claudio Semini $^{1}$ and Michele Focchi$^{1, \, 2, ^*}$}
 
@@
 $^{1}$ \quad Dynamic Legged Systems (DLS) lab, Istituto Italiano di Tecnologia (IIT), Genoa, Italy\\
-$^{2}$ \quad Industrial Engineering Department (DII), University of Trento, Trento, Italy\\
-$^{3}$ \quad Department of Information Engineering and Computer Science (DISI), University of Trento, Trento, Italy}
+$^{2}$ \quad Department of Information Engineering and Computer Science (DISI), University of Trento, Trento, Italy\\
+$^{3}$ \quad Industrial Engineering Department (DII), University of Trento, Trento, Italy\\
+}
 

### 2022-12-02 mdpi2022roscia.tex
contribution acceleration
add i.e.

@@
 \subsection{Preliminary Analysis}
-Most of quadruped robots are designed in such a way that the largest amount of the mass is located at the main body. As a consequence that the moments given the other bodies is moderate. \\
+Most of quadruped robots are designed in such a way that the largest amount of the mass is located at the main body.
+As a consequence, the contribution of the moments due to the acceleration of other bodies (i.e.
+leg links) is moderate.
+\\
 As an example, let consider the case of Solo12.
@@
 In this work, we have presented the design of a novel \gls{ocs}, composed of two flywheels, that enables to control the trunk of a legged robot platform, 
-increasing the accuracy of aerial maneuvers during the under-actuated phases (i.e., flight phase) as well as 
+increasing the accuracy of aerial maneuvers during the under-actuated phases (i.e. flight phase) as well as 
 enhancing stability when in contact by damping oscillations.

### 2022-12-06 mdpi2022roscia.tex
typo correction on affiliations

@@
 % Authors, for the paper (add full first names)
-\Author{Francesco Roscia $^{1}$, Andrea Cumerlotti $^{1, \, 2}$, Andrea Del Prete $^{3}$, Claudio Semini $^{1}$ and Michele Focchi$^{1, \, 2, ^*}$}
+\Author{Francesco Roscia $^{1}$, Andrea Cumerlotti $^{1, \, 2}$, Andrea Del Prete $^{2}$, Claudio Semini $^{1}$ and Michele Focchi$^{1, \, 3, ^*}$}
 
@@
 $^{1}$ \quad Dynamic Legged Systems (DLS) lab, Istituto Italiano di Tecnologia (IIT), Genoa, Italy\\
-$^{2}$ \quad Department of Information Engineering and Computer Science (DISI), University of Trento, Trento, Italy\\
-$^{3}$ \quad Industrial Engineering Department (DII), University of Trento, Trento, Italy\\
+$^{2}$ \quad Industrial Engineering Department (DII), University of Trento, Trento, Italy\\
+$^{3}$ \quad Department of Information Engineering and Computer Science (DISI), University of Trento, Trento, Italy\\
 }

### 2022-12-06 mdpi2022roscia.tex
light blue area

@@
 in a continuous manner. 
-\item They enable to track a time-varying reference (e.g. to have the robot land with a desired angular velocity (possibly zero) and orientation).
+\item They enable to track a time-varying reference, e.g. to have the robot land with a desired angular velocity (possibly zero) and orientation.
 \item  Even in presence of contacts, they can enhance the landing phase or stabilize dynamic gaits (e.g. trot), by significantly reducing trunk oscillations.  
@@
 	\centering
-	\includegraphics[width=.7\linewidth]{figures/fw_jump.eps}
+	\includegraphics[width=.7\linewidth]{figures/drawing.png}
 	\caption{\small First Test: simulation results showing the $\mathrm{CoM}_x$ and pitch trajectory vs.
@@
 A disturbance moment $\bm{\tau}_{dist}$ on the trunk can be compensated only if the \gls{ocs} is enabled.
-If it is disabled, the robot is unable to restore a safe configuration after the touch-down and eventually falls.}
+If it is disabled, the robot is unable to restore a safe configuration after the touch-down and eventually falls.
+The light blue area represents the interval of time in which the disturbance is applied.}
 	\label{fig:fw_jump}

### 2023-01-03 mdpi2022roscia.tex
imporvements after reviews

@@
 Legged robots are designed for traversing rough terrain.
-Different types of gaits, such as trot or crawl, have been developed for quadrupedal robots. 
+Different types of gaits, such as trot \MF{\cite{barasuol2013reactive}} or crawl \MF{\cite{focchi2020heuristic}}, have been developed for quadrupedal robots. 
 Thanks to the progress of the last two decades, robots have become lighter and stronger, which has enabled them to perform agile locomotion.
@@
 This means that it is possible to control the base angular velocity by changing the inertia of the robot, e.g., changing the joints configuration.
-
 Nevertheless, the majority of quadrupeds are designed with light legs, resulting in limbs that have small influence on the total angular momentum.
@@
 However, the placement of the additional link makes the resulting robot asymmetric.
-Moreover, due to its limited range of motion, a tail can be used only for a single jump, not for a repeated sequence~\cite{johnson2012tail}.
+Moreover, due to its limited range of motion, a tail can be used only for a single jump, not for a repeated sequence \cite{johnson2012tail}.
+\MF{To circumvent these drawbacks, in \cite{an2022design}, \cite{an2020development}, \cite{tang2022towards} the authors attach a morphable inertial tail with 3-\gls{dofs} (pitch, yaw, and telescoping) on a monopod, a biped and a quadruped, respectively, to enhance more agile locomotion and improve safe landing.}
+\\
 % Legs reconfig
-It is possible to obtain a similar result by creating repetitive circular motions with the feet, like in \cite{hoffman2021exploiting} and \cite{kurtz2021mini}.
-In the latter work, the authors proposed special heavy boots for Mini Cheetah and used a neural network to calculate online joint trajectories.
+It is possible to obtain a similar result by creating repetitive circular motions with the feet, like in \cite{hoffman2021exploiting}.
+\MF{Moving the feet outwards increases the robot total inertia, so if a leg is extended during half of the motion and retracted in the other half, a net angular momentum will result on the trunk, causing a rotation.
+In \cite{kurtz2022mini}} the authors proposed special heavy boots for Mini Cheetah and used a neural network to calculate online joint trajectories.
 However, this solution unnecessarily increases the inertia of the legs, complicating the locomotion problem, which can no longer rely on the massless leg assumption.
-%\annotation{trova un modo per dire che questa complicazione è data dalla massless assumption che non è più valida}
+\\
 % Gyros
@@
 This system is widespread in spacecrafts reorientation \cite{yoon2002spacecraft}, but less frequently exploited in robot locomotion, either wheeled \cite{brown1996single} or legged \cite{mikhalkov2021gyrubot}.
-The \gls{cmg} presents interesting capabilities, but the presence of a pan-tilt unit to drive the gyroscope makes it impractical to mount it on a small, lightweight robot.
-
+The \gls{cmg} presents interesting capabilities, but the presence of a pan-tilt unit to drive the gyroscope makes it impractical to mount it on a small, lightweight robot.\\
 % flywheel
@@
 \subsection{Proposed Approach and Contribution}
-In this work, we present a compact \gls{ocs} based on two flywheels, mounted on the trunk of the lightweight quadruped Solo12 \cite{grimminger2020open}, see Fig. \ref{fig:solo12flywheels}.
+In this work, we \MF{introduce} a compact \gls{ocs} based on two flywheels, mounted on the trunk of the lightweight quadruped Solo12 \cite{grimminger2020open}, see Fig. \ref{fig:solo12flywheels}.
 The contributions of the paper are:
@@
 Even if each limb accounts for about $13.4\%$ of the total mass, changes of the total robot inertia come with the motion of only the upperleg (thigh) and the lowerleg (calf), since the displacement of distribution of the hip mass with the system \acrshort{com} is approximately constant.
-In view of this, each limb accounts for only $7.5\%$ of the total mass in varying the the total robot inertia.
+\MF{In view of this, each limb is responsible for only $7.5\%$ of the total mass to vary the total inertia of the robot}.
 Flywheels can be used for alleviating the lack of control authority.
@@
 We want to drive the robot to a desired pitch rate $\dot{\theta}_{des}$ by acting on the wheel speed.
-Without loss of generality, we can assume the flywheel is stationary with the robot at the instant at which the reorientation maneuver starts ($\dot{\gamma}_0 = 0$), having a system angular momentum of $L_0 = \left(I_r + I_{f}\right)\dot{\theta}_0$.
+Without loss of generality, \MF{one} can assume the flywheel is stationary with the robot at the instant at which the reorientation maneuver starts ($\dot{\gamma}_0 = 0$), having a system angular momentum of $L_0 = \left(I_r + I_{f}\right)\dot{\theta}_0$.
 Under the condition of conservation of the angular momentum, this quantity is constant over time and it is possible to estimate the lower bound for $I_{f}$ given a desired pitch rate and the maximum velocity of the flywheels $\dot{\gamma}_{max}$:
@@
 \subsection{Flywheels Axes of Rotation}
+\label{subseq:flywheel_axes}
 \begin{figure}[t]
@@
 	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/axes.eps}
-	\caption{\small Representation of the flywheels rotation axes, seen from the top of the robot.}
+	\caption{\small \MF{Top view of Solo12 with the proposed \gls{ocs}.
+The $x-$ and $y-$axes embody the forward and left directions of the robot base, respectively.
+The picture reports also the unit vectors $\hat{\bm{a}}_l$ and $\hat{\bm{a}}_r$  that represent the axes of rotation of the left flywheel and of the right one: $\alpha$ is the angle of incidence defined in \ref{subseq:flywheel_axes}.}}
 	\label{fig:axes}
@@
 \end{equation*}
-We can use the PD action $\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}_r}$, where $\bm{K}_p$ and $\bm{K}_d$ are diagonal and positive-definite gain matrices for the error in attitude and angular velocity.
+We can use the \MF{proportional and derivative} action $\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}_r}$, where $\bm{K}_p$ and $\bm{K}_d$ are diagonal and positive-definite gain matrices for the error in attitude and angular velocity.
 The orientation error $\bm{e} \in SO(3)$ needs the algebra of the special rotational group to be computed. To avoid singular configurations, we represent the orientation with quaternions.
@@
 and work in parallel with the joints of the legs to achieve a highly dynamic motion. 
-All the simulations are performed in Gazebo.
-The robot full dynamics is modeled with Pinocchio \cite{carpentier2019pinocchio}.
-The references for the joints of the legs are computed off-line using Crocoddyl \cite{mastalli2020crocoddyl} and tracked with a PD joint controller.
-The \gls{ocs} is commanded to
-track base orientation references using the control law introduced in Subsection \ref{subsec:control_law}.\footnote{The video with all the simulations is available at web link \url{https://www.dropbox.com/scl/fo/wevvl0zvfp8gofie6x7rw/h?dl=0&rlkey=pap76cr813a0skrvrxl7qgrb9}.}
+All the simulations are performed using \MF{ROS \cite{quigley2009ros} together with Gazebo \cite{koenig2004design}.
+Physics related parameters are reported in Table \ref{tab:pys_param}.} The robot full dynamics is modeled with Pinocchio \cite{carpentier2019pinocchio}.
+The references for the joints of the legs are computed off-line using Crocoddyl \cite{mastalli2020crocoddyl} and tracked with a proportional-derivative joint controller.
+The \gls{ocs} is commanded to  track base orientation references using the control law introduced in Subsection \ref{subsec:control_law}.\footnote{The video with all the simulations is available at web link \url{https://www.dropbox.com/scl/fo/wevvl0zvfp8gofie6x7rw/h?dl=0&rlkey=pap76cr813a0skrvrxl7qgrb9}.}
 
-\begin{figure}[t!]
+\begin{table}[t!] 
+	\caption{\MF{Physics related parameters used for simulate the robot dynamics in Gazebo}}
+	\label{tab:pys_param}
+	\newcolumntype{C}{>{\centering\arraybackslash}X}
+	\begin{tabularx}{\textwidth}{CC}
+		\toprule
+		\textbf{Parameter} & \textbf{Value} \\
+		\midrule
+		Step size & $0.001 \ \mathrm{s}$\\
+		Real time update rate & $250$ \\
+		Physics engine & ODE \\
+		Solver & Quick (Projected Gauss-Seidel method)\\
+		Iterations & $50$\\
+		Successive Over Relaxation parameter & $1.3$\\
+		Rescaling Moment of Inertia  & no \\
+		Friction model & Pyramid \\		
+		\bottomrule	\end{tabularx}
+\end{table}
+\begin{figure}[t]
 	\centering
@@
 \end{figure}
-\begin{figure}[H]
+
+\begin{figure}[t!]
 	\centering
@@
 	\label{fig:fall_plot}
+\end{figure}
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]{figures/backflip.png}
+	\caption{\small \MF{Third Test: sequence of snapshots of Solo12 performing a back-flip in Gazebo environment.
+The \gls{ocs} alleviates the effort requested to the leg joints, that in this case can be used to accomplish only the vertical motion.}}
+	\label{fig:backflip}
 \end{figure}
@@
 
-Finally, we demonstrate how flywheels can alleviate the effort of the leg joints to achieve a highly dynamic motion: a back-flip. 
-For this, we target a space application carrying out a simulation with Moon gravity ($1.62 \ \mathrm{m/s^2}$).
+Finally, we demonstrate how flywheels can alleviate the effort of the leg joints to achieve a highly dynamic motion: a back-flip (Fig. \ref{fig:backflip}). 
+For this, we target a space application carrying out a simulation with \MF{lunar gravity acceleration} ($1.62 \ \mathrm{m/s^2}$).
 In this way, it is possible to obtain high jumps with a long flight phase, without having to select more powerful actuators.  
@@
 %\entrylink{The Link to this entry published on the encyclopedia platform.}
-
 \abbreviations{Abbreviations}{
@@
 \begin{tabular}{@{}ll}
+DoF & Degree of Freedom\\
 CoM & Center Of Mass\\

### 2023-01-04 mdpi2022roscia.tex
all equations are numbered

@@
 The angle is selected considering the ratio of the eigenvalues along the $x$ and $y$ directions of the ellipsoid of the robot inertia(Fig. \ref{fig:ellipsoid}):
-\begin{equation*}
+\MF{\begin{equation}
 \alpha^{*}=\tan^{-1}\left(\dfrac{I_{r, xx}}{I_{r, yy}}\right).
-\end{equation*}
+\end{equation}}
 \begin{figure}[t]
@@
 To derive a control law based on the robot base orientation, we make use of \eqref{eq:cons_ang_mom}, expressing all the contributions to the time derivative of total angular momentum with the respect to the base reference frame:
-\begin{equation*}
+\MF{\begin{equation}
 \tensor[_b]{\bm{I}}{_f} \, \tensor[_b]{\dot{\bm{\omega}}}{_f} + \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\dot{\bm{\omega}}}{_r} + \tensor[_b]{\bm{\omega}}{_r} \times \left( \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\bm{\omega}}{_r}\right) = 0
-\end{equation*}
+\end{equation}}
 From this expression, we can define the moment on the base caused by the acceleration of the flywheels, that can be used as a feedback torque $\bm{\tau}_{fb}$:
-\begin{equation*}
+\MF{\begin{equation}
 \begin{split}
@@
 \end{split}
-\end{equation*}
+\end{equation}}
 We can use the \MF{proportional and derivative} action $\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}_r}$, where $\bm{K}_p$ and $\bm{K}_d$ are diagonal and positive-definite gain matrices for the error in attitude and angular velocity.
@@
 The derivative error can be computed using $\dot{\bm{e}} = \tensor[_b]{\bm{\omega}}{_r^{des}} - \tensor[_b]{\bm{\omega}}{_r}$, in which $\tensor[_b]{\bm{\omega}}{_r^{des}}$ and $\tensor[_b]{\bm{\omega}}{_r}$ are, respectively, the desired and actual angular velocity of the base:
-\begin{equation*}
+\MF{\begin{equation}
 \bm{\tau}_{fb} = - \tensor[_b]{\bm{I}}{_r} \left(\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}}\right) - \tensor[_b]{\bm{\omega}}{_r} \times \left( \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\bm{\omega}}{_r}\right).
-\end{equation*}
+\end{equation}}
 Projecting the moment $\bm{\tau}_{fb}$ onto the flywheel axes with $\bm{C}^T$, we obtain the control action $\bm{u}$

### 2023-01-04 mdpi2022roscia.tex
added python

@@
 The references for the joints of the legs are computed off-line using Crocoddyl \cite{mastalli2020crocoddyl} and tracked with a proportional-derivative joint controller.
-The \gls{ocs} is commanded to  track base orientation references using the control law introduced in Subsection \ref{subsec:control_law}.\footnote{The video with all the simulations is available at web link \url{https://www.dropbox.com/scl/fo/wevvl0zvfp8gofie6x7rw/h?dl=0&rlkey=pap76cr813a0skrvrxl7qgrb9}.}
+The \gls{ocs} is commanded to  track base orientation references using the control law introduced in Subsection \ref{subsec:control_law} \MF{coded in Python}.\footnote{The video with all the simulations is available at web link \url{https://www.dropbox.com/scl/fo/wevvl0zvfp8gofie6x7rw/h?dl=0&rlkey=pap76cr813a0skrvrxl7qgrb9}.}
 

### 2023-01-04 mdpi2022roscia.tex
cite odri github

@@
 \subsection{Proposed Approach and Contribution}
-In this work, we \MF{introduce} a compact \gls{ocs} based on two flywheels, mounted on the trunk of the lightweight quadruped Solo12 \cite{grimminger2020open}, see Fig. \ref{fig:solo12flywheels}.
+In this work, we \MF{introduce} a compact \gls{ocs} based on two flywheels, mounted on the trunk of the lightweight quadruped Solo12 \cite{grimminger2020open}.
+\MF{The open-source hardware and software are reported in \cite{github_solo}.
+Fig.
+\ref{fig:solo12flywheels} shows the \gls{ocs} mounted on the trunk of the robot.
+}
 The contributions of the paper are:

### 2023-01-04 mdpi2022roscia.tex
references update

@@
 Referring to legged robots, this condition occurs when the system is not in contact with the ground or other objects, e.g., during a fall or the flight phase of a jump.
-In this case, it is possible to change the angular velocity of the base link changing the joint positions and velocities, as a result of the \textit{non-holonomy} of the total angular momentum \cite{Wieber16}: if the angular momentum of a certain body changes, the one of the others must change to maintain the total sum constant.
+In this case, it is possible to change the angular velocity of the base link changing the joint positions and velocities, as a result of the \textit{non-holonomy} of the total angular momentum \cite{wieber2016modeling}: if the angular momentum of a certain body changes, the one of the others must change to maintain the total sum constant.
 \subsection{Preliminary Analysis}

### 2023-01-04 mdpi2022roscia.tex
small modifications

@@
 		Real time update rate & $250$ \\
-		Physics engine & ODE \\
+		Physics engine & Open Dynamics Engine (ODE) \\
 		Solver & Quick (Projected Gauss-Seidel method)\\
@@
 	\caption{\small \MF{Third Test: sequence of snapshots of Solo12 performing a back-flip in Gazebo environment.
-The \gls{ocs} alleviates the effort requested to the leg joints, that in this case can be used to accomplish only the vertical motion.}}
+The \gls{ocs} alleviates the effort requested to the leg joints, that in this case can be used to accomplish only the vertical motion.
+The red, green and blue lines represent the axes of the inertial (world) reference frame.}}
 	\label{fig:backflip}

### 2023-01-05 answersToReviewers.tex
update after reviews

%  LaTeX support: latex@mdpi.com 
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
\documentclass[sensors,reply,submit,pdftex,moreauthors]{Definitions/mdpi} 
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, adolescents, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, alloys, analytica, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, arts, asc, asi, astronomy, atmosphere, atoms, audiolres, automation, axioms, bacteria, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomass, biomechanics, biomed, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, blsf, brainsci, breath, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, chips, cimb, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coasts, coatings, colloids, colorants, commodities, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, covid, crops, cryptography, crystals, csmf, ctn, curroncol, currophthalmol, cyber, dairy, data, dentistry, dermato, dermatopathology, designs, diabetology, diagnostics, dietetics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, ent, entomology, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fintech, fire, fishes, fluids, foods, forecasting, forensicsci, forests, foundations, fractalfract, fuels, futureinternet, futureparasites, futurepharmacol, futurephys, futuretransp, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, hazardousmatters, healthcare, hearts, hemato, heritage, highthroughput, histories, horticulturae, humanities, humans, hydrobiology, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijtm, ijtpp, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jal, jcdd, jcm, jcp, jcs, jdb, jeta, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jzbg, kidney, kidneydial, knowledge, land, languages, laws, life, liquids, literature, livers, logics, logistics, lubricants, lymphatics, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, merits, metabolites, metals, meteorology, methane, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, microplastics, minerals, mining, modelling, molbank, molecules, mps, msf, mti, muscles, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutraceuticals, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, organoids, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacoepidemiology, pharmacy, philosophies, photochem, photonics, phycology, physchem, physics, physiologia, plants, plasma, pollutants, polymers, polysaccharides, poultry, powders, preprints, proceedings, processes, prosthesis, proteomes, psf, psych, psychiatryint, psychoactives, publications, quantumrep, quaternary, qubs, radiation, reactions, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, rheumato, risks, robotics, ruminants, safety, sci, scipharm, seeds, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, software, soilsystems, solar, solids, sports, standards, stats, stresses, surfaces, surgeries, suschem, sustainability, symmetry, synbio, systems, taxonomy, technologies, telecom, test, textiles, thalassrep, thermo, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by: 
% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
% MDPI internal commands
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2022}
\copyrightyear{2022}
%\externaleditor{Academic Editor: Firstname Lastname}
\datereceived{} 
%\daterevised{} % Only for the journal Acoustics
\dateaccepted{} 
\datepublished{} 
%\datecorrected{} % Corrected papers include a "Corrected: XXX" date in the original paper.
%\dateretracted{} % Corrected papers include a "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
\usepackage{glossaries}
\usepackage{bm}
\usepackage{tensor}
%\input{includes/used_packages.tex}
\input{includes/acronyms.tex}
%\input{includes/math_commands.tex}
%\input{includes/newcommands.tex}

\newcommand{\MF}[1]{\textcolor{red}{\textbf{mfocchi}: #1}}
\newcommand{\CS}[1]{\textcolor{violet}{\textbf{csemini}: #1}}
\newcommand{\FR}[1]{\textcolor{teal}{\textbf{froscia}: #1}}
\newcommand{\ADP}[1]{\textcolor{blue}{\textbf{adelprete}: #1}}
%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Orientation Control System: Enhancing Aerial Maneuvers for Quadruped Robots}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Orientation Control System: Enhancing Aerial Maneuvers for Quadruped Robots}

% Author Orchid ID: enter ID or remove command
%\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Francesco Roscia $^{1}$, Andrea Cumerlotti $^{1, \, 2}$, Andrea Del Prete $^{2}$, Claudio Semini $^{1}$ and Michele Focchi$^{1, \, 3, ^*}$}

%\longauthorlist{yes}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Francesco Roscia, Andrea Cumerlotti, Andrea Del Prete, Claudio Semini and Michele Focchi}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Roscia, F.; Cumerlotti, A.; Del Prete, A., Semini, C.; Focchi, M.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Dynamic Legged Systems (DLS) lab, Istituto Italiano di Tecnologia (IIT), Genoa, Italy\\
$^{2}$ \quad Industrial Engineering Department (DII), University of Trento, Trento, Italy\\
$^{3}$ \quad Department of Information Engineering and Computer Science (DISI), University of Trento, Trento, Italy\\
}

% Contact information of the corresponding author
\corres{Correspondence: michele.focchi@unitn.it}

% Current address and/or shared authorship
%\firstnote{Current address: Affiliation 3.} 
%\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\) 
\makeglossaries
% Keywords

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data
%\dataset{DOI number or link to the deposited data set if the data set is published separately. If the data set shall be published as a supplement to this paper, this field will be filled by the journal editors. In this case, please submit the data set as a supplement.}
%\datasetlicense{License under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Encyclopedia
%\encyclopediadef{For entry manuscripts only: please provide a brief overview of the entry title instead of an abstract.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Advances in Respiratory Medicine
%\addhighlights{yes}
%\renewcommand{\addhighlights}{%

%\noindent This is an obligatory section in “Advances in Respiratory Medicine”, whose goal is to increase the discoverability and readability of the article via search engines and other scholars. Highlights should not be a copy of the abstract, but a simple text allowing the reader to quickly and simplified find out what the article is about and what can be cited from it. Each of these parts should be devoted up to 2~bullet points.\vspace{3pt}\\
%\textbf{What are the main findings?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}\vspace{3pt}
%\textbf{What is the implication of the main finding?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Dear Ms. Vera Popovic,\\
Thank you for giving us the chance to submit a revised draft version of the manuscript titled "Orientation Control System: Enhancing Aerial Maneuvers for Quadruped Robots" to the MDPI journal \textit{Sensors}. We are grateful for the time and the effort you and the reviewers have dedicated to providing insightful comments on the original manuscript. We have incorporated changes to reflect most of the provided suggestions: they are highlighted in red within the manuscript. For removing the highlights, please comment line 73 and uncomment line 74 in the file named \texttt{mdpi2022roscia.tex}.\\
Here is a point-by-point response to the reviewers' comments and concerns.

\section*{Comments for Reviewer 1}
\begin{itemize}
	\item \textbf{Comment 1:} \textit{Extend the text of the manuscript (e.g. introduction or conclusion) with specific results in the world and Europe, - Improve the quality of the paper by presenting the results of publications by researchers and experts who are involved in this field and registered in world databases (wos). These are e.g: Measurement of industrial robot pose repeatability, Investigation of snake robot locomotion possibilities in a pipe, thanks.}
	\item[] \textbf{Response:} Thank you for this advice. However, in the case of our study, it seems slightly out of range because we tackle of the angular dynamics of the main body of a quadruped robot, that is observable.	\\
	\item \textbf{Comment 2:} \textit{Figure 4 should be contrasting and readable}
	\item[] \textbf{Response:} Thank you for pointing this out. We have incorporated your suggestion in the revised manuscript. \\
	\item \textbf{Comment 3:} \textit{ Conclusions and future work should be extended to contain practical applications based on research described in this paper - expand references.}
	\item[] \textbf{Response:} We are glad you made such an important hint to improve the quality of the manuscript. Possible applications of the presented OCS include but are not limited to efficiently adjust the posture of quadruped robots walking or jumping on uneven terrains. As proved in the third simulation (back-flip), our approach improves the capabilities of quadrupeds in space application enabling fast locomotion by means of leaps, ensuring a reactive control action on the robot angular momentum. Furthermore, the method presented in Section 3 for designing the OCS does not depend on the specific platform, thus it can be replicated for reorienting mechanical structures with different morphology, e.g, monopods or bipeds. A construction worker's backpack can contain two flywheels with incident rotation axes: in the event of a fall from scaffolding, they can be used to reorient the human body to avoid impact of the head with the ground with the same controller proposed in this work. \\
	\item \textbf{Comment 4:} \textit{Number all mathematical equations}
	\item[] \textbf{Response:} Thank you for bringing this to our attention. We applied the comment into the document.\\	
	\item \textbf{Comment 5:} \textit{For article type, 20 references are not enough. Please add more references (>20) during your revisions.}
	\item[] \textbf{Response:} We really appreciate your suggestion, which helped us to improve the manuscript. We included other relevant works in the literature review, e.g. [An, 2022], [Tang, 2022], and [Kurtz, 2022].\\
	\item \textbf{Comment 6:} \textit{Standardization of literature input (into one style)}
	\item[] \textbf{Response:} Agree. We have corrected the literature input.\\
\end{itemize}

\section*{Comments for Reviewer 2}
\begin{itemize}
	\item \textbf{Comment 1:} \textit{'All the simulations are performed in Gazebo'. The paper must have more details related with the simulation, including a snapshot of the simulated robot in its environment.}
	\item[] \textbf{Response:} Thank you for the insightful comment. We summarized the details of the simulation in \ref{tab:pys_param}, which is incorporated to the manuscript too. Moreover, inspired by your suggestion, we attached a sequence of snapshots for the third simulation (back-flip), showing the robot in the Gazebo environment.
	\begin{table}[H] 
		\caption{Physics related parameters used for simulate the robot dynamics in Gazebo}
		\label{tab:pys_param}
		\newcolumntype{C}{>{\centering\arraybackslash}X}
		\begin{tabularx}{\textwidth}{CC}
			\toprule
			\textbf{Parameter} & \textbf{Value} \\
			\midrule
			Step size & $0.001 \ \mathrm{s}$\\
			Real time update rate & $250$ \\
			Physics engine & Open Dynamics Engine (ODE) \\
			Solver & Quick (Projected Gauss-Seidel method)\\
			Iterations & $50$\\
			Successive Over Relaxation parameter & $1.3$\\
			Rescaling Moment of Inertia  & no \\
			Friction model & Pyramid \\		
			\bottomrule	\end{tabularx}
	\end{table} 
	\item \textbf{Comment 2:} \textit{The authors should give more details regarding the next sentence, being crucial for the paper outputs: "The robot full dynamics is modeled with Pinocchio". Currently Gazebo supports 4 physics engines: ode, bullet, simbody and dart, being the default physics engine ode. Since the simulator models the robot dynamics, why use Pinocchio software?}
	\item[] \textbf{Response:} We acknowledge that we could have made a better explanation here. We rephrased "The robot full dynamics is modeled with Pinocchio. The references for the joints of the legs are computed off-line using Crocoddyl and tracked with a proportional-derivative joint controller." with "References for the joints of the legs are computed off-line using Crocoddyl, an optimal control library for robots based on Differential Dynamic Programming (DDP) algorithms. It uses Pinocchio for fast computation of robots dynamics and their analytical derivatives. References $\bm{q}_{ref}$, $\dot{\bm{q}}_{ref}$ and $\bm{\tau}_{ref}$ for joint positions, velocities and torques are then executed on-line with a proportional-derivative joint controller:
	\begin{equation*}
	\bm{\tau}_j = \bm{K}_{p,\, j} (\bm{q}_{ref} - \bm{q}) + \bm{K}_{d,\, j} (\dot{\bm{q}}_{ref} - \dot{\bm{q}}) + \bm{\tau}_{ref}"
	\end{equation*}
	\item \textbf{Comment 3:} \textit{ Section 2 should be expanded with more citations, an describing more related works.}
	\item[] \textbf{Response:} Thank you for the suggestion. Reviewer 1 pointed out the same advise, then you could refer to the response to Comment 5 of the the previous section. 
	\item \textbf{Comment 4:} \textit{Ide for robot code development}
	\item[] \textbf{Response:} We coded the controller using Python v3.8, with the IDE Pycharm v3.2 (\url{https://www.jetbrains.com/pycharm/}). The manuscript lacks about this information because we believe that, in our work, the use of a specific IDE is not relevant.
\end{itemize}
\vspace{12 pt}

In addition to the above observations, all orthographic and grammatical errors mentioned by the reviewers have been corrected. \\
We look forward to hearing from you in due time regarding our submission and to respond to any further questions and comments you may have.
Sincerely,

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following supporting information can be downloaded at:  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
%\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
%\appendixstart
%\appendix
%\section[\appendixname~\thesection]{}
%\subsection[\appendixname~\thesubsection]{}
%The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data. Mathematical proofs of results not central to the paper can be added as an appendix.

%\begin{table}[H] 
%\caption{This is a table caption.\label{tab5}}
%\newcolumntype{C}{>{\centering\arraybackslash}X}
%\begin{tabularx}{\textwidth}{CCC}
%\toprule
%\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
%\midrule
%Entry 1		& Data			& Data\\
%Entry 2		& Data			& Data\\
%\bottomrule
%\end{tabularx}
%\end{table}

%\section[\appendixname~\thesection]{}
%All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{adjustwidth}{-\extralength}{0cm}
%\printendnotes[custom] % Un-comment to print a list of endnotes

%\reftitle{References}

% Please provide either the correct journal abbreviation (e.g. according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: external bibliography
%=====================================
%\bibliography{references/bibliography.bib}

%=====================================
% References, variant B: internal bibliography
%=====================================
%\begin{thebibliography}{999}
% Reference 1
%\bibitem[Author1(year)]{ref-journal}
%Author~1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
% Reference 2
%\bibitem[Author2(year)]{ref-book1}
%Author~2, L. The title of the cited contribution. In {\em The Book Title}; Editor 1, F., Editor 2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
% Reference 3
%\bibitem[Author3(year)]{ref-book2}
%Author 1, A.; Author 2, B. \textit{Book Title}, 3rd ed.; Publisher: Publisher Location, Country, 2008; pp. 154--196.
% Reference 4
%\bibitem[Author4(year)]{ref-unpublish}
%Author 1, A.B.; Author 2, C. Title of Unpublished Work. \textit{Abbreviated Journal Name} year, \textit{phrase indicating stage of publication (submitted; accepted; in press)}.
% Reference 5
%\bibitem[Author5(year)]{ref-communication}
%Author 1, A.B. (University, City, State, Country); Author 2, C. (Institute, City, State, Country). Personal communication, 2012.
% Reference 6
%\bibitem[Author6(year)]{ref-proceeding}
%Author 1, A.B.; Author 2, C.D.; Author 3, E.F. Title of presentation. In Proceedings of the Name of the Conference, Location of Conference, Country, Date of Conference (Day Month Year); Abstract Number (optional), Pagination (optional).
% Reference 7
%\bibitem[Author7(year)]{ref-thesis}
%Author 1, A.B. Title of Thesis. Level of Thesis, Degree-Granting University, Location of University, Date of Completion.
% Reference 8
%\bibitem[Author8(year)]{ref-url}
%Title of Site. Available online: URL (accessed on Day Month Year).
%\end{thebibliography}

% If authors have biography, please use the format below
%\section*{Short Biography of Authors}
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
%{\textbf{Firstname Lastname} Biography of first author}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
%{\textbf{Firstname Lastname} Biography of second author}

% For the MDPI journals use author-date citation, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\end{adjustwidth}
\end{document}

### 2023-01-05 mdpi2022roscia.tex
updated pdf
update after reviews

@@
 
-\newcommand{\MF}[1]{\textcolor{red}{\textbf{mfocchi}: #1}}
+\newcommand{\MF}[1]{\textcolor{red}{#1}}
+%\newcommand{\MF}[1]{\textcolor{black}{#1}}
 \newcommand{\CS}[1]{\textcolor{violet}{\textbf{csemini}: #1}}
@@
 The axes of rotation of the flywheels are designed to be incident, leading to a compact orientation control system that is capable of controlling both roll and pitch angles, considering the different moment of inertia in the two directions.
-We tested the concept with simulations on the robot Solo12.}
+\MF{The concept is tested with simulations on the robot Solo12.}}
 
@@
 
-\begin{figure}[t!]
-	\centering
-	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/solo_fw_pic_compressed.png}
-	\caption{\small The proposed Orientation Control System for 
-		for the $2.5 \ \mathrm{kg}$ quadruped robot Solo12 consists of two $0.1 \ \mathrm{kg}$ flywheels with incident rotation axes.
-Each wheel is located in a 3D-printed shell and mounted on the trunk body.}
-	\label{fig:solo12flywheels}
-\end{figure}
-
 Quadrupedal animals, like cats, can rearrange tail and trunk to correct the orientation during a fall \cite{kane1969dynamical}.
@@
 \subsection{Proposed Approach and Contribution}
-In this work, we \MF{introduce} a compact \gls{ocs} based on two flywheels, mounted on the trunk of the lightweight quadruped Solo12 \cite{grimminger2020open}.
+\begin{figure}[t!]
+	\centering
+	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/solo_fw_pic_compressed.png}
+	\caption{\small The proposed Orientation Control System for 
+		for the $2.5 \ \mathrm{kg}$ quadruped robot Solo12 consists of two $0.1 \ \mathrm{kg}$ flywheels with incident rotation axes.
+Each wheel is located in a 3D-printed shell and mounted on the trunk body.}
+	\label{fig:solo12flywheels}
+\end{figure} 
+\MF{This work introduces} a compact \gls{ocs} based on two flywheels, mounted on the trunk of the lightweight quadruped Solo12 \cite{grimminger2020open}.
 \MF{The open-source hardware and software are reported in \cite{github_solo}.
@@
 \label{sec:ocs}
-In this section, we present a procedure to select the inertia of the flywheels, and we show how to exploit them for simultaneously controlling the robot roll and pitch orientation.
+In this section, a procedure to select the inertia of the flywheels \MF{is presented, together with a strategy} to exploit them for simultaneously controlling the robot roll and pitch orientation.
 
@@
 Investigation for the selection of the flywheels' inertia can be performed with the Elroy's Beanie model, depicted in Fig. \ref{fig:elroy-beanie}.
-\begin{figure}
+\begin{figure}[t!]
 	\centering
@@
 \end{figure} 
-\begin{figure}
+\begin{figure}[t]
 	\centering
@@
 The aim is to examine the rotational dynamics of the system as a whole.
-To simplify the analysis, in the following we will refer only to the effects on the pitch angle of the robot, keeping in mind that the same arguments apply also to the roll.
+To simplify the analysis, in the following \MF{only the effects on the pitch angle of the robot is considered}, keeping in mind that the same arguments apply also to the roll.
 Let us identify the moment of inertia of the robot in the nominal configuration as $I_r$ and the moment of inertia of the two flywheels as $I_f$, both referred to the system \gls{com}.
@@
 where $\dot{\theta}$ and $\dot{\gamma}$ are respectively the robot pitch rate and the angular speed of the wheel.
-We want to drive the robot to a desired pitch rate $\dot{\theta}_{des}$ by acting on the wheel speed.
+\MF{The robot is driven to reach} a desired pitch rate $\dot{\theta}_{des}$ by acting on the wheel speed.
 Without loss of generality, \MF{one} can assume the flywheel is stationary with the robot at the instant at which the reorientation maneuver starts ($\dot{\gamma}_0 = 0$), having a system angular momentum of $L_0 = \left(I_r + I_{f}\right)\dot{\theta}_0$.
@@
 \ref{fig:axes}).
-To remove unnecessary complications, we designed two identical modules, to be mounted on the trunk of Solo12, symmetrically with the respect of the robot sagittal plane.
-We denote with $\alpha \leq \pi/2$ the non-negative incident angle between the wheels' axes of rotation and the robot lateral direction.
+To remove unnecessary complications, \MF{the authors} designed two identical modules, to be mounted on the trunk of Solo12, symmetrically with the respect of the robot sagittal plane.
+\MF{Let} denote with $\alpha \leq \pi/2$ the non-negative incident angle between the wheels' axes of rotation and the robot lateral direction.
 The matrix
@@
 Once the desired inertia $I_f$ is chosen, according to the 2D Elroy's Beanie model, it is used to realize the 3D \gls{ocs} (Fig. \ref{fig:flywheel}). 
-We decided to design the flywheels in the shape of hollow cylinders.
+\MF{Two identical flywheels are designed to have} the shape of hollow cylinders.
 This shape increases the inertia by locating the mass far away from the rotation axis.
@@
 All the parameters are reported in Table \ref{tab:dim_fw}, together with the selected inertia and mass of a single flywheel.
-\begin{table}[H] 
+\begin{figure}[t!]
+	\centering
+	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/flywheel_opaque.png}
+	\caption{\small The final design of the flywheel, shown together with its inertia principal axes. Here, $h$ is the wheel thickness, $r$ is the inner radius of the wheel and $R$ is the outer one.}
+	\label{fig:flywheel}
+\end{figure}
+\begin{table}[t] 
 	\caption{Sizes and dynamic parameters of a single flywheel}
@@
 \end{table}
-\begin{figure}[t!]
-	\centering
-	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/flywheel_opaque.png}
-	\caption{\small The final design of the flywheel, shown together with its inertia principal axes. Here, $h$ is the wheel thickness, $r$ is the inner radius of the wheel and $R$ is the outer one.}
-	\label{fig:flywheel}
-\end{figure}
 \unskip
@@
 \end{equation}}
-From this expression, we can define the moment on the base caused by the acceleration of the flywheels, that can be used as a feedback torque $\bm{\tau}_{fb}$:
+From this expression, the moment on the base caused by the acceleration of the flywheels \MF{is defined}, that can be used as a feedback torque $\bm{\tau}_{fb}$:
 \MF{\begin{equation}
@@
 \end{equation}}
-We can use the \MF{proportional and derivative} action $\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}_r}$, where $\bm{K}_p$ and $\bm{K}_d$ are diagonal and positive-definite gain matrices for the error in attitude and angular velocity.
-The orientation error $\bm{e} \in SO(3)$ needs the algebra of the special rotational group to be computed. To avoid singular configurations, we represent the orientation with quaternions.
+\MF{Proportional and derivative} action $\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}_r}$ \MF{is used}, where $\bm{K}_p$ and $\bm{K}_d$ are diagonal and positive-definite gain matrices for the error in attitude and angular velocity.
+The orientation error $\bm{e} \in SO(3)$ needs the algebra of the special rotational group to be computed. To avoid singular configurations, orientation \MF{are represented} with quaternions.
 The derivative error can be computed using $\dot{\bm{e}} = \tensor[_b]{\bm{\omega}}{_r^{des}} - \tensor[_b]{\bm{\omega}}{_r}$, in which $\tensor[_b]{\bm{\omega}}{_r^{des}}$ and $\tensor[_b]{\bm{\omega}}{_r}$ are, respectively, the desired and actual angular velocity of the base:
@@
 \end{equation}}
-Projecting the moment $\bm{\tau}_{fb}$ onto the flywheel axes with $\bm{C}^T$, we obtain the control action $\bm{u}$
+Projecting the moment $\bm{\tau}_{fb}$ onto the flywheel axes with $\bm{C}^T$, the control action $\bm{u}$ \MF{yields}
 \begin{equation}
@@
 \label{sec:simulations}
-To validate our \gls{ocs}, we present three simulations on different scenarios. We want to test the capability of the proposed approach to: 
+To validate our \gls{ocs}, three simulations on different scenarios \MF{are reported}. \MF{The authors} want to test the capability of the proposed approach to: 
 reject a disturbance when the robot is in the flight phase of a jump, damp trunk oscillations after the touch-down, 
@@
 All the simulations are performed using \MF{ROS \cite{quigley2009ros} together with Gazebo \cite{koenig2004design}.
-Physics related parameters are reported in Table \ref{tab:pys_param}.} The robot full dynamics is modeled with Pinocchio \cite{carpentier2019pinocchio}.
-The references for the joints of the legs are computed off-line using Crocoddyl \cite{mastalli2020crocoddyl} and tracked with a proportional-derivative joint controller.
-The \gls{ocs} is commanded to  track base orientation references using the control law introduced in Subsection \ref{subsec:control_law} \MF{coded in Python}.\footnote{The video with all the simulations is available at web link \url{https://www.dropbox.com/scl/fo/wevvl0zvfp8gofie6x7rw/h?dl=0&rlkey=pap76cr813a0skrvrxl7qgrb9}.}
+Physics related parameters are reported in Table \ref{tab:pys_param}.} \MF{References for the joints of the legs are computed off-line using Crocoddyl, an optimal control library for robots based on \gls{ddp} algorithms.
+It uses Pinocchio for fast computation of robots dynamics and their analytical derivatives.
+References $\bm{q}_{ref}$, $\dot{\bm{q}}_{ref}$ and $\bm{\tau}_{ref}$ for joint positions, velocities and torques are then executed on-line with a proportional-derivative joint controller:
+\begin{equation}
+\bm{\tau}_j = \bm{K}_{p,\, j} (\bm{q}_{ref} - \bm{q}) + \bm{K}_{d,\, j} (\dot{\bm{q}}_{ref} - \dot{\bm{q}}) + \bm{\tau}_{ref}
+\end{equation}
+} The \gls{ocs} is commanded to track base orientation references using the control law introduced in Subsection \ref{subsec:control_law} \MF{coded in Python}.\footnote{The video with all the simulations is available at web link \url{https://www.dropbox.com/scl/fo/wevvl0zvfp8gofie6x7rw/h?dl=0&rlkey=pap76cr813a0skrvrxl7qgrb9}.}
 
@@
 
-In the second test, we demonstrate the ability of the \gls{ocs} to reorient the base.
-We let the robot fall from a height of $1 \ \mathrm{m}$ with an initial pitch orientation of $30^\circ$ and zero base angular velocity.
-If we do not actuate the flywheels, the robot touches the ground with the same initial orientation and the trunk oscillates.
+\MF{The second test demonstrates} the ability of the \gls{ocs} to reorient the base.
+\MF{The robot is falling} from a height of $1.0 \ \mathrm{m}$ with an initial pitch orientation of $30^\circ$ and zero base angular velocity.
+\MF{If the flywheels are not actuated}, the robot touches the ground with the same initial orientation and the trunk oscillates.
 If the \gls{ocs} is enabled, it drives the robot base to be horizontal when it is still in the air and the oscillations after the touch-down are drastically reduced, both in the pitch angle and forward direction (Fig.
@@
 
-Finally, we demonstrate how flywheels can alleviate the effort of the leg joints to achieve a highly dynamic motion: a back-flip (Fig. \ref{fig:backflip}). 
+\MF{Finally, the ability of the flywheels to relieve the effort on the leg joints to achieve a highly dynamic movement is shown with a back-flip} (Fig. \ref{fig:backflip}). 
 For this, we target a space application carrying out a simulation with \MF{lunar gravity acceleration} ($1.62 \ \mathrm{m/s^2}$).
@@
 
-In this work, we have presented the design of a novel \gls{ocs}, composed of two flywheels, that enables to control the trunk of a legged robot platform, 
-increasing the accuracy of aerial maneuvers during the under-actuated phases (i.e. flight phase) as well as 
-enhancing stability when in contact by damping oscillations.
-The novelty of the design, that involves the flywheels to be attached with incident rotational axes on the trunk, allows to control the orientation
-in both the roll and pitch directions while keeping the device compact. Their effect is optimized considering the inertial property of the mechanical structure. 
-Several simulations are reported with the quadruped Solo12 that demonstrate the effectiveness of the proposed approach to: 
-reject disturbances during the flight phase, stabilize the platform after the touch-down even in absence of a specific landing strategy and 
-achieve a fast reorientation maneuver (a back-flip) in a reduced gravity environment.
+\MF{This work presented} the design of a novel \gls{ocs}, composed of two flywheels, that enables to control the trunk of a legged robot platform, increasing the accuracy of aerial maneuvers during the under-actuated phases (i.e.
+flight phase) as well as enhancing stability when in contact by damping oscillations.
+The novelty of the design, that involves the flywheels to be attached with incident rotational axes on the trunk, allows to control the orientation in both the roll and pitch directions while keeping the device compact.
+Their effect is optimized considering the inertial property of the mechanical structure.
+Several simulations are reported with the quadruped Solo12 that demonstrate the effectiveness of the proposed approach to: reject disturbances during the flight phase, stabilize the platform after the touch-down even in absence of a specific landing strategy and achieve a fast reorientation maneuver (a back-flip) in a reduced gravity environment.
 
-In future works, we plan to demonstrate our concept performing experiment with the real platform.
+\MF{Possible applications of the presented \gls{ocs} include but are not limited to efficiently adjust the posture of quadruped robots walking or jumping on uneven terrains.
+As proved in the third simulation (back-flip), our approach improves the capabilities of quadrupeds in space application enabling fast locomotion by means of leaps, ensuring a reactive control action on the robot angular momentum.
+Furthermore, the method presented in Section 3 for designing the OCS does not depend on the specific platform, thus it can be replicated for reorienting mechanical structures with different morphology, e.g, monopods or bipeds.
+A construction worker's backpack can contain two flywheels with incident rotation axes: in the event of a fall from scaffolding, they can be used to reorient the human body to avoid impact of the head with the ground with the same controller proposed in this work.}
+
+In future works, \MF{the authors} plan to demonstrate our concept performing experiment with the real platform.
 The control strategy for defining the base desired angular velocity can be improved using \gls{nmpc}, that takes into consideration the future samples of the orientation reference.
-We expect this feature will allows the stabilization of the yaw to a desired value, that if it is not locally controllable, enhancing the non-holonomy property of the angular momentum, i.e.
+\MF{This feature is expected to allow} the stabilization of the yaw to a desired value, that if it is not locally controllable, enhancing the non-holonomy property of the angular momentum, i.e.
 doing a preliminary roll and pitch maneuver.
@@
 CMG & Control Moment Gyroscope\\
+\MF{DDP} & \MF{Differential Dynamic Programming}\\
 NMPC & Nonlinear Model Predictive Control

### 2023-01-05 answersToReviewers.tex
signature and style cleaning

@@
 \vspace{12 pt} 
+\noindent In addition to the above observations, all orthographic and grammatical errors mentioned have been corrected.
 
-In addition to the above observations, all orthographic and grammatical errors mentioned by the reviewers have been corrected. \\
-We look forward to hearing from you in due time regarding our submission and to respond to any further questions and comments you may have.
-Sincerely,
+We look forward to hear regarding our submission and to respond to any further questions and comments.
+\vspace{18pt}
+\begin{flushright}
+	Sincerely,\\
+	Michele Focchi, Ph.D.	
+\end{flushright}
 

### 2023-01-05 answersToReviewers.tex
typo

@@
 	\item[] \textbf{Response:} Thank you for this advice.
-However, in the case of our study, it seems slightly out of range because we tackle of the angular dynamics of the main body of a quadruped robot, that is observable.	\\
+However, in the case of our study, it seems slightly out of range because we tackle the angular dynamics of the main body of a quadruped robot, that is observable.	\\
 	\item \textbf{Comment 2:} \textit{Figure 4 should be contrasting and readable}

### 2023-01-05 answersToReviewers.tex
typo

@@
 Possible applications of the presented OCS include but are not limited to efficiently adjust the posture of quadruped robots walking or jumping on uneven terrains.
-As proved in the third simulation (back-flip), our approach improves the capabilities of quadrupeds in space application enabling fast locomotion by means of leaps, ensuring a reactive control action on the robot angular momentum.
+As proved in the third simulation (back-flip), our approach improves the capabilities of quadrupeds in space applications enabling fast locomotion by means of leaps, ensuring a reactive control action on the robot angular momentum.
 Furthermore, the method presented in Section 3 for designing the OCS does not depend on the specific platform, thus it can be replicated for reorienting mechanical structures with different morphology, e.g, monopods or bipeds.

### 2023-01-05 answersToReviewers.tex
typo

@@
 Furthermore, the method presented in Section 3 for designing the OCS does not depend on the specific platform, thus it can be replicated for reorienting mechanical structures with different morphology, e.g, monopods or bipeds.
-A construction worker's backpack can contain two flywheels with incident rotation axes: in the event of a fall from scaffolding, they can be used to reorient the human body to avoid impact of the head with the ground with the same controller proposed in this work.
+A construction worker's backpack can contain two flywheels with incident rotation axes: in the event of a fall from scaffolding, they can be used to reorient the human body to avoid the impact of the head with the ground with the same controller proposed in this work.
 \\

### 2023-01-05 answersToReviewers.tex
added fig

@@
 	\item[] \textbf{Response:} Thank you for the insightful comment.
-We summarized the details of the simulation in \ref{tab:pys_param}, which is incorporated to the manuscript too.
-Moreover, inspired by your suggestion, we attached a sequence of snapshots for the third simulation (back-flip), showing the robot in the Gazebo environment.
+We summarized the details of the simulation in Table \ref{tab:pys_param}, which is incorporated to the manuscript too.
+Moreover, inspired by your suggestion, we attached a sequence of snapshots for the third simulation (back-flip), showing the robot in the Gazebo environment (Fig.
+\ref{fig:backflip}).
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=\linewidth]{figures/backflip.png}
+		\caption{\small Third Test: sequence of snapshots of Solo12 performing a back-flip in Gazebo environment.
+The \gls{ocs} alleviates the effort requested to the leg joints, that in this case can be used to accomplish only the vertical motion.
+The red, green and blue lines represent the axes of the inertial (world) reference frame.}
+		\label{fig:backflip}
+	\end{figure}
 	\begin{table}[H] 

### 2023-01-09 answersToReviewers.tex
added Claudio's suggestions

@@
 
-\newcommand{\MF}[1]{\textcolor{red}{\textbf{mfocchi}: #1}}
+\newcommand{\MF}[1]{\textcolor{red}{#1}}
 \newcommand{\CS}[1]{\textcolor{violet}{\textbf{csemini}: #1}}
@@
 	\item[] \textbf{Response:} Thank you for this advice.
-However, in the case of our study, it seems slightly out of range because we tackle the angular dynamics of the main body of a quadruped robot, that is observable.	\\
+However, in the case of our study, it seems slightly out of range because we tackle the angular dynamics of the main body of a quadruped robot, that is observable.
+\MF{Thank you for having recommended us wos.
+The suggested papers seem a bit beyond of our topic.
+Nevertheless, we found on wos a relevant work about morphable 3d tail for quadruped robots [An, 2022], that we included on our manuscript.}	\\
 	\item \textbf{Comment 2:} \textit{Figure 4 should be contrasting and readable}
@@
 \\
+	\MF{We also expanded the number of references, please refer to Comment 5 below.}\\
 	\item \textbf{Comment 4:} \textit{Number all mathematical equations}

### 2023-01-09 mdpi2022roscia.tex
typo

@@
 \begin{table}[t!] 
-	\caption{\MF{Physics related parameters used for simulate the robot dynamics in Gazebo}}
+	\caption{\MF{Physics related parameters used to simulate the robot dynamics in Gazebo}}
 	\label{tab:pys_param}
@@
 \MF{Finally, the ability of the flywheels to relieve the effort on the leg joints to achieve a highly dynamic movement is shown with a back-flip} (Fig. \ref{fig:backflip}). 
-For this, we target a space application carrying out a simulation with \MF{lunar gravity acceleration} ($1.62 \ \mathrm{m/s^2}$).
+For this, we target a space applications carrying out a simulation with \MF{lunar gravity acceleration} ($1.62 \ \mathrm{m/s^2}$).
 In this way, it is possible to obtain high jumps with a long flight phase, without having to select more powerful actuators.  
@@
 As proved in the third simulation (back-flip), our approach improves the capabilities of quadrupeds in space application enabling fast locomotion by means of leaps, ensuring a reactive control action on the robot angular momentum.
-Furthermore, the method presented in Section 3 for designing the OCS does not depend on the specific platform, thus it can be replicated for reorienting mechanical structures with different morphology, e.g, monopods or bipeds.
-A construction worker's backpack can contain two flywheels with incident rotation axes: in the event of a fall from scaffolding, they can be used to reorient the human body to avoid impact of the head with the ground with the same controller proposed in this work.}
+Furthermore, the method presented in Section 3 to design the OCS does not depend on the specific platform, thus it can be replicated for reorienting mechanical structures with different morphology, e.g, monopods or bipeds.
+A construction worker's backpack can contain two flywheels with incident rotation axes: in the event of a fall from scaffolding, they can be used to reorient the human body to impact on the ground with an upright posture with the same controller proposed in this work.}
 

### 2023-01-09 answersToReviewers.tex
new comments of claudio

--- 2023-01-05 answersToReviewers.tex
@@
 These are e.g: Measurement of industrial robot pose repeatability, Investigation of snake robot locomotion possibilities in a pipe, thanks.}
-	\item[] \textbf{Response:} Thank you for this advice.
-However, in the case of our study, it seems slightly out of range because we tackle the angular dynamics of the main body of a quadruped robot, that is observable.	\\
+	\item[] \textbf{Response:} 
+	\MF{Thank you for the constructive comment.
+We have extended the introduction of the manuscript with additional explanations and references to papers about the specific research of orientation control of legged robots during aerial maneuvers.
+The authors of the additionally cited papers are world experts and the majority of the newly added papers are listed in the Web of Science.
+The most recent publications of 2022 are not yet listed there.
+We analysed the two suggested examples of potentially relevant publications to add as a reference.
+We concluded that they are not directly related to the contributions of our work, because we tackle the angular dynamics of the main body of a quadruped robot, that is observable.}	\\
 	\item \textbf{Comment 2:} \textit{Figure 4 should be contrasting and readable}
@@
 	\item[] \textbf{Response:} We are glad you made such an important hint to improve the quality of the manuscript. 
+	\MF{In the conclusions section we have added the following paragraph to discuss the possible practical applications.}\\	
 	Possible applications of the presented OCS include but are not limited to efficiently adjust the posture of quadruped robots walking or jumping on uneven terrains.

### 2023-01-18 mdpi2022roscia_v7/sensors-2114505.tex
for final review

--- 2023-01-04 mdpi2022roscia.tex
@@
 %=================================================================
-\documentclass[sensors,article,submit,pdftex,moreauthors]{Definitions/mdpi} 
+\documentclass[sensors,article,accept,pdftex,moreauthors]{Definitions/mdpi} 
 % For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept".
@@
 \articlenumber{0}
-\pubyear{2022}
-\copyrightyear{2022}
-%\externaleditor{Academic Editor: Firstname Lastname}
-\datereceived{} 
+\pubyear{2023}
+\copyrightyear{2023}
+%学术编辑+dates
+\externaleditor{Academic Editor: Emanuele Lindo Secco 
+}
+\datereceived{6 December 2022 %MDPI: Please add the date.
+}
+\daterevised{6 January 2023 %MDPI: Please add the date.
+}
 %\daterevised{} % Only for the journal Acoustics
-\dateaccepted{} 
-\datepublished{} 
+\dateaccepted{16 January 2023 %MDPI: Please add the date.
+}
+\datepublished{{~} 
+}
 %\datecorrected{} % Corrected papers include a "Corrected: XXX" date in the original paper.
 %\dateretracted{} % Corrected papers include a "Retracted: XXX" date in the original paper.
+
 \hreflink{https://doi.org/} % If needed use \linebreak
@@
 
-\newcommand{\MF}[1]{\textcolor{red}{\textbf{mfocchi}: #1}}
+\newcommand{\MF}[1]{\textcolor{black}{#1}}
+%\newcommand{\MF}[1]{\textcolor{black}{#1}}
 \newcommand{\CS}[1]{\textcolor{violet}{\textbf{csemini}: #1}}
@@
 % Authors, for the paper (add full first names)
-\Author{Francesco Roscia $^{1}$, Andrea Cumerlotti $^{1, \, 2}$, Andrea Del Prete $^{2}$, Claudio Semini $^{1}$ and Michele Focchi$^{1, \, 3, ^*}$}
+\Author{\hl{Francesco} %MDPI: Please carefully check the accuracy of names and affiliations.
+ Roscia $^{1}$, Andrea Cumerlotti $^{1,2}$, Andrea Del Prete $^{2}$, Claudio Semini $^{1}$ and Michele Focchi $^{1,3,}$*}
 
@@
 % MDPI internal command: Authors, for citation in the left column
-\AuthorCitation{Roscia, F.; Cumerlotti, A.; Del Prete, A., Semini, C.; Focchi, M.}
+\AuthorCitation{Roscia, F.; Cumerlotti, A.; Del Prete, A.; Semini, C.; Focchi, M.}
 % If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
@@
 \address{%
-$^{1}$ \quad Dynamic Legged Systems (DLS) lab, Istituto Italiano di Tecnologia (IIT), Genoa, Italy\\
-$^{2}$ \quad Industrial Engineering Department (DII), University of Trento, Trento, Italy\\
-$^{3}$ \quad Department of Information Engineering and Computer Science (DISI), University of Trento, Trento, Italy\\
+$^{1}$ \quad Dynamic Legged Systems (DLS) Lab, Istituto Italiano di Tecnologia (IIT), \hl{Genoa,} %MDPI: Please add the postal code (or ZIP code in the U.S.).
+If the postal code is not available, please provide the P.O.
+Box.
+The following highlights are the same.
+ Italy; \linebreak  \hl{francesco.roscia@iit.it} %MDPI: We added all the emails in affiliations according to the submitting system. Please confirm.
+ (F.R.); andrea.cumerlotti@studenti.unitn.it (A.C.); claudio.semini@iit.it (C.S.)\\
+$^{2}$ \quad Industrial Engineering Department (DII), University of Trento, \hl{Trento}, Italy; andrea.delprete@unitn.it (A.D.P.)\\
+$^{3}$ \quad Department of Information Engineering and Computer Science (DISI), University of Trento, \hl{Trento}, Italy\\
 }
@@
 
-% Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{For legged robots, aerial motions are the only option to overpass obstacles that cannot be circumvent with standard locomotion gaits.
+% Abstract (Do not insert blank lines, i.e., \\) 
+\abstract{For legged robots, aerial motions are the only option to overpass obstacles that cannot be circumvented with standard locomotion gaits.
 In these cases, the robot must perform a leap to either jump onto the obstacle or fly over it.
-However, these movements represent a challenge because during the flight phase the \gls{com} cannot be controlled, and the robot orientation has limited controllability.
-This paper focuses on the latter issue and proposes an \gls{ocs} consisting of two rotating and actuated masses (flywheels or reaction wheels) to gain control authority on the robot orientation.
-Because of the conservation of angular momentum, their rotational velocity can be adjusted to steer the robot orientation even when there are no contacts with the ground.
-The axes of rotation of the flywheels are designed to be incident, leading to a compact orientation control system that is capable of controlling both roll and pitch angles, considering the different moment of inertia in the two directions.
-We tested the concept with simulations on the robot Solo12.}
+However, these movements represent a challenge, because, during the flight phase, the Center of Mass (CoM) cannot be controlled, and there is limited controllability over the orientation of the robot.
+This paper focuses on the latter issue and proposes an  Orientation Control System (OCS), consisting of two rotating and actuated masses (flywheels or reaction wheels), to gain control authority on the orientation of the robot.
+Due to the conservation of angular momentum, the rotational velocity if the robot can be adjusted to steer the robot's orientation, even when the robot has no contact with the ground.
+The axes of rotation of the flywheels are designed to be incident, leading to a compact orientation control system that is capable of controlling both roll and pitch angles, considering the different moments of inertia in the two directions.
+\MF{The concept was tested by means of simulations on the \mbox{robot Solo12}.}}
 
@@
 % Keywords
-\keyword{Legged robot, Orientation Control, Articulated Multi-body System}
+\keyword{legged robot; orientation control; articulated multi-body system}
 
@@
 Different types of gaits, such as trot \MF{\cite{barasuol2013reactive}} or crawl \MF{\cite{focchi2020heuristic}}, have been developed for quadrupedal robots. 
-Thanks to the progress of the last two decades, robots have become lighter and stronger, which has enabled them to perform agile locomotion.
-However, sometimes it is not possible to get around or over an obstacle with the gaits mentioned above, and jumps are required.
+Thanks to progress over the last two decades, robots have become lighter and stronger, which has enabled them to perform with agile locomotion.
+However, sometimes it is not possible for the robot to move around or over an obstacle with the gaits mentioned above, and jumps are required.
 
-When the robot is in the air, the \gls{com} moves on the ballistic trajectory, that is completely defined by the lift-off position and velocity.
-On the other hand, the base orientation can be changed exploiting the conservation of the system angular momentum.
-This means that it is possible to control the base angular velocity by changing the inertia of the robot, e.g., changing the joints configuration.
-Nevertheless, the majority of quadrupeds are designed with light legs, resulting in limbs that have small influence on the total angular momentum.
+When the robot is in the air, the CoM moves on a ballistic trajectory, and this is completely defined by the lift-off position and velocity.
+On the other hand, the base orientation can be changed so as to exploit the conservation of the system's angular momentum.
+This means that it is possible to control the base angular velocity by changing the inertia of the robot, e.g., changing the configuration of the joints.
+Nevertheless, the majority of quadrupeds are designed with light legs, resulting in limbs that have little influence on the total angular momentum.
 
-\begin{figure}[t!]
-	\centering
+Quadrupedal animals, like cats, can rearrange their tails and trunks to correct their orientations during a fall \cite{kane1969dynamical}.
+% Tail
+Much work in the field of robotics has made use of an additional link, such as a tail, as in \cite{chu2019null,wenger2016frontal}.
+This link rotates around an axis that does not pass through the robot's CoM.
+The distances between the axis of rotation and the CoM of both trunk and tail have a large effect on the total inertia, even with a small tail mass.
+However, the placement of the additional link makes the resulting robot asymmetric.
+Moreover, due to its limited range of motion, a tail can be used only for a single jump, not for a repeated sequence \cite{johnson2012tail}.
+\MF{To circumvent these drawbacks, in \cite{an2022design,an2020development,tang2022towards} the authors attached a morphable inertial tail with 3 Degrees of Freedom (DoFs) (pitch, yaw, and telescoping) on a monopod, a biped and a quadruped, respectively, to enhance the agility of locomotion and to improve safety in landing.}
+
+% Legs reconfig
+It is possible to obtain a similar result by creating repetitive circular motions with the feet, like in \cite{hoffman2021exploiting}.
+\MF{Moving the feet outwards increases the robot's total inertia, so if a leg is extended during half of the motion and retracted in the other half, a net angular momentum results on the trunk, causing a rotation.
+In \cite{kurtz2022mini}} the authors proposed special heavy boots for Mini Cheetah and used a neural network to calculate online joint trajectories.
+However, this solution unnecessarily increased the inertia of the legs, complicating the locomotion problem, which could no longer rely on the massless leg assumption.
+
+% Gyros
+Another option is to use a  Control Moment Gyroscope (CMG).
+This consists of a wheel, spinning at a constant angular velocity, inside two or three actuated gimbals.
+Tilting the wheel's axis of rotation generates gyroscopic torque.
+This system is widespread in spacecraft reorientation \cite{yoon2002spacecraft}, but less frequently exploited in robot locomotion, either wheeled \cite{brown1996single} or legged \cite{mikhalkov2021gyrubot}.
+The CMG presents interesting capabilities, but the presence of a pan-tilt unit to drive the gyroscope makes it impractical to mount it on a small, lig\mbox{htweight rob}ot.
+
+% flywheel
+Flywheels represent an additional option for controlling a robot's orientation.
+Changing the angular velocity of a rotating mass attached to the trunk generates a torque that can reorient the robot.
+This device was first applied for spacecraft orientation \cite{oland2009reaction}, and only sporadically investigated in legged locomotion for controlling pitch orientation, both for bipeds \cite{Brown2016,xiong2020sequential} and quadrupeds \cite{kolvenbach2019towards,vasilopoulos2016quadruped}.
+Compared with tails, flywheels do not have position limits, and, since their rotational axes pass through the CoM, angular momentum results in holonomic \cite{machairas2015quadruped}.
+To get a fast response, it is necessary to have an abrupt change in the flywheel's angular velocity (angular acceleration).
+Using a brake avoids the employment of a motor able to deliver higher torques \cite{gajamohan2012cubli}, 
+keeping the system compact. The motor slowly accelerates the wheel to a certain speed to store angular momentum and, when a reorientation is required, the break stops its spin.
+Since the effect of the break is unidirectional, a limitation of this approach lies in the fact that it is only possible to generate a rotation of the base in the opposite direction of the flywheel's angular velocity, making the approach unsuitable for applications where \textit{continuous} controllability is required.
+On the other hand, direct-drive controlled flywheels can create accelerations in both directions,  and can implement continuous control laws.
+
+The benefits of this kind of OCS are multiple. 
+\begin{enumerate}
+\item Flywheels can correct orientation errors (e.g., due to disturbances or tracking inaccuracies in the angular momentum achieved at the lift-off) during flight    % and non-idealities).  
+in a conti\mbox{nuous ma}nner. 
+\item They can track a time-varying reference, e.g., the robot can land with a desired angular velocity (possibly zero) and orientation.
+\item  Even in the presence of contact, the landing phase can be enhanced or dynamic gaits, such as a trot, stabilized, by significantly reducing trunk oscillations.  
+\item  The presence of these additional joints, the only functions of which  are to control orientation, provides the possibility of relieving the efforts of the legs.
+\end{enumerate}
+
+In more complex scenarios, like for a somersault, limbs and OCS can operate in parallel to achieve a rotational angle larger than that achievable only with legs (e.g., due to to\mbox{rque limita}tion).
+
+\subsection{Proposed Approach and Contribution}
+
+\MF{This work introduces} a compact OCS, based on two flywheels, mounted on the trunk of the lightweight quadruped Solo12 \cite{grimminger2020open}.
+\MF{The open-source hardware and software are reported in \cite{github_solo}.
+Figure \ref{fig:solo12flywheels} shows the OCS mounted on the trunk of the robot.} The contributions of the paper are the following:
+\begin{itemize}
+	\item the design of a novel OCS that enables effective control of the orientation of a legged robot during a jump, while keeping the design simple. 
+	In particular, the axes of rotation of the flywheels are set to be incident, enabling \hl{\textit{continuous}} %MDPI: Please confirm if all the italics in the maintext should be retained.
+ controllability in both directions of roll and pitch, while keeping the device compact.
+	\item simulations  with the quadruped Solo12 were conducted and demonstrated the effectiveness of the proposed approach.
+\end{itemize}
+
+\begin{figure}[H]%[t!] 1
+
 	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/solo_fw_pic_compressed.png}
@@
 		for the $2.5 \ \mathrm{kg}$ quadruped robot Solo12 consists of two $0.1 \ \mathrm{kg}$ flywheels with incident rotation axes.
-Each wheel is located in a 3D-printed shell and mounted on the trunk body.}
+Each wheel is located in a 3D-printed shell and mounted on the trunk of the body.}
 	\label{fig:solo12flywheels}
@@
 
-Quadrupedal animals, like cats, can rearrange tail and trunk to correct the orientation during a fall \cite{kane1969dynamical}.
-% Tail
-Many works in the field of robotics used an additional link as a tail, like in \cite{chu2019null} and \cite{wenger2016frontal}.
-This link rotates around an axis that does not pass through the robot \gls{com}: the distances between the axis of rotation and the \gls{com} of both trunk and tail result in a large effect on the total inertia even with a small tail mass.
-However, the placement of the additional link makes the resulting robot asymmetric.
-Moreover, due to its limited range of motion, a tail can be used only for a single jump, not for a repeated sequence \cite{johnson2012tail}.
-\MF{To circumvent these drawbacks, in \cite{an2022design}, \cite{an2020development}, \cite{tang2022towards} the authors attach a morphable inertial tail with 3-\gls{dofs} (pitch, yaw, and telescoping) on a monopod, a biped and a quadruped, respectively, to enhance more agile locomotion and improve safe landing.}
-\\
-% Legs reconfig
-It is possible to obtain a similar result by creating repetitive circular motions with the feet, like in \cite{hoffman2021exploiting}.
-\MF{Moving the feet outwards increases the robot total inertia, so if a leg is extended during half of the motion and retracted in the other half, a net angular momentum will result on the trunk, causing a rotation.
-In \cite{kurtz2022mini}} the authors proposed special heavy boots for Mini Cheetah and used a neural network to calculate online joint trajectories.
-However, this solution unnecessarily increases the inertia of the legs, complicating the locomotion problem, which can no longer rely on the massless leg assumption.
-\\
-% Gyros
-Another option is to use a \gls{cmg}.
-It consists of a wheel, spinning at a constant angular velocity inside two or three actuated gimbals.
-Tilting the wheel's axis of rotation generates the gyroscopic torque.
-This system is widespread in spacecrafts reorientation \cite{yoon2002spacecraft}, but less frequently exploited in robot locomotion, either wheeled \cite{brown1996single} or legged \cite{mikhalkov2021gyrubot}.
-The \gls{cmg} presents interesting capabilities, but the presence of a pan-tilt unit to drive the gyroscope makes it impractical to mount it on a small, lightweight robot.\\
-% flywheel
-Flywheels represent an additional option for controlling the robot orientation.
-Changing the angular velocity of a rotating mass attached to the trunk generates a torque that can reorient the robot.
-This device has been first applied for spacecraft orientation \cite{oland2009reaction}, and later it has been only sporadically investigated in legged locomotion for controlling pitch orientation, both for bipeds \cite{Brown2016}, \cite{xiong2020sequential} and quadrupeds \cite{kolvenbach2019towards}, \cite{vasilopoulos2016quadruped}.
-Compared with tails, flywheels do not have position limits, and since their rotation axes pass through the \gls{com}, its angular momentum results holonomic \cite{machairas2015quadruped}.
-To get a fast response, it is necessary to have an abrupt change in the flywheels angular velocity (angular acceleration).
-Using a brake avoids the employment of a motor able to deliver higher torques \cite{gajamohan2012cubli}, 
-keeping the system compact: the motor slowly accelerates the wheel to a certain speed to store angular momentum and, when a reorientation is required, the break stops its spin.
-Since the effect of the break is unidirectional, a limitation of this approach is given by the fact that it is possible to generate a rotation of the base only in the opposite direction of the flywheel angular velocity making them unsuitable for applications where \textit{continuous} controllability is required.
-On the other hand, direct-drive controlled flywheels enable to create accelerations in both directions,  and  to implement continuous control laws.
-\\
-The benefits of this kind of \gls{ocs} are multiple. 
-\begin{enumerate}
-\item Flywheels can correct orientation errors during the flight (e.g. due to disturbances or tracking inaccuracies in the angular momentum achieved at the lift-off)   % and non-idealities).  
-in a continuous manner. 
-\item They enable to track a time-varying reference, e.g. to have the robot land with a desired angular velocity (possibly zero) and orientation.
-\item  Even in presence of contacts, they can enhance the landing phase or stabilize dynamic gaits (e.g. trot), by significantly reducing trunk oscillations.  
-\item  The presence of these additional joints, whose only function is to control the orientation, gives the possibility to relieve the effort of the legs.
-\end{enumerate}
-In more complex scenarios, like for a somersault, limbs and \gls{ocs} can operate in parallel to achieve a rotation angle larger than the one achievable only with legs (e.g., due to torque limitation).
-
-\subsection{Proposed Approach and Contribution}
-In this work, we \MF{introduce} a compact \gls{ocs} based on two flywheels, mounted on the trunk of the lightweight quadruped Solo12 \cite{grimminger2020open}.
-\MF{The open-source hardware and software are reported in \cite{github_solo}.
-Fig.
-\ref{fig:solo12flywheels} shows the \gls{ocs} mounted on the trunk of the robot.
-}
-The contributions of the paper are:
-\begin{itemize}
-	\item the design of a novel \gls{ocs} that enables to control the orientation of a legged robot during a jump in an effective way, while keeping the design simple. 
-	In particular, the axes of rotation of the flywheels are set to be incident, enabling \textit{continuous} controllability in both the directions (roll and pitch) while keeping the device compact.
-	\item simulations  with the quadruped Solo12 that demonstrate the effectiveness of the proposed approach.
-\end{itemize}
-
 \subsection{Outline}
-The remainder of this paper is organized as follow.
-In Section \ref{sec:background}, the law of conservation of the total angular momentum is recalled.
-In Section \ref{sec:ocs}, the design principles are presented together with the strategy for simultaneously control the robot roll and pitch orientations.
-Section \ref{sec:simulations} presents simulation in different scenarios that demonstrate the the capability of our \gls{ocs} to reject disturbances and to track angular references when there is no contact with the ground, and damping of base oscillation after the touch-down.
-Conclusions and a possible evolution of the work ar reported in Section \ref{sec:conclusions}.
+The remainder of this paper is organized as follows.
+In Section \ref{sec:background}, the law of conservation of the total angular momentum is described.
+In Section \ref{sec:ocs}, the design principles are presented, together with the strategy for simultaneously controlling the robot's roll and pitch orientations.
+Section \ref{sec:simulations} presents the simulation in different scenarios that demonstrate the capability of our OCS to reject disturbances and to track angular references when there is no contact with the ground, and to dampen base oscillation after touch-down.
+Conclusions and a possible evolution of the work are reported in Section \ref{sec:conclusions}.
 
@@
 \label{sec:background}
-The starting point for any \gls{ocs} is the Euler's equation.
-For any mechanical system, the time derivative of the angular momentum $\bm{L}$ computed with the respect to a reference point $O$, fixed in an inertial frame, equals the sum of the moments $\bm{M}_i$ applied to the system with respect to the same reference point:
+The starting point for any OCS is the Euler's equation.
+For any mechanical system, the time derivative of the angular \hl{momentum}, %MDPI: Please confirm if all the bold in equations should be retained.
+ $\bm{L}$ computed in respect to a reference point $O$, fixed in an inertial frame, equals the sum of the moments $\bm{M}_i$ applied to the system with respect to the same reference point:
 \begin{equation}
@@
 \end{equation}
-When the resultant of the external moments applied to the system is zero, the Euler's equation simplifies to:
+\hl{When} %MDPI: Please confirm if the noindent should be retained. The following highlights are the same.
+ the result of the external moments applied to the system is zero, and the Euler's equation simplifies to:
  \begin{equation}
@@
 which is known as conservation of angular momentum.
-Referring to legged robots, this condition occurs when the system is not in contact with the ground or other objects, e.g., during a fall or the flight phase of a jump.
-In this case, it is possible to change the angular velocity of the base link changing the joint positions and velocities, as a result of the \textit{non-holonomy} of the total angular momentum \cite{wieber2016modeling}: if the angular momentum of a certain body changes, the one of the others must change to maintain the total sum constant.
-\subsection{Preliminary Analysis}
-Most of quadruped robots are designed in such a way that the largest amount of the mass is located at the main body.
-As a consequence, the contribution of the moments due to the acceleration of other bodies (i.e.
-leg links) is moderate.
+Referring to legged robots, this condition occurs when the system is not in contact with the ground or with other objects, e.g., during a fall or the flight phase of a jump.
+In this case, it is possible to change the angular velocity of the base link, thereby changing the joint positions and velocities, as a result of the \textit{non-holonomy} of the total angular momentum \cite{wieber2016modeling}.
+If the angular momentum of a certain body changes, then one of the others must change to maintain the total sum as constant.
+\subsection*{Preliminary Analysis}
+Most quadruped robots are designed in such a way that the largest amount of the mass is located in the main body.
+As a consequence, the contribution of the moments due to the acceleration of other bodies (i.e., leg links) is moderate.
 \\
-As an example, let consider the case of Solo12.
-Even if each limb accounts for about $13.4\%$ of the total mass, changes of the total robot inertia come with the motion of only the upperleg (thigh) and the lowerleg (calf), since the displacement of distribution of the hip mass with the system \acrshort{com} is approximately constant.
-\MF{In view of this, each limb is responsible for only $7.5\%$ of the total mass to vary the total inertia of the robot}.
-Flywheels can be used for alleviating the lack of control authority.
+As an example, consider the case of Solo12.
+Even if each limb accounts for about $13.4\%$ of the total mass, changes in the total inertia of the robot come with the motion of only the upper leg (thigh) and the lower leg (calf), since the displacement of distribution of the hip mass with the system's CoM is approximately constant.
+\MF{In view of this, each limb is responsible for only $7.5\%$ of the total mass in varying the total inertia of the robot}.
+Flywheels can be used to alleviate the lack of control authority.
 
@@
 \label{sec:ocs}
-In this section, we present a procedure to select the inertia of the flywheels, and we show how to exploit them for simultaneously controlling the robot roll and pitch orientation.
+In this section, a procedure to select the inertia of the flywheels \MF{is presented, together with a strategy} to exploit them so as to simultaneously control the robot's roll and p\mbox{itch orien}tations.
 
 \subsection{Bounds on the Inertia}
-Investigation for the selection of the flywheels' inertia can be performed with the Elroy's Beanie model, depicted in Fig. \ref{fig:elroy-beanie}.
-\begin{figure}
-	\centering
+Investigation for the selection of the flywheels' inertia can be performed with the Elroy's Beanie model, depicted in Figure \ref{fig:elroy-beanie}.
+\begin{figure}[H]%[t!]2
+
 	\includegraphics[width=.7\linewidth]{figures/elroys-beanie.eps}
@@
 \end{figure}
-\begin{figure}
-	\centering
-	\includegraphics[width=.8\linewidth]{figures/lowerbound2.eps}
-	\caption{\small Minimum inertia of the \gls{ocs} (normalized by the robot inertia) necessary to achieve a base velocity variation $\Delta \dot{\theta}$, given the actuation bound $\dot \gamma^{max}$.}
-	\label{fig:inertia_limits}
-\end{figure}
-This consists of two rigid bodies connected through their \gls{com} with a revolute joint.
-One of the bodies represents the robot in its nominal configuration.
-The other models the \gls{ocs}, here represented by a single wheel for the sake of simplicity.
-The aim is to examine the rotational dynamics of the system as a whole.
-To simplify the analysis, in the following we will refer only to the effects on the pitch angle of the robot, keeping in mind that the same arguments apply also to the roll.
-Let us identify the moment of inertia of the robot in the nominal configuration as $I_r$ and the moment of inertia of the two flywheels as $I_f$, both referred to the system \gls{com}.
-The angular momentum $L$ of this system can be written as
+
+This consists of two rigid bodies connected through their CoMs with a revolute joint.
+One of the bodies represents the robot in its nominal configuration and the other models the OCS, here represented by a single wheel for the sake of simplicity.
+The aim was to examine the rotational dynamics of the system as a whole.
+To simplify the analysis, in the following \MF{only the effects on the pitch angle of the robot was considered}, keeping in mind that the same arguments also apply to the roll.
+Let us identify the moment of inertia of the robot in the nominal configuration as $I_r$ and the moment of inertia of the two flywheels as $I_f$, both referred to the system's CoM.
+The angular momentum $L$ of this system can be written as:
 \begin{equation}\label{eq:ang_mom_elroy}
@@
 \end{equation}
-where $\dot{\theta}$ and $\dot{\gamma}$ are respectively the robot pitch rate and the angular speed of the wheel.
-We want to drive the robot to a desired pitch rate $\dot{\theta}_{des}$ by acting on the wheel speed.
+where $\dot{\theta}$ and $\dot{\gamma}$ are, respectively, the robot;s pitch rate and the angular speed of the wheel.
+\MF{The robot is driven to reach} a desired pitch rate $\dot{\theta}_{des}$ by acting on the wheel speed.
 Without loss of generality, \MF{one} can assume the flywheel is stationary with the robot at the instant at which the reorientation maneuver starts ($\dot{\gamma}_0 = 0$), having a system angular momentum of $L_0 = \left(I_r + I_{f}\right)\dot{\theta}_0$.
-Under the condition of conservation of the angular momentum, this quantity is constant over time and it is possible to estimate the lower bound for $I_{f}$ given a desired pitch rate and the maximum velocity of the flywheels $\dot{\gamma}_{max}$:
+Under the condition of conservation of the angular momentum, this quantity is constant over time and it is possible to estimate the lower bound for $I_{f}$, given a desired pitch rate and the maximum velocity of the flywheels $\dot{\gamma}_{max}$:
 \begin{equation}
@@
 where $\Delta \dot{\theta} =  \left\lvert \dot{\theta}_0-\dot{\theta}_{des} \right\rvert$ is the base velocity variation.
-Fig. \ref{fig:inertia_limits} reports the lower bound of $I_{f}$ given the desired base velocity variation and the maximum velocity of the actuator.
+Figure \ref{fig:inertia_limits} reports the lower bound of $I_{f}$ given the desired base velocity variation and the maximum velocity of the actuator.
 
-\subsection{Flywheels Axes of Rotation}
+\vspace{-9pt}
+
+\begin{figure}[H]%[t]3
+
+	\includegraphics[width=.8\linewidth]{figures/lowerbound2.eps}
+	\caption{\small Minimum inertia of the OCS (normalized by the robot's inertia) necessary to achieve a base velocity variation $\Delta \dot{\theta}$, given the actuation bound $\dot \gamma^{max}$.}
+	\label{fig:inertia_limits}
+\end{figure}
+
+\subsection{Flywheels' Axes of Rotation}
 \label{subseq:flywheel_axes}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/axes.eps}
-	\caption{\small \MF{Top view of Solo12 with the proposed \gls{ocs}.
-The $x-$ and $y-$axes embody the forward and left directions of the robot base, respectively.
-The picture reports also the unit vectors $\hat{\bm{a}}_l$ and $\hat{\bm{a}}_r$  that represent the axes of rotation of the left flywheel and of the right one: $\alpha$ is the angle of incidence defined in \ref{subseq:flywheel_axes}.}}
-	\label{fig:axes}
-\end{figure}
+
 The orientation of the axis of rotation of each flywheel influences their contribution to the total angular momentum.
-To meet the specification of controlling both the robot roll and pitch, the axes of rotation of the left and right wheel, identified in the base reference frame with the unit vectors $\hat{\bm{a}}_l$ and $\hat{\bm{a}}_r$, are set to be incident, laying on a plane parallel to the $xy-$plane of the base reference frame (see Fig.
-\ref{fig:axes}).
-To remove unnecessary complications, we designed two identical modules, to be mounted on the trunk of Solo12, symmetrically with the respect of the robot sagittal plane.
-We denote with $\alpha \leq \pi/2$ the non-negative incident angle between the wheels' axes of rotation and the robot lateral direction.
-The matrix
+To meet the specification of controlling both the robot roll and pitch, the axes of rotation of the left and right wheels, identified in the base reference frame with the unit vectors $\hat{\bm{a}}_l$ and $\hat{\bm{a}}_r$, are set to be incident, laying on a plane parallel to the $xy-$plane of the base reference frame (see Figure \ref{fig:axes}).
+To remove unnecessary complications, \MF{the authors} designed two identical modules, to be mounted on the trunk of Solo12, symmetrically, with respect to the robot's sagittal plane.
+\MF{Let} denote $\alpha \leq \pi/2$ with the non-negative incident angle between the wheels' axes of rotation and the robot's lateral direction.
+The matrix:
 \begin{equation}
@@
 maps the flywheel moments into scalar torques $\bm{u} = \left[\begin{array}{cc} \tau_{fl} & \tau_{fr} \end{array} \right]^T$ about the flywheel axes $\hat{\bm{a}}_l$ and $\hat{\bm{a}}_r$ (expressed in the base frame).
-As long as $\bm{C}$ is full column rank, it is possible to control both roll and pitch angles.
-If $\alpha = 0$, the roll results to be uncontrollable through the \gls{ocs}; otherwise, if $\alpha = \pi/2$, the pitch is uncontrollable.
-The angle is selected considering the ratio of the eigenvalues along the $x$ and $y$ directions of the ellipsoid of the robot inertia(Fig. \ref{fig:ellipsoid}):
+As long as $\bm{C}$ is in full column rank, it is possible to control both roll and pitch angles.
+If $\alpha = 0$, the roll becomes uncontrollable through the OCS; otherwise, if $\alpha = \pi/2$, the pitch becomes uncontrollable.
+The angle is selected considering the ratio of the eigenvalues along the $x$ and $y$ directions of the ellipsoid of the robot's inertia(Figure \ref{fig:ellipsoid}):
 \MF{\begin{equation}
@@
 \end{equation}}
-\begin{figure}[t]
-	\centering
+
+\begin{figure}[H]%[t]4
+
+	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/axes.eps}
+
+	\caption{\MF{Top view of Solo12 with the proposed OCS.
+The $x-$ and $y-$axes embody the forward and left directions of the robot base, respectively.
+The figure also illustrates the unit vectors $\hat{\bm{a}}_l$ and $\hat{\bm{a}}_r$  that represent the axes of rotation of the left flywheel and of the right one and $\alpha$ is the angle of incidence, defined in Section \ref{subseq:flywheel_axes}.}}
+	\label{fig:axes}
+\end{figure}
+\vspace{-11pt}
+\begin{figure}[H]%[t]5
+ \hspace{-16pt}
 	\includegraphics[width=1\linewidth]{figures/inertiaEllipsoid.eps}
-	\caption{\small The inertia tensor can be seen as an ellipsoid. Its principal axes in the direction of the eigenvectors of the tensor and their length depends on the eigenvalues.}
+	\vspace{-9pt}
+	\caption{\hl{The} %MDPI: Please change the hyphen (-) into minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1".
+ inertia tensor can be seen as an ellipsoid. Its principal axes are in the direction of the eigenvectors of the tensor and their lengtha depend on the eigenvalues.}
 	\label{fig:ellipsoid}
 \end{figure}
-In the case of Solo12, this angle is about $40^{\circ}$. With these considerations, the angular momentum produced by the wheels, expressed in the robot base frame, is:
+In the case of Solo12, this angle was about $40^{\circ}$. With these considerations in mind, the angular momentum produced by the wheels, expressed in the robot base frame, was:
 \begin{equation}
@@
 in which $\bm{\omega}_{fl}= \omega_{fl} \, \hat{\bm{a}}_{l}$ and $\bm{\omega}_{fr}= \omega_{fr} \, \hat{\bm{a}}_{r}$ are the angular velocity vectors of the two wheels, and $\omega_{fl}$ and $\omega_{fr}$ are the angular speeds provided to each flywheel by its actuation system.
-The latter equation shows that the difference of the two angular speeds impacts on roll rotations, while their difference can be used to adjust the pitch, see Fig.
-\ref{fig:sum-diff}.
-Using the definition of $\bm{C}$, \eqref{eq:bLf} rewrites as
+The latter equation shows that the difference of the two angular speeds impacts on roll rotations, while their difference can be used to adjust the pitch, see Figure \ref{fig:sum-diff}.
+Using the definition of $\bm{C}$, \eqref{eq:bLf} is rewritten as:
 \begin{equation}
@@
 \end{equation}
-\begin{figure}
-	\centering
+\vspace{-6pt}
+\begin{figure}[H]%6
+
 	\includegraphics[width=.7\linewidth]{figures/sum_diff_mod.eps}
-	\caption{\small Having incident rotation axes, the \gls{ocs} allows to control both the robot roll and pitch.
+	\caption{Having incident rotation axes, the OCS allows control of both the robot's roll and pitch.
 Notice that the roll angle is influenced by the difference of the angular speeds of the flywheels, $\omega_x = \left(\omega_{fl} - \omega_{fr}\right)\sin(\alpha)$.
@@
 \subsection{Inertia Selection}
-Once the desired inertia $I_f$ is chosen, according to the 2D Elroy's Beanie model, it is used to realize the 3D \gls{ocs} (Fig. \ref{fig:flywheel}). 
-We decided to design the flywheels in the shape of hollow cylinders.
+Once the desired inertia $I_f$ is selected, according to the 2D Elroy's Beanie model, it is used to realize the 3D OCS (Figure \ref{fig:flywheel}). 
+\MF{Two identical flywheels were designed to have} the shape of hollow cylinders.
 This shape increases the inertia by locating the mass far away from the rotation axis.
-The inertia tensor expressed in its principal axes, $\bm{I}_f = \mathrm{diag}\left\{ I_{f, xx}, \, I_{f, yy}, \, I_{f, zz}  \right\}$, depends on the cylinders inner and outer radii, $r$ and $R$, its height $h$ and the material density $\rho$:
+The inertia tensor expressed in its principal axes, \linebreak  $\bm{I}_f = \mathrm{diag}\left\{ I_{f, xx}, \, I_{f, yy}, \, I_{f, zz}  \right\}$, depends on the cylinders' inner and outer radii, $r$ and $R$, its height $h$ and material density $\rho$:
 \begin{equation}
@@
 \end{equation}
-Notice that in the Elroy's Beanie model, there is a single body that models the complete \gls{ocs}.
+\hl{Notice} that in the Elroy's Beanie model, there is a single body that models the complete OCS.
 The inertia introduced in the Elroy's Beanie model has to be split between the two flywheels: $I_{f, zz} = I_f/\left(2\cos(\alpha)\right)$.
-The parameters $R$ and $h$ can be set to have a compact \gls{ocs} and $\rho$ depends on the chosen material, that is stainless steel in our case.
-The inner radius $r$ can be adjusted to obtain the desired inertia.
-Spokes with negligible mass connect the wheel to the motor shaft.
+The parameters $R$ and $h$ can be set to have a compact OCS and $\rho$ depends on the chosen material, which was stainless steel in our case.
+The inner radius $r$ could be adjusted to obtain the desired inertia.
+Spokes with negligible mass connected the wheel to the motor shaft.
 All the parameters are reported in Table \ref{tab:dim_fw}, together with the selected inertia and mass of a single flywheel.
-\begin{table}[H] 
-	\caption{Sizes and dynamic parameters of a single flywheel}
+
+\begin{figure}[H]%[t!]7
+
+	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/flywheel_opaque.png}
+	\caption{\small The final design of the flywheel, shown together with its principal axes of inertia.
+Here, $h$ is the wheel thickness, $r$ is the inner radius of the wheel and $R$ is the radius of the outer one.}
+	\label{fig:flywheel}
+\end{figure}
+
+\vspace{-6pt}
+\begin{table}[H]%[t] 1
+	\caption{\hl{Sizes} %MDPI: Please confirm if all the "\cdot" between numbers should be a multiplication sign ("\times" U+00D7).
+ and dynamic parameters of a single flywheel.}
 	\label{tab:dim_fw}
@@
 		\midrule
-		$r$ & $2.20\cdot 10^{-2}$ & $\mathrm{m}$ \\
+		$r$ & $2.20 \cdot 10^{-2}$ & $\mathrm{m}$ \\
 		$R$ & $3.00\cdot 10^{-2}$ & $\mathrm{m}$ \\
@@
 \end{table}
-\begin{figure}[t!]
-	\centering
-	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/flywheel_opaque.png}
-	\caption{\small The final design of the flywheel, shown together with its inertia principal axes. Here, $h$ is the wheel thickness, $r$ is the inner radius of the wheel and $R$ is the outer one.}
-	\label{fig:flywheel}
-\end{figure}
-\unskip
 
-\subsection{Flywheels Control Law}
+\subsection{Flywheels' Control Law}
 \label{subsec:control_law}
-To derive a control law based on the robot base orientation, we make use of \eqref{eq:cons_ang_mom}, expressing all the contributions to the time derivative of total angular momentum with the respect to the base reference frame:
-\MF{\begin{equation}
+To derive a control law based on the robot's base orientation, we made use of \eqref{eq:cons_ang_mom}, expressing all the contributions to the time derivative of total angular momentum with respect to the base reference frame:
+\begin{equation}
 \tensor[_b]{\bm{I}}{_f} \, \tensor[_b]{\dot{\bm{\omega}}}{_f} + \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\dot{\bm{\omega}}}{_r} + \tensor[_b]{\bm{\omega}}{_r} \times \left( \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\bm{\omega}}{_r}\right) = 0
-\end{equation}}
-From this expression, we can define the moment on the base caused by the acceleration of the flywheels, that can be used as a feedback torque $\bm{\tau}_{fb}$:
-\MF{\begin{equation}
+\end{equation}
+\hl{From} this expression, the moment on the base caused by the acceleration of the flywheels \MF{is defined}, and can be used as a feedback torque $\bm{\tau}_{fb}$:
+\begin{equation}
 \begin{split}
@@
 \end{split}
-\end{equation}}
-We can use the \MF{proportional and derivative} action $\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}_r}$, where $\bm{K}_p$ and $\bm{K}_d$ are diagonal and positive-definite gain matrices for the error in attitude and angular velocity.
-The orientation error $\bm{e} \in SO(3)$ needs the algebra of the special rotational group to be computed. To avoid singular configurations, we represent the orientation with quaternions.
-The derivative error can be computed using $\dot{\bm{e}} = \tensor[_b]{\bm{\omega}}{_r^{des}} - \tensor[_b]{\bm{\omega}}{_r}$, in which $\tensor[_b]{\bm{\omega}}{_r^{des}}$ and $\tensor[_b]{\bm{\omega}}{_r}$ are, respectively, the desired and actual angular velocity of the base:
-\MF{\begin{equation}
+\end{equation}
+\hl{Proportional} and derivative action $\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}_r}$ \MF{is used}, where $\bm{K}_p$ and $\bm{K}_d$ are diagonal and positive definite gain matrices for the error in attitude and angular velocity.
+The orientation error $\bm{e} \in SO(3)$ requires that the algebra of the special rotational group is computed. To avoid singular configurations, orientations \MF{are represented} with quaternions.
+The derivative error can be computed using $\dot{\bm{e}} = \tensor[_b]{\bm{\omega}}{_r^{des}} - \tensor[_b]{\bm{\omega}}{_r}$, in which $\tensor[_b]{\bm{\omega}}{_r^{des}}$ and $\tensor[_b]{\bm{\omega}}{_r}$ are, respectively, the desired and actual angular velocities of the base:
+\begin{equation}
 \bm{\tau}_{fb} = - \tensor[_b]{\bm{I}}{_r} \left(\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}}\right) - \tensor[_b]{\bm{\omega}}{_r} \times \left( \tensor[_b]{\bm{I}}{_r} \tensor[_b]{\bm{\omega}}{_r}\right).
-\end{equation}}
-Projecting the moment $\bm{\tau}_{fb}$ onto the flywheel axes with $\bm{C}^T$, we obtain the control action $\bm{u}$
+\end{equation}
+\hl{Projecting} the moment $\bm{\tau}_{fb}$ onto the flywheel axes with $\bm{C}^T$, the control action $\bm{u}$ \MF{yields}
 \begin{equation}
@@
 \label{sec:simulations}
-To validate our \gls{ocs}, we present three simulations on different scenarios. We want to test the capability of the proposed approach to: 
-reject a disturbance when the robot is in the flight phase of a jump, damp trunk oscillations after the touch-down, 
-and work in parallel with the joints of the legs to achieve a highly dynamic motion. 
-All the simulations are performed using \MF{ROS \cite{quigley2009ros} together with Gazebo \cite{koenig2004design}.
-Physics related parameters are reported in Table \ref{tab:pys_param}.} The robot full dynamics is modeled with Pinocchio \cite{carpentier2019pinocchio}.
-The references for the joints of the legs are computed off-line using Crocoddyl \cite{mastalli2020crocoddyl} and tracked with a proportional-derivative joint controller.
-The \gls{ocs} is commanded to  track base orientation references using the control law introduced in Subsection \ref{subsec:control_law} \MF{coded in Python}.\footnote{The video with all the simulations is available at web link \url{https://www.dropbox.com/scl/fo/wevvl0zvfp8gofie6x7rw/h?dl=0&rlkey=pap76cr813a0skrvrxl7qgrb9}.}
+To validate our OCS, three simulations on different scenarios {are reported.~The authors} wanted to test the capability of the proposed approach in terms of the following: 
+to reject a disturbance when the robot was in the flight phase of a jump, and to dampen trunk oscillations after touch-down, 
+and to work in parallel with the joints of the legs to achieve a highly dynamic motion. 
+All the simulations were performed using \MF{ROS \cite{quigley2009ros}, together with Gazebo \cite{koenig2004design}.
+The physics-related parameters are reported in Table \ref{tab:pys_param}.} References for the joints of the legs were computed off-line using Crocoddyl, an optimal control library for robots based on Differential Dynamic Programming (DDP) algorithms.
+It uses Pinocchio for fast computation of robot dynamics and the analytical derivatives.
+References $\bm{q}_{ref}$, $\dot{\bm{q}}_{ref}$ and $\bm{\tau}_{ref}$ for joint positions, velocities and torques were then executed on-line with a proportional-derivative joint controller:
+\begin{equation}
+\bm{\tau}_j = \bm{K}_{p,\, j} (\bm{q}_{ref} - \bm{q}) + \bm{K}_{d,\, j} (\dot{\bm{q}}_{ref} - \dot{\bm{q}}) + \bm{\tau}_{ref}
+\end{equation}
+\hl{The} OCS was commanded to track base orientation references using the control law introduced in Section \ref{subsec:control_law} \MF{coded in Python}.
+({\hl{The} %MDPI: Footnote is not permitted in our journal.
+We include this paragraph to the maintext.
+Please confirm.
+ video, with all the simulations, is available at web link \url{https://www.dropbox.com/scl/fo/wevvl0zvfp8gofie6x7rw/h?dl=0&rlkey=pap76cr813a0skrvrxl7qgrb9}  (\hl{accessed} %MDPI: Please add the access date (Format: Date Month Year).
+e.g., (accessed on 1 January 2020).
+ on).)}
 
-\begin{table}[t!] 
-	\caption{\MF{Physics related parameters used for simulate the robot dynamics in Gazebo}}
+\begin{table}[H]%[t!] 
+	\caption{\MF{Physics-related parameters used to simulate the robot dynamics in Gazebo.}}
 	\label{tab:pys_param}
@@
 \end{table}
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.7\linewidth]{figures/drawing.png}
-	\caption{\small First Test: simulation results showing the $\mathrm{CoM}_x$ and pitch trajectory vs.
-time plots.
-A disturbance moment $\bm{\tau}_{dist}$ on the trunk can be compensated only if the \gls{ocs} is enabled.
-If it is disabled, the robot is unable to restore a safe configuration after the touch-down and eventually falls.
-The light blue area represents the interval of time in which the disturbance is applied.}
-	\label{fig:fw_jump}
-\end{figure}
 
-\begin{figure}[t!]
-	\centering
-	\includegraphics[width=.7\linewidth]{figures/fall_plot.eps}
-	\caption{\small Second Test: simulation results showing the $\mathrm{CoM}_x$ and pitch trajectory vs.
-time plots.
-The \gls{ocs} drives the robot orientation during a fall to be horizontal.
-This allows to damp for damping the base oscillations after the touch-down (vertical dashed line), even without implementing a landing strategy}
-	\label{fig:fall_plot}
-\end{figure}
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=\linewidth]{figures/backflip.png}
-	\caption{\small \MF{Third Test: sequence of snapshots of Solo12 performing a back-flip in Gazebo environment.
-The \gls{ocs} alleviates the effort requested to the leg joints, that in this case can be used to accomplish only the vertical motion.
-The red, green and blue lines represent the axes of the inertial (world) reference frame.}}
-	\label{fig:backflip}
-\end{figure}
-
-%In a second test, we demonstrate the versatility use the presented \gls{ocs} to dump the base oscillations generated at the landing. 
+%In a second test, we demonstrate the versatility use the presented OCS to dump the base oscillations generated at the landing. 
 %% , even though we designed it to control the orientation during the flight phase.
@@
 is the same jump?}
-%The results point out the possibility to use this device even when the angular momentum is not conserved (i.e. when contacts are present).
+%The results point out the possibility to use this device even when the angular momentum is not conserved (i.e., when contacts are present).
 %this is the main test! 
-The necessity of having an \gls{ocs} is revealed with the first test: disturbance that occurs when the robot has no contact with the ground. 
-During the flight phase of a forward jump, $0.1~\mathrm{s}$ after the lift-off, an external disturbance moment is applied to the robot base, deflecting the robot orientation: this disturbance is set to $\bm{\tau}_{dist} = \left[\begin{array}{ccc}
+\textls[-15]{The necessity of having an OCS was revealed in the first test, as disturbance occurred when the robot had no contact with the ground. 
+During the flight phase of a forward jump, $0.1~\mathrm{s}$ after the lift-off, an external disturbance moment was applied to the robot's base, deflecting the robot's orientation.
+This disturbance was set to $\bm{\tau}_{dist} = \left[\begin{array}{ccc}
 0.2 & 0.8 & 0.0
@@
 \right]^T \ \mathrm{N \cdot m}$ and applied for $0.05 \ \mathrm{s}$. 
-If the flywheels are not used, Solo12 falls after the touch-down. Instead, 
+If the flywheels were not used, Solo12 fell after touch-down. Instead, when
 using the flywheels %pairing the legs joint PD controller with \ref{eq:ctrl_law},
-it is possible to drive the robot to a safe configuration after the landing (Fig. \ref{fig:fw_jump}), without need to implement of a specific landing strategy, as the one in \cite{jeon2021real}.
+it was possible to drive the robot to a safe configuration after landing (Figure \ref{fig:fw_jump}), without the need to implement a specific landing strategy, such as the one in \cite{jeon2021real}.
+}
 
-In the second test, we demonstrate the ability of the \gls{ocs} to reorient the base.
-We let the robot fall from a height of $1 \ \mathrm{m}$ with an initial pitch orientation of $30^\circ$ and zero base angular velocity.
-If we do not actuate the flywheels, the robot touches the ground with the same initial orientation and the trunk oscillates.
-If the \gls{ocs} is enabled, it drives the robot base to be horizontal when it is still in the air and the oscillations after the touch-down are drastically reduced, both in the pitch angle and forward direction (Fig.
-\ref{fig:fall_plot}).
+\MF{The second test demonstrated} the ability of the OCS to reorient the base.
+\MF{The robot fell} from a height of $1.0 \ \mathrm{m}$ with an initial pitch orientation of $30^\circ$ and zero base angular velocity.
+\MF{If the flywheels were not actuated}, the robot touched the ground with the same initial orientation and the trunk oscillated.
+If the OCS was enabled, it drove the robot's base to be horizontal when it was still in the air and the oscillations after touch-down were drastically reduced, both in the pitch angle and forward direction (Figure \ref{fig:fall_plot}).
 
-Finally, we demonstrate how flywheels can alleviate the effort of the leg joints to achieve a highly dynamic motion: a back-flip (Fig. \ref{fig:backflip}). 
-For this, we target a space application carrying out a simulation with \MF{lunar gravity acceleration} ($1.62 \ \mathrm{m/s^2}$).
-In this way, it is possible to obtain high jumps with a long flight phase, without having to select more powerful actuators.  
-The leg joint trajectory computed off-line describes a purely vertical jump of $1 \ \mathrm{m}$, having a flight phase that lasts $2 \ \mathrm{s}$.
-Right after the lift-off, the flywheels start the reorientation task of performing a back-flip, that is a spin of $360^\circ$ on the pitch.
-For this maneuver, the value of the incident angle $\alpha$ is set to $0^\circ$ since no roll rotation is required.
-% and it is preferred to use all the contributions of the \gls{ocs} on the pitch.
-Our hardware design allows the manual change of this value before performing the task.
-This simulation demonstrates that the \gls{ocs} alleviate the effort to be applied on the legs: indeed, that it would not have been possible to do a back-flip without the flywheels because the legs are used only to achieve the linear motion.
+\MF{Finally, the ability of the flywheels to relieve the efforts of the leg joints to achieve a highly dynamic movement was shown with a back-flip} (Figure \ref{fig:backflip}). 
+For this, we targeted a space application, carrying out a simulation with \MF{lunar gravity acceleration} ($1.62 \ \mathrm{m/s^2}$).
+In this way, it was possible to obtain high jumps with a long flight phase, without having to select more powerful actuators.  
+The leg joint trajectory computed off-line described a purely vertical jump of $1 \ \mathrm{m}$, having a flight phase that lasted $2 \ \mathrm{s}$.
+Right after the lift-off, the flywheels started the reorientation task of performing a back-flip, by means of a spin of $360^\circ$ on the pitch.
+For this maneuver, the value of the incident angle $\alpha$ was set to $0^\circ$, since no roll rotation was required.
+% and it is preferred to use all the contributions of the OCS on the pitch.
+Our hardware design allowed manual change of this value before performing the task.
+This simulation demonstrated that the OCS alleviated the effort applied on the legs.
+Indeed, it would not have been possible to do a back-flip without the flywheels, because the legs were only used to achieve the linear motion.
+%\vspace{-12pt}
+\begin{figure}[H]%[t]8
+
+	\includegraphics[width=.7\linewidth]{figures/drawing.png}
+	\caption{\small First Test: simulation results showing the $\mathrm{CoM}_x$ and pitch trajectory vs.
+time plots.
+A disturbance moment $\bm{\tau}_{dist}$ on the trunk could be compensated for only if the OCS was enabled.
+If it was disabled, the robot was unable to restore a safe configuration after touch-down and, eventually, fell.
+The light blue area represents the interval of time in which the disturbance was applied.}
+	\label{fig:fw_jump}
+\end{figure}
+\vspace{-8pt}
+\begin{figure}[H]%[t!]9
+
+	\includegraphics[width=.7\linewidth]{figures/fall_plot.eps}
+	\caption{Second Test: simulation results showing the $\mathrm{CoM}_x$ and pitch trajectory vs.
+time plots.
+The OCS drove the robot's orientation during a fall so as to be horizontal.
+This allowed dampening of the base oscillations after touch-down (vertical dashed line), even without implementing a l\mbox{anding strat}egy.}
+	\label{fig:fall_plot}
+\end{figure}
+
+\begin{figure}[H]%[h!]10
+
+	\includegraphics[width=\linewidth]{figures/backflip.png}
+	\caption{{Third Test: sequence of snapshots of Solo12 performing a back-flip in the Gazebo environment.
+The OCS alleviated the effort requested of the leg joints, that, in this case, could only be used to accomplish the vertical motion.
+The red, green and blue lines represent the axes of the inertial (world) reference frame.}}
+	\label{fig:backflip}
+\end{figure}
 
@@
 
-In this work, we have presented the design of a novel \gls{ocs}, composed of two flywheels, that enables to control the trunk of a legged robot platform, 
-increasing the accuracy of aerial maneuvers during the under-actuated phases (i.e. flight phase) as well as 
-enhancing stability when in contact by damping oscillations.
-The novelty of the design, that involves the flywheels to be attached with incident rotational axes on the trunk, allows to control the orientation
-in both the roll and pitch directions while keeping the device compact. Their effect is optimized considering the inertial property of the mechanical structure. 
-Several simulations are reported with the quadruped Solo12 that demonstrate the effectiveness of the proposed approach to: 
-reject disturbances during the flight phase, stabilize the platform after the touch-down even in absence of a specific landing strategy and 
-achieve a fast reorientation maneuver (a back-flip) in a reduced gravity environment.
+\MF{This work presented} the design of a novel OCS, composed of two flywheels, that enable control of the trunk of a legged robot's platform, increasing the accuracy of aerial maneuvers during the under-actuated phases (i.e., flight phase), as well as enhancing stability when in contact by damping the oscillations.
+The novelty of the design, which involves flywheels attached with incident rotational axes on the trunk, is that it allows control of the orientations in both the roll and pitch directions, while keeping the device compact.
+The effect is optimized considering the inertial property of the mechanical structure.
+Several simulations were performed with the quadruped Solo12 that demonstrated the effectiveness of the proposed approach in the following: to reject disturbances during the flight phase, to stabilize the platform after touch-down, even in absence of a specific landing strategy, and to achieve a fast reorientation maneuver (a back-flip) in a reduced gravity environment.
 
-In future works, we plan to demonstrate our concept performing experiment with the real platform.
-The control strategy for defining the base desired angular velocity can be improved using \gls{nmpc}, that takes into consideration the future samples of the orientation reference.
-We expect this feature will allows the stabilization of the yaw to a desired value, that if it is not locally controllable, enhancing the non-holonomy property of the angular momentum, i.e.
-doing a preliminary roll and pitch maneuver.
+\MF{Possible applications of the presented OCS include, but are not limited to, efficient adjustment of the posture of quadruped robots walking or jumping on uneven terrains.
+As proved in the third simulation (back-flip), our approach improves the capabilities of quadrupeds in space application, enabling fast locomotion by means of leaps, ensuring a reactive control action on the robot's angular momentum.
+Furthermore, the method presented in Section \ref{sec:ocs} in designing the OCS does not depend on a specific platform, and, thus, it can be replicated for reorienting mechanical structures with different morphologies, e.g, monopods or bipeds.
+A construction worker's backpack could contain two flywheels with incident rotation axes, which, in the event of a fall from scaffolding, could be used to reorient a human body for impact on the ground with an upright posture, using the same controller proposed in this work.}
+
+In future works, \MF{the authors} plan to demonstrate our concept performing experiments with a real platform.
+The control strategy for defining the base's desired angular velocity could be improved using  Nonlinear Model Predictive Control (NMPC), which takes into consideration future samples of the orientation reference.
+\MF{This feature is expected to allow} the stabilization of the yaw to a desired value, so that, if it is not locally controllable, enhances the non-holonomy property of the angular momentum, i.e., doing a preliminary roll and pitch maneuver.
 A control strategy that may lead to the same result is dynamic feedback linearization, widely used for flight with quadricopters.
@@
 % If you wish to submit a video article, please do so with any other supplementary material.
-% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
-A supporting video article is available at doi: link.}
+ \supplementary{\hl{The} %MDPI: Please add supplementary material citation in main  text, e.g., (see Supplementary Material)
+ following supporting information can be downloaded at: \linksupplementary{s1},  \hl{Video S1: title.}%MDPI: please complete it
+}
 
@@
 and A.C.; supervision, A.D.P.
-and M.F; project administration, F.R.
-and M.F.; funding acquisition, M.F., C.S.
+and M.F.; project administration, F.R.
+and M.F.; funding acquisition, M.F.
+and C.S.
 All authors have read and agreed to the published version of the manuscript.}
@@
 
-\funding{The publication was created with the co-financing of the European Union FSE-REACT-EU, PON Research and Innovation 2014-2020 DM1062 / 2021.}
+\funding{The publication was created with the co-financing of the European Union FSE-REACT-EU, PON Research and Innovation 2014-2020 DM1062/2021.}
 
@@
 
-\acknowledgments{All the authors want to thank Dr. Roy Featherstone for interesting tips that started this work. }
+\acknowledgments{All the authors want to thank Roy Featherstone for interesting tips that started this work. }
 
@@
 DoF & Degree of Freedom\\
-CoM & Center Of Mass\\
+CoM & Center of Mass\\
 OCS & Orientation Control System\\
 CMG & Control Moment Gyroscope\\
+\MF{DDP} & \MF{Differential Dynamic Programming}\\
 NMPC & Nonlinear Model Predictive Control
@@
 
-%\begin{table}[H] 
+%\begin{table}[H]%[H] 
 %\caption{This is a table caption.\label{tab5}}
@@
 
-% Please provide either the correct journal abbreviation (e.g.
-according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
+% Please provide either the correct journal abbreviation (e.g., according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
 % Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
@@
 %=====================================
-\bibliography{references/bibliography.bib}
+\begin{thebibliography}{999}
+
+\bibitem[Barasuol \em{et~al.}(2013)Barasuol, Buchli, Semini, Frigerio,
+  De~Pieri, and Caldwell]{barasuol2013reactive}
+Barasuol, V.; Buchli, J.; Semini, C.; Frigerio, M.; De~Pieri, E.R.; Caldwell,
+  D.G.
+\newblock A reactive controller framework for quadrupedal locomotion on
+  challenging terrain.
+\newblock In Proceedings of the 2013 IEEE International Conference on Robotics
+  and Automation, \hl{Karlsruhe, Germany, 6--10 May 2013;} %MDPI: Newly added information. Please confirm. The following highlights are the same.
+ IEEE: \hl{Piscataway, NJ, USA,}  2013; pp.~2554--2561.
+
+\bibitem[Focchi \em{et~al.}(2020)Focchi, Orsolino, Camurri, Barasuol, Mastalli,
+  Caldwell, and Semini]{focchi2020heuristic}
+Focchi, M.; Orsolino, R.; Camurri, M.; Barasuol, V.; Mastalli, C.; Caldwell,
+  D.G.; Semini, C.
+\newblock Heuristic planning for rough terrain locomotion in presence of
+  external disturbances and variable perception quality. In {\em Advances in
+  Robotics Research: From Lab to Market}; Springer: \hl{Berlin/Heidelberg, Germany,} %We added the location of publisher. Please confirm
+  2020; pp. 165--209.
+
+\bibitem[Kane and Scher(1969)]{kane1969dynamical}
+Kane, T.; Scher, M.
+\newblock A dynamical explanation of the falling cat phenomenon.
+\newblock {\em Int. J. Solids Struct.} {\bf 1969}, {\em
+  5},~663--670.
+
+\bibitem[Chu \em{et~al.}(2019)Chu, Lo, Ma, and Au]{chu2019null}
+Chu, X.; Lo, C.H.D.; Ma, C.; Au, K.W.S.
+\newblock Null-Space-Avoidance-Based Orientation Control Framework for
+  Underactuated, Tail-Inspired Robotic Systems in Flight Phase.
+\newblock {\em IEEE Robot. Autom. Lett.} {\bf 2019}, {\em
+  4},~3916--3923.
+
+\bibitem[Wenger \em{et~al.}(2016)Wenger, De, and Koditschek]{wenger2016frontal}
+Wenger, G.; De, A.; Koditschek, D.E.
+\newblock Frontal plane stabilization and hopping with a 2DOF tail.~In Proceedings of the 2016 IEEE/RSJ International Conference on
+  Intelligent Robots and Systems (IROS), \hl{Daejeon, Korea, 9--14 October 2016}; IEEE: \hl{\mbox{Piscataway}, NJ, USA,}  2016; pp.~567--573.
+
+\bibitem[Johnson \em{et~al.}(2012)Johnson, Libby, Chang-Siu, Tomizuka, Full,
+  and Koditschek]{johnson2012tail}
+Johnson, A.M.; Libby, T.; Chang-Siu, E.; Tomizuka, M.; Full, R.J.; Koditschek,
+  D.E.
+\newblock Tail assisted dynamic self righting. In {\em Adaptive Mobile
+  Robotics}; World Scientific: \hl{Singapore},
+  2012; pp. 611--620.
+
+\bibitem[An \em{et~al.}(2022)An, Ma, Lo, Ng, Chu, and Au]{an2022design}
+An, J.; Ma, X.; Lo, C.H.D.; Ng, W.; Chu, X.; Au, K.W.S.
+\newblock Design and Experimental Validation of a Monopod Robot with 3-DoF
+  Morphable Inertial Tail for Somersault.
+\newblock {\em IEEE/ASME Trans. Mechatron.} {\bf 2022}, \hl{\emph{7}, 5072--5083}.
+
+\bibitem[An \em{et~al.}(2020)An, Chung, Lo, Ma, Chu, and Au]{an2020development}
+An, J.; Chung, T.; Lo, C.H.D.; Ma, C.; Chu, X.; Au, K.S.
+\newblock Development of a bipedal hopping robot with morphable inertial tail
+  for agile locomotion.
+\newblock In Proceedings of the 2020 8th IEEE RAS/EMBS International Conference
+  for Biomedical Robotics and Biomechatronics (BioRob),  \hl{New York City, NY, USA, 29 November--1 December 2020}; IEEE: \hl{Piscataway, NJ, USA},  2020; \mbox{pp.
+  132--139.}
+
+\bibitem[Tang \em{et~al.}(2022)Tang, An, Chu, Wang, Wong, and
+  Au]{tang2022towards}
+Tang, Y.; An, J.; Chu, X.; Wang, S.; Wong, C.Y.; Au, K.
+\newblock Towards Safe Landing of Falling Quadruped Robots Using a 3-DoF
+  Morphable Inertial Tail.
+\newblock {\em arXiv} {\bf 2022},  arXiv:2209.15337.
+
+\bibitem[Hoffman and Paolillo(2021)]{hoffman2021exploiting}
+Hoffman, E.M.; Paolillo, A.
+\newblock Exploiting visual servoing and centroidal momentum for whole-body
+  motion control of humanoid robots in absence of contacts and gravity.
+\newblock In Proceedings of the 2021 IEEE International Conference on Robotics
+  and Automation (ICRA), \hl{Xi'an, China, 30 May--5 June 2021}; IEEE: \hl{Piscataway, NJ, USA},  2021; pp.~2979--2985.
+
+\bibitem[Kurtz \em{et~al.}(2022)Kurtz, Li, Wensing, and Lin]{kurtz2022mini}
+Kurtz, V.; Li, H.; Wensing, P.M.; Lin, H.
+\newblock Mini cheetah, the falling cat: A case study in machine learning and
+  trajectory optimization for robot acrobatics.
+\newblock In Proceedings of the 2022 International Conference on Robotics and
+  Automation (ICRA), \hl{Philadelphia , PA, USA, 23--27 May 2022}; IEEE: \hl{Piscataway, NJ, USA},  2022; pp.~4635--4641.
+
+\bibitem[Yoon and Tsiotras(2002)]{yoon2002spacecraft}
+Yoon, H.; Tsiotras, P.
+\newblock Spacecraft adaptive attitude and power tracking with variable speed
+  control moment gyroscopes.
+\newblock {\em J. Guid. Control. Dyn.} {\bf 2002}, {\em
+  25},~1081--1090.
+
+\bibitem[Brown and Xu(1996)]{brown1996single}
+Brown, H.B.; Xu, Y.
+\newblock A single-wheel, gyroscopically stabilized robot.
+\newblock In Proceedings of the  IEEE International Conference on
+  Robotics and Automation, \hl{Minneapolis, MN, USA, 22--28 April 1996}; IEEE: \hl{Piscataway, NJ, USA},  1996, Vol.~4; pp.~3658--3663.
+
+\bibitem[Mikhalkov \em{et~al.}(2021)Mikhalkov, Prutskiy, Sechenev, Kazakov,
+  Simulin, Sokolov, and Ryadchikov]{mikhalkov2021gyrubot}
+Mikhalkov, N.; Prutskiy, A.; Sechenev, S.; Kazakov, D.; Simulin, A.; Sokolov,
+  D.; Ryadchikov, I.
+\newblock Gyrubot: Nonanthropomorphic stabilization for a biped.
+\newblock In Proceedings of the 2021 IEEE International Conference on Robotics
+  and Automation (ICRA), \hl{30 May--5 June 2021, Xi'an, China}; IEEE: \hl{Piscataway, NJ, USA},  2021; pp.~4976--4982.
+
+\bibitem[Oland and Schlanbusch(2009)]{oland2009reaction}
+Oland, E.; Schlanbusch, R.
+\newblock Reaction wheel design for cubesats.
+\newblock In Proceedings of the 2009 4th International Conference on Recent
+  Advances in Space Technologies, \hl{Istanbul, Turkey, 11--13June 2009}; IEEE: \hl{Piscataway, NJ, USA},  2009; pp.~778--783.
+
+\bibitem[Brown and Schmiedeler(2016)]{Brown2016}
+Brown, T.L.; Schmiedeler, J.P.
+\newblock Reaction Wheel Actuation for Improving Planar Biped Walking
+  Efficiency.
+\newblock {\em IEEE Trans. Robot.} {\bf 2016}, {\em 32},~1290--1297.
+\newblock {{https://doi.org/10.1109/TRO.2016.2593484}}.
+
+\bibitem[Xiong and Ames(2020)]{xiong2020sequential}
+Xiong, X.; Ames, A.D.
+\newblock Sequential motion planning for bipedal somersault via flywheel slip
+  and momentum transmission with task space control.
+\newblock In Proceedings of the 2020 IEEE/RSJ International Conference on
+  Intelligent Robots and Systems (IROS), \hl{Las~Vegas, Nevada, USA, 25--29 October 2020}; IEEE: \hl{Piscataway, NJ, USA},  2020; pp.~3510--3517.
+
+\bibitem[Kolvenbach \em{et~al.}(2019)Kolvenbach, Hampp, Barton, Zenkl, and
+  Hutter]{kolvenbach2019towards}
+Kolvenbach, H.; Hampp, E.; Barton, P.; Zenkl, R.; Hutter, M.
+\newblock Towards jumping locomotion for quadruped robots on the moon.
+\newblock In Proceedings of the 2019 IEEE/RSJ International Conference on
+  Intelligent Robots and Systems (IROS), \hl{Macau, 3--8 November 2019}; IEEE: \hl{Piscataway, NJ, USA},  2019; pp.~5459--5466.
+
+\bibitem[Vasilopoulos \em{et~al.}(2016)Vasilopoulos, Machairas, and
+  Papadopoulos]{vasilopoulos2016quadruped}
+Vasilopoulos, V.; Machairas, K.; Papadopoulos, E.
+\newblock Quadruped pronking on compliant terrains using a reaction wheel.
+\newblock In Proceedings of the 2016 IEEE International Conference on Robotics
+  and Automation (ICRA), \hl{Stockholm, Sweden, 16--21 May 2016;} IEEE: \hl{Piscataway, NJ, USA},  2016; pp.~3590--3595.
+
+\bibitem[Machairas and Papadopoulos(2015)]{machairas2015quadruped}
+Machairas, K.; Papadopoulos, E.
+\newblock On quadruped attitude dynamics and control using reaction wheels and
+  tails.
+\newblock In Proceedings of the 2015 European Control Conference (ECC), \hl{Linz, Austria, 15--17 July 2015}; IEEE: \hl{Piscataway, NJ, USA},
+  2015; pp.~753--758.
+
+\bibitem[Gajamohan \em{et~al.}(2012)Gajamohan, Merz, Thommen, and
+  D'Andrea]{gajamohan2012cubli}
+Gajamohan, M.; Merz, M.; Thommen, I.; D'Andrea, R.
+\newblock The cubli: A cube that can jump up and balance.
+\newblock In Proceedings of the 2012 IEEE/RSJ International Conference on
+  Intelligent Robots and Systems, \hl{Vilamoura-Algarve, Portugal, 7--12 October 2012}; IEEE: \hl{Piscataway, NJ, USA},  2012; pp.~3722--3727.
+
+\bibitem[Grimminger \em{et~al.}(2020)Grimminger, Meduri, Khadiv, Viereck,
+  W{\"u}thrich, Naveau, Berenz, Heim, Widmaier, Flayols,
+  et~al.]{grimminger2020open}
+Grimminger, F.; Meduri, A.; Khadiv, M.; Viereck, J.; W{\"u}thrich, M.; Naveau,
+  M.; Berenz, V.; Heim, S.; Widmaier, F.; Flayols, T.;  et~al.
+\newblock An open torque-controlled modular robot architecture for legged
+  locomotion research.
+\newblock {\em IEEE Robot. Autom. Lett.} {\bf 2020}, {\em
+  5},~3650--3657.
+
+\bibitem[{Open Dynamic Robot Initiative (ODRI)}()]{github_solo}
+{Open Dynamic Robot Initiative (ODRI)}.
+\newblock \emph{Open Robot Actuator Hardware}\hl{; .} %MDPI: Please add the name of the publisher and the location of it.
+
+\bibitem[Wieber \em{et~al.}(2016)Wieber, Tedrake, and
+  Kuindersma]{wieber2016modeling}
+Wieber, P.B.; Tedrake, R.; Kuindersma, S.
+\newblock Modeling and control of legged robots. In {\em Springer Handbook of
+  Robotics}; Springer: \hl{Berlin/Heidelberg, Germany},  2016; pp. 1203--1234.
+
+\bibitem[Quigley \em{et~al.}(2009)Quigley, Conley, Gerkey, Faust, Foote, Leibs,
+  Wheeler, Ng, et~al.]{quigley2009ros}
+Quigley, M.; Conley, K.; Gerkey, B.; Faust, J.; Foote, T.; Leibs, J.; Wheeler,
+  R.; Ng, A.Y. 
+\newblock ROS: An Open-Source Robot Operating System.
+\newblock In Proceedings of the ICRA Workshop Open Source Softw, Kobe, Japan,  \hl{12--17 May} 2009; Volume 5, p.~5.
+
+\bibitem[Koenig and Howard(2004)]{koenig2004design}
+Koenig, N.; Howard, A.
+\newblock Design and use paradigms for gazebo, an open-source multi-robot
+  simulator.
+\newblock In Proceedings of the 2004 IEEE/RSJ International Conference on
+  Intelligent Robots and Systems (IROS) (IEEE Cat. No. 04CH37566), \hl{Sendai, Japan, 28 September 2004--2 October 2004}; IEEE: \hl{Piscataway, NJ, USA},  2004;
+  Volume~3; pp.~2149--2154.
+
+\bibitem[Jeon \em{et~al.}(2021)Jeon, Kim, and Kim]{jeon2021real}
+Jeon, S.H.; Kim, S.; Kim, D.
+\newblock Real-time Optimal Landing Control of the MIT Mini Cheetah.
+\newblock {\em arXiv} {\bf 2021},  arXiv:2110.02799.
+
+\end{thebibliography}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\PublishersNote{}
 \end{adjustwidth}

### 2023-01-18 sensors-2114505.tex
for final review

--- 2023-01-18 mdpi2022roscia_v7/sensors-2114505.tex
@@
 % Authors, for the paper (add full first names)
-\Author{\hl{Francesco} %MDPI: Please carefully check the accuracy of names and affiliations.
+\Author{Francesco %MDPI: Please carefully check the accuracy of names and affiliations.
  Roscia $^{1}$, Andrea Cumerlotti $^{1,2}$, Andrea Del Prete $^{2}$, Claudio Semini $^{1}$ and Michele Focchi $^{1,3,}$*}
@@
 \address{%
-$^{1}$ \quad Dynamic Legged Systems (DLS) Lab, Istituto Italiano di Tecnologia (IIT), \hl{Genoa,} %MDPI: Please add the postal code (or ZIP code in the U.S.).
+$^{1}$ \quad Dynamic Legged Systems (DLS) Lab, Istituto Italiano di Tecnologia (IIT), 16163 Genova, %MDPI: Please add the postal code (or ZIP code in the U.S.).
 If the postal code is not available, please provide the P.O.
@@
 The following highlights are the same.
- Italy; \linebreak  \hl{francesco.roscia@iit.it} %MDPI: We added all the emails in affiliations according to the submitting system. Please confirm.
- (F.R.); andrea.cumerlotti@studenti.unitn.it (A.C.); claudio.semini@iit.it (C.S.)\\
-$^{2}$ \quad Industrial Engineering Department (DII), University of Trento, \hl{Trento}, Italy; andrea.delprete@unitn.it (A.D.P.)\\
-$^{3}$ \quad Department of Information Engineering and Computer Science (DISI), University of Trento, \hl{Trento}, Italy\\
+ Italy; \linebreak francesco.roscia@iit.it %MDPI: We added all the emails in affiliations according to the submitting system. Please confirm.
+ (F.R.); claudio.semini@iit.it (C.S.)\\
+$^{2}$ \quad Industrial Engineering Department (DII), University of Trento, 38123 Trento, Italy; andrea.cumerlotti@studenti.unitn.it (A.C.); andrea.delprete@unitn.it (A.D.P.)\\
+$^{3}$ \quad Department of Information Engineering and Computer Science (DISI), University of Trento, 38123 Trento, Italy\\
 }
@@
 % Contact information of the corresponding author
-\corres{Correspondence: michele.focchi@unitn.it}
+\corres{Correspondence: michele.focchi@unitn.it (M.F.)}
 
@@
 	\item the design of a novel OCS that enables effective control of the orientation of a legged robot during a jump, while keeping the design simple. 
-	In particular, the axes of rotation of the flywheels are set to be incident, enabling \hl{\textit{continuous}} %MDPI: Please confirm if all the italics in the maintext should be retained.
+	In particular, the axes of rotation of the flywheels are set to be incident, enabling \textit{continuous} %MDPI: Please confirm if all the italics in the maintext should be retained.
  controllability in both directions of roll and pitch, while keeping the device compact.
@@
 \begin{figure}[H]%[t!] 1
-
 	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/solo_fw_pic_compressed.png}
@@
 \label{sec:background}
-The starting point for any OCS is the Euler's equation.
-For any mechanical system, the time derivative of the angular \hl{momentum}, %MDPI: Please confirm if all the bold in equations should be retained.
+The starting point for any OCS is the Euler's equation. For any mechanical system, the time derivative of the angular momentum, %MDPI: Please confirm if all the bold in equations should be retained.
  $\bm{L}$ computed in respect to a reference point $O$, fixed in an inertial frame, equals the sum of the moments $\bm{M}_i$ applied to the system with respect to the same reference point:
@@
 \end{equation}
-\hl{When} %MDPI: Please confirm if the noindent should be retained. The following highlights are the same.
+When %MDPI: Please confirm if the noindent should be retained. The following highlights are the same.
  the result of the external moments applied to the system is zero, and the Euler's equation simplifies to:
@@
 \end{figure} 
-
-This consists of two rigid bodies connected through their CoMs with a revolute joint.
+\noindent This consists of two rigid bodies connected through their CoMs with a revolute joint.
 One of the bodies represents the robot in its nominal configuration and the other models the OCS, here represented by a single wheel for the sake of simplicity.
@@
 	\vspace{-9pt}
-	\caption{\hl{The} %MDPI: Please change the hyphen (-) into minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1".
- inertia tensor can be seen as an ellipsoid. Its principal axes are in the direction of the eigenvectors of the tensor and their lengtha depend on the eigenvalues.}
+	\caption{The %MDPI: Please change the hyphen (-) into minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1".
+ inertia tensor can be seen as an ellipsoid. Its principal axes are in the direction of the eigenvectors of the tensor and their lengths depend on the eigenvalues.}
 	\label{fig:ellipsoid}
@@
 \end{equation}
-\hl{Notice} that in the Elroy's Beanie model, there is a single body that models the complete OCS.
+Notice that in the Elroy's Beanie model, there is a single body that models the complete OCS.
 The inertia introduced in the Elroy's Beanie model has to be split between the two flywheels: $I_{f, zz} = I_f/\left(2\cos(\alpha)\right)$.
@@
 \begin{table}[H]%[t] 1
-	\caption{\hl{Sizes} %MDPI: Please confirm if all the "\cdot" between numbers should be a multiplication sign ("\times" U+00D7).
+	\caption{Sizes %MDPI: Please confirm if all the "\cdot" between numbers should be a multiplication sign ("\times" U+00D7).
  and dynamic parameters of a single flywheel.}
@@
 		\midrule
-		$r$ & $2.20 \cdot 10^{-2}$ & $\mathrm{m}$ \\
-		$R$ & $3.00\cdot 10^{-2}$ & $\mathrm{m}$ \\
-		$h$   & $1.02\cdot 10^{-2}$ & $\mathrm{m}$ \\
-		$\rho$   & $7.86 \cdot 10^{+3}$ & $\mathrm{kg/m^3}$ \\
-		$m$   & $1.02\cdot 10^{-1}$ & $\mathrm{kg}$ \\
-		$I_{xx}, I_{yy}$   & $3.64 \cdot 10^{-5}$ & $\mathrm{kg \cdot m^2}$ \\
-		$I_{zz}$   & $7.11 \cdot 10^{-5}$ & $ \mathrm{kg \cdot m^2}$\\
+		$r$ & $2.20 \times 10^{-2}$ & $\mathrm{m}$ \\
+		$R$ & $3.00\times 10^{-2}$ & $\mathrm{m}$ \\
+		$h$   & $1.02\times 10^{-2}$ & $\mathrm{m}$ \\
+		$\rho$   & $7.86 \times 10^{+3}$ & $\mathrm{kg/m^3}$ \\
+		$m$   & $1.02\times 10^{-1}$ & $\mathrm{kg}$ \\
+		$I_{xx}, I_{yy}$   & $3.64 \times 10^{-5}$ & $\mathrm{kg \cdot m^2}$ \\
+		$I_{zz}$   & $7.11 \times 10^{-5}$ & $ \mathrm{kg \cdot m^2}$\\
 		\bottomrule	\end{tabularx}
@@
 \end{equation}
-\hl{From} this expression, the moment on the base caused by the acceleration of the flywheels \MF{is defined}, and can be used as a feedback torque $\bm{\tau}_{fb}$:
+From this expression, the moment on the base caused by the acceleration of the flywheels \MF{is defined}, and can be used as a feedback torque $\bm{\tau}_{fb}$:
 \begin{equation}
@@
 \end{equation}
-\hl{Proportional} and derivative action $\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}_r}$ \MF{is used}, where $\bm{K}_p$ and $\bm{K}_d$ are diagonal and positive definite gain matrices for the error in attitude and angular velocity.
+Proportional and derivative action $\bm{K}_p \bm{e} + \bm{K}_d \dot{\bm{e}_r}$ \MF{is used}, where $\bm{K}_p$ and $\bm{K}_d$ are diagonal and positive definite gain matrices for the error in attitude and angular velocity.
 The orientation error $\bm{e} \in SO(3)$ requires that the algebra of the special rotational group is computed. To avoid singular configurations, orientations \MF{are represented} with quaternions.
@@
 \end{equation}
-\hl{Projecting} the moment $\bm{\tau}_{fb}$ onto the flywheel axes with $\bm{C}^T$, the control action $\bm{u}$ \MF{yields}
+Projecting the moment $\bm{\tau}_{fb}$ onto the flywheel axes with $\bm{C}^T$, the control action $\bm{u}$ \MF{yields}
 \begin{equation}
@@
 \end{equation}
-\hl{The} OCS was commanded to track base orientation references using the control law introduced in Section \ref{subsec:control_law} \MF{coded in Python}.
-({\hl{The} %MDPI: Footnote is not permitted in our journal.
+The OCS was commanded to track base orientation references using the control law introduced in Section \ref{subsec:control_law} \MF{coded in Python}.
+(The %MDPI: Footnote is not permitted in our journal.
 We include this paragraph to the maintext.
 Please confirm.
- video, with all the simulations, is available at web link \url{https://www.dropbox.com/scl/fo/wevvl0zvfp8gofie6x7rw/h?dl=0&rlkey=pap76cr813a0skrvrxl7qgrb9}  (\hl{accessed} %MDPI: Please add the access date (Format: Date Month Year).
+video, with all the simulations, is available at web link \url{https://www.dropbox.com/s/6py2avxj2f9kksn/video.mp4?dl=0}, accessed %MDPI: Please add the access date (Format: Date Month Year).
 e.g., (accessed on 1 January 2020).
- on).)}
+ on 18 January 2023).
 
@@
 \newblock In Proceedings of the 2013 IEEE International Conference on Robotics
-  and Automation, \hl{Karlsruhe, Germany, 6--10 May 2013;} %MDPI: Newly added information. Please confirm. The following highlights are the same.
- IEEE: \hl{Piscataway, NJ, USA,}  2013; pp.~2554--2561.
+  and Automation, Karlsruhe, Germany, 6--10 May 2013; %MDPI: Newly added information. Please confirm. The following highlights are the same.
+ IEEE: Piscataway, NJ, USA,  2013; pp.~2554--2561.
 
@@
   external disturbances and variable perception quality. In {\em Advances in
-  Robotics Research: From Lab to Market}; Springer: \hl{Berlin/Heidelberg, Germany,} %We added the location of publisher. Please confirm
+  Robotics Research: From Lab to Market}; Springer: Berlin/Heidelberg, Germany, %We added the location of publisher. Please confirm
   2020; pp. 165--209.
@@
 \newblock Frontal plane stabilization and hopping with a 2DOF tail.~In Proceedings of the 2016 IEEE/RSJ International Conference on
-  Intelligent Robots and Systems (IROS), \hl{Daejeon, Korea, 9--14 October 2016}; IEEE: \hl{\mbox{Piscataway}, NJ, USA,}  2016; pp.~567--573.
+  Intelligent Robots and Systems (IROS), Daejeon, Korea, 9--14 October 2016; IEEE: \mbox{Piscataway}, NJ, USA,  2016; pp.~567--573.
 
@@
 \newblock Tail assisted dynamic self righting. In {\em Adaptive Mobile
-  Robotics}; World Scientific: \hl{Singapore},
+  Robotics}; World Scientific: Singapore,
   2012; pp. 611--620.
@@
   Morphable Inertial Tail for Somersault.
-\newblock {\em IEEE/ASME Trans. Mechatron.} {\bf 2022}, \hl{\emph{7}, 5072--5083}.
+\newblock {\em IEEE/ASME Trans. Mechatron.} {\bf 2022}, \emph{7}, 5072--5083.
 
@@
 \newblock In Proceedings of the 2020 8th IEEE RAS/EMBS International Conference
-  for Biomedical Robotics and Biomechatronics (BioRob),  \hl{New York City, NY, USA, 29 November--1 December 2020}; IEEE: \hl{Piscataway, NJ, USA},  2020; \mbox{pp.
+  for Biomedical Robotics and Biomechatronics (BioRob),  New York City, NY, USA, 29 November--1 December 2020; IEEE: Piscataway, NJ, USA,  2020; \mbox{pp.
   132--139.}
@@
 \newblock In Proceedings of the 2021 IEEE International Conference on Robotics
-  and Automation (ICRA), \hl{Xi'an, China, 30 May--5 June 2021}; IEEE: \hl{Piscataway, NJ, USA},  2021; pp.~2979--2985.
+  and Automation (ICRA), Xi'an, China, 30 May--5 June 2021; IEEE: Piscataway, NJ, USA,  2021; pp.~2979--2985.
 
@@
 \newblock In Proceedings of the 2022 International Conference on Robotics and
-  Automation (ICRA), \hl{Philadelphia , PA, USA, 23--27 May 2022}; IEEE: \hl{Piscataway, NJ, USA},  2022; pp.~4635--4641.
+  Automation (ICRA), Philadelphia , PA, USA, 23--27 May 2022; IEEE:Piscataway, NJ, USA,  2022; pp.~4635--4641.
 
@@
 \newblock In Proceedings of the  IEEE International Conference on
-  Robotics and Automation, \hl{Minneapolis, MN, USA, 22--28 April 1996}; IEEE: \hl{Piscataway, NJ, USA},  1996, Vol.~4; pp.~3658--3663.
+  Robotics and Automation, Minneapolis, MN, USA, 22--28 April 1996; IEEE: Piscataway, NJ, USA,  1996, Vol.~4; pp.~3658--3663.
 
@@
 \newblock In Proceedings of the 2021 IEEE International Conference on Robotics
-  and Automation (ICRA), \hl{30 May--5 June 2021, Xi'an, China}; IEEE: \hl{Piscataway, NJ, USA},  2021; pp.~4976--4982.
+  and Automation (ICRA), 30 May--5 June 2021, Xi'an, China; IEEE: Piscataway, NJ, USA,  2021; pp.~4976--4982.
 
@@
 \newblock In Proceedings of the 2009 4th International Conference on Recent
-  Advances in Space Technologies, \hl{Istanbul, Turkey, 11--13June 2009}; IEEE: \hl{Piscataway, NJ, USA},  2009; pp.~778--783.
+  Advances in Space Technologies, Istanbul, Turkey, 11--13 June 2009; IEEE: Piscataway, NJ, USA,  2009; pp.~778--783.
 
@@
 \newblock In Proceedings of the 2020 IEEE/RSJ International Conference on
-  Intelligent Robots and Systems (IROS), \hl{Las~Vegas, Nevada, USA, 25--29 October 2020}; IEEE: \hl{Piscataway, NJ, USA},  2020; pp.~3510--3517.
+  Intelligent Robots and Systems (IROS), Las~Vegas, Nevada, USA, 25--29 October 2020; IEEE: Piscataway, NJ, USA,  2020; pp.~3510--3517.
 
@@
 \newblock In Proceedings of the 2019 IEEE/RSJ International Conference on
-  Intelligent Robots and Systems (IROS), \hl{Macau, 3--8 November 2019}; IEEE: \hl{Piscataway, NJ, USA},  2019; pp.~5459--5466.
+  Intelligent Robots and Systems (IROS),Macau, 3--8 November 2019; IEEE: Piscataway, NJ, USA,  2019; pp.~5459--5466.
 
@@
 \newblock In Proceedings of the 2016 IEEE International Conference on Robotics
-  and Automation (ICRA), \hl{Stockholm, Sweden, 16--21 May 2016;} IEEE: \hl{Piscataway, NJ, USA},  2016; pp.~3590--3595.
+  and Automation (ICRA), Stockholm, Sweden, 16--21 May 2016; IEEE: Piscataway, NJ, USA,  2016; pp.~3590--3595.
 
@@
   tails.
-\newblock In Proceedings of the 2015 European Control Conference (ECC), \hl{Linz, Austria, 15--17 July 2015}; IEEE: \hl{Piscataway, NJ, USA},
-  2015; pp.~753--758.
+\newblock In Proceedings of the 2015 European Control Conference (ECC), Linz, Austria, 15--17 July 2015; IEEE: Piscataway, NJ, USA,2015; pp.~753--758.
 
@@
 \newblock In Proceedings of the 2012 IEEE/RSJ International Conference on
-  Intelligent Robots and Systems, \hl{Vilamoura-Algarve, Portugal, 7--12 October 2012}; IEEE: \hl{Piscataway, NJ, USA},  2012; pp.~3722--3727.
+  Intelligent Robots and Systems, Vilamoura-Algarve, Portugal, 7--12 October 2012; IEEE: Piscataway, NJ, USA,  2012; pp.~3722--3727.
 
@@
 {Open Dynamic Robot Initiative (ODRI)}.
-\newblock \emph{Open Robot Actuator Hardware}\hl{; .} %MDPI: Please add the name of the publisher and the location of it.
+\newblock \url{https://open-dynamic-robot-initiative.github.io/} %MDPI: Please add the name of the publisher and the location of it.
 
@@
 \newblock Modeling and control of legged robots. In {\em Springer Handbook of
-  Robotics}; Springer: \hl{Berlin/Heidelberg, Germany},  2016; pp. 1203--1234.
+  Robotics}; Springer: Berlin/Heidelberg, Germany,  2016; pp. 1203--1234.
 
@@
 \newblock ROS: An Open-Source Robot Operating System.
-\newblock In Proceedings of the ICRA Workshop Open Source Softw, Kobe, Japan,  \hl{12--17 May} 2009; Volume 5, p.~5.
+\newblock In Proceedings of the ICRA Workshop Open Source Softw, Kobe, Japan,  12--17 May 2009; Volume 5, p.~5.
 
@@
 \newblock In Proceedings of the 2004 IEEE/RSJ International Conference on
-  Intelligent Robots and Systems (IROS) (IEEE Cat. No. 04CH37566), \hl{Sendai, Japan, 28 September 2004--2 October 2004}; IEEE: \hl{Piscataway, NJ, USA},  2004;
+  Intelligent Robots and Systems (IROS) (IEEE Cat. No. 04CH37566), Sendai, Japan, 28 September 2004--2 October 2004; IEEE: Piscataway, NJ, USA,  2004;
   Volume~3; pp.~2149--2154.

### 2023-01-18 mdpi2022roscia_v7/matlab/inertiaEllipsoid/arrow3d.m

@@ -1,0 +1,155 @@
+function h = arrow3d(p1,vec,varargin)
+%mArrow3 - plot a 3D arrow as patch object (cylinder+cone)
+%
+% syntax:   h = mArrow3(p1,p2)
+%           h = mArrow3(p1,p2,'propertyName',propertyValue,...)
+%
+% with:     p1:         tail location
+%           p2:         vector you want to plot
+%           properties: 'color':      color according to MATLAB specification
+%                                     (see MATLAB help item 'ColorSpec')
+%                       'stemWidth':  width of the line
+%                       'tipWidth':   width of the cone                       
+%
+%           Additionally, you can specify any patch object properties. (For
+%           example, you can make the arrow semitransparent by using
+%           'facealpha'.)
+%                       
+% example1: h = mArrow3([0 0 0],[1 1 1])
+%           (Draws an arrow from [0 0 0] to [1 1 1] with default properties.)
+%
+% example2: h = mArrow3([0 0 0],[1 1 1],'color','red','stemWidth',0.02,'facealpha',0.5)
+%           (Draws a red semitransparent arrow with a stem width of 0.02 units.)
+%
+% hint:     use light to achieve 3D impression
+%
+
+propertyNames = {'edgeColor'};
+propertyValues = {'none'};    
+
+%% evaluate property specifications
+for argno = 1:2:nargin-2
+    switch varargin{argno}
+        case 'color'
+            propertyNames = {propertyNames{:},'facecolor'};
+            propertyValues = {propertyValues{:},varargin{argno+1}};
+        case 'stemWidth'
+            if isreal(varargin{argno+1})
+                stemWidth = varargin{argno+1};
+            else
+                warning('mArrow3:stemWidth','stemWidth must be a real number');
+            end
+        case 'tipWidth'
+            if isreal(varargin{argno+1})
+                tipWidth = varargin{argno+1};
+            else
+                warning('mArrow3:tipWidth','tipWidth must be a real number');
[…]

### 2023-01-18 mdpi2022roscia_v7/matlab/inertiaEllipsoid/computeInflateFactor.m

@@ -1,0 +1,17 @@
+function inflateFactor = computeInflateFactor(A,longestAxisLength)
+
+
+    [eigvecs, eigvals] = eig(A);
+    eigs= diag(eigvals);
+    % Position (1,2 or 3) of the minimum inertia
+    minpos = find(eigs == min(eigs)); 
+
+     
+    % the radii are the singular value of a rectangula matrix J+ and they are the sqrt(eigenvalues of the square matrix J+'J+) 
+    radii(1) = sqrt(1/eigs(1));
+    radii(2) = sqrt(1/eigs(2));
+    radii(3) = sqrt(1/eigs(3));
+        
+    % Resize the radii to match user desires
+    inflateFactor = (radii(minpos)*2) / longestAxisLength;
+end

### 2023-01-18 mdpi2022roscia_v7/matlab/inertiaEllipsoid/plotEigenvectorsAndEllipsoid.m

@@ -1,0 +1,11 @@
+clear all
+close all
+I = diag([0.00578574, 0.01938108,0.02476124]);
+inflateFactor = computeInflateFactor( I, 0.2);
+plotEllispoidInertia(I,[0;0;0],  inflateFactor)
+
+% save the plot
+set(gcf, 'Paperunits' , 'centimeters')
+set(gcf, 'PaperSize', [45 19]);
+set(gcf, 'PaperPosition', [0 0 45 19]);
+print(gcf, '-dpdf','../../figures/inertiaEllipsoid.pdf')

### 2023-01-18 mdpi2022roscia_v7/matlab/inertiaEllipsoid/plotEllispoidInertia.m

@@ -1,0 +1,56 @@
+ function  plotEllispoidInertia(A, center, inflateFactor)
+    
+ 
+    [eigvecs, eigvals] = eig(A);
+         
+    % The eigenvalues are the diagonal elements of the returned matrix
+    eigs= diag(eigvals);
+    if(any(eigs < 0))
+        error('Looks like the inertia tensor is not positive definite');
+    end
+
+    % The equation of the inertia ellipsoid is
+    %    X x^2 + Y y^2 + Z z^2 = 1
+    % Matlab computes an ellipsoid with this equation
+    %    (x^2)/(a^2) + (y^2)/(b^2) + (z^2)/(c^2) = 1
+    % I call 'a', 'b' and 'c' "radii"; they are half of the ellipsoid axes
+    %size of axis are the singularvalues = square root of the eigen values
+    %of (I)^-1
+
+    % the radii are the singular value of a rectangula matrix J+ and they are the sqrt(1/eigenvalues of the square matrix J+'J+) 
+    radii(1) = sqrt(1/eigs(1));
+    radii(2) = sqrt(1/eigs(2));
+    radii(3) = sqrt(1/eigs(3));
+        
+    radii= radii/inflateFactor;
+    %the ellispoid is created 
+    [X,Y,Z] = ellipsoid(0,0,0, radii(1),radii(2),radii(3));
+
+    % Rotate the coordinates of all the points to the base frame (they are
+    % generated in the ellipsoidframe), and translate them to p3
+    for r=1:size(X,1)
+        for c=1:size(X,2)
+            tmp =  eigvecs* [X(r,c); Y(r,c); Z(r,c);]; %%rotate back to base coordinates by mult for the eigvecs vatrix that contains the axis of the ellipsoid in base coords
+            X(r,c) = tmp(1) + center(1); %translate to end effector
+            Y(r,c) = tmp(2) + center(2);
+            Z(r,c) = tmp(3) + center(3);
+        end
+    end
+    surf(X, Y, Z, 'FaceColor',[0.8 0.1 0.1],'FaceAlpha', 0.5, 'EdgeAlpha', 0.1 ,'EdgeColor','none');hold on;
+    colormap jet
+    shading interp;
+    light;
+    lighting phong;
+    
+
+    %plot axes 
[…]
